### 2022-08-12 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
First draft

---
title: "Meta-Analysis of Variation in Sport and Exercise Science"
subtitle: "Examples of Application Within Resistance Training Research"
author: 
  - James Steele
  - James Fisher
  - Dave Smith
  - Andrew Vigotsky
  - Brad Schoenfeld
  - Yefeng Yang
  - Shinichi Nakagawa
date: "2022-08-04"
abstract: |
  TO COMPLETE.
output: bookdown::pdf_document2
bibliography: references.bib
toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
options(knitr.kable.NA = '')
```
```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Load relevant packages
library(here)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# load the raw data to the environment
Data <- read.csv(here::here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))

# load the summary dataframes to the environment
load(here("models/SMD_SDir_logVR"))
load(here("models/mods_SMD_logCVR"))

```
# Introduction
Though the quantitative synthesis of results across studies has existed since the 17th century [@plackett_studies_1958], the modern day term "meta-analysis" as it is now referred to was coined by Gene Glass in [-@glass_primary_1976]. Since that time use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@hagger_meta-analysis_2022]. One area of sport and exercise science which has comprised a considerable proportion of that growth appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)).
```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976"}

load(here("plots/trends_plot"))
trends_plot

```
As with many other fields [@nakagawa_meta-analysis_2015; @usui_meta-analysis_2021; @mills_detecting_2021] likely the most familiar aim with the use of meta-analysis, and indeed primary empirical research too, in sport and exercise science is to make comparisons between the means of measurements taken across different categorical grouping variables; for example, the comparison of an intervention group and a control group, or the comparison of one intervention group to another. Indeed, a recent umbrella review [@bernardez-vazquez_resistance_2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables on hypertrophy outcomes all of which focused on the comparison of mean changes between different intervention groups.

Most commonly a magnitude based[^1] effect size statistic [@caldwell_case_2020], the standardised mean difference (SMD), is used to compare means between groups or conditions. This is usually Cohen's *d* [@cohen_statistical_1988], or it's bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007]. The SMD[^2], and its sampling variance, $s^2_{SMD}$ are given by:

\begin{equation}
SMD=\frac{\overline{x}_E - \overline{x}_C}{s_{pooled}}J
(\#eq:SMD-eq)
\end{equation}

\begin{equation}
J=1-\frac{3}{4(n_{C}+n_{E})-2)-1}J
(\#eq:biasJ-eq)
\end{equation}

\begin{equation}
s_{pooled}=\sqrt{\frac{(n_{C}-1)s^2_{C}(n_{E}-1)s^2_{E}+}{n_{C}+n_{E}-2}}
(\#eq:s-pooled-eq)
\end{equation}

\begin{equation}
s^2_{SMD}=\frac{n_{C}+n_{E}}{n_{C}n_{E}}+\frac{SMD^2}{2(n_{E}+n_{C})}
(\#eq:SMDvar-eq)
\end{equation}

where $\overline{x}_C$ and $\overline{x}_E$ are the sample means of the control group (C) and experimental (E) or intervention group respectively, $s_C$ and $s_E$ are the standard deviations of the two groups, $n_C$ and $n_E$ are the sample sizes of the two groups, and $J$ is a bias correction for small sample sizes.

The natural logarithm of the ratio of two means ($\textrm{ln}RR$) is also another effect size statistic that can be used [@curtis_meta-analysis_1998; @hedges_meta-analysis_1999; @lajeunesse_bias_2015]. The lnRR, and its sampling variance, $s^2_{lnRR}$ are given by:

\begin{equation}
\textrm{ln}RR=\textrm{ln}\frac{\overline{x}_E}{\overline{x}_C}
(\#eq:lnRR-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}RR}=\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}+\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}}
(\#eq:lnRRvar-eq)
\end{equation}

In comparison to the SMD, which due to its calculation is affected not only by the difference in means of the two groups C and E, but is also affected by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$. The $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), and only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)). Despite this, use of $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@deb_quantifying_2018] and to our knowledge only one meta-analysis in RT has used this effect size (Swinton et al., in press).

Though the focus in sport and exercise science among other fields has been in estimating the average effects of interventions for both primary research and synthesis through meta-analysis, the field has been aware for some time that responses to interventions do vary potentially on a subgroup or even individual basis. The increased interested in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@hecksteden_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @ross_precision_2019; @swinton_statistical_2018; @hopkins_individual_2015; @kelley_precision_2022; @hrubeniuk_directions_2022; @pickering_non-responders_2019]. However, despite the availability of approaches to compare variances between groups, in sports and exercise science this is rarely explored in primary research [@bonafiglia_interindividual_2022] and, though there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022]. Examination of interindividual heterogeneity in response to interventions presents considerable value to researchers and practitioners in sport and exercise science; interventions with low interindividual variation are likely to be widely generalisable, whilst an intervention with high interindividual variation is likely to have effects that are either subgroup or individual specific. The former kind of intervention might be widely applicable across individuals, whilst the latter kind of intervention requires specific research to tease apart subgroup- or participant-by-intervention interactions to enable successful practical application.

Comparison of heterogeneity in responses such as post-scores or change scores to longitudinal interventions are however not the only possible use of statistical methods for comparing variances. For example, in other fields such as ecology there have been calls to shift focus onto the exploration of dispersion of traits between groups in non-experimental or interventional designs [@nakagawa_mean_2012]. Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between participant acute response variation for the purposes of identifying methods[^3] to reduce RT stimulus heterogeneity [@exner_does_2022] as well as meta-analysis exploring between participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperin_accuracy_2022].

Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet the lack of application in research synthesis by way of meta-analysis, we present and discuss effect size approaches and models for meta-analysis of variation. We provide examples of the approaches presented using data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021].

# Effect size statistics for meta-analytic comparisons of variation
Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework. However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; @hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015).

## Standard deviation for individual responses ($SD_{ir}$)
In the context of *precision* or *personalised* approaches to exercise prescription comparison of variation between two groups, C and E, the $SD_{ir}$ has been proposed as an approach to determining the extent to which individual responses manifest [@hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019]. The standard deviation of changes scores (post-intervention scores minus pre-intervention scores) within E reflects the gross combination of a number of sources of variation including actual subject-by-intervention interactions (i.e., actual individual responsiveness), within-subject variability in intervention response (i.e., variability in response to the same intervention administered to the same subject), and random error (i.e., from pre and post measurements; @hecksteden_individual_2015). The standard deviation of changes scores from the C group by contrast only reflects random error [@hecksteden_individual_2015]. As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention that might reflect individual responses. The $SD_{ir}$, and its sampling variance, $s^2_{SD_{ir}}$ are given by:

\begin{equation}
SD_{ir}=\sqrt{s^2_{E} - s^2_{C}}
(\#eq:SDir-eq)
\end{equation}

\begin{equation}
s^2_{SD_{ir}}=2(\frac{s^4_{E}}{n_{E}-1}+\frac{s^4_{C}}{n_{C}-1})
(\#eq:sdIRvar-eq)
\end{equation}

Thus, the $SD_{ir}$ reflects a comparison of the absolute variance in change scores between C and E. Whilst the $SD_{ir}$ has been proposed and used primarily in the context of individual response variation to interventions, it should be noted that this kind of absolute comparison of variance between groups or conditions is not limited to such applications.

## Log ratio of standard deviations ($\textrm{ln}VR$)
A similar effect size statistic for the comparison of absolute variance between groups, and one which has had wide applications in more than just intervention response variability within fields like ecology and evolution, is the $\textrm{ln}VR$ [@nakagawa_meta-analysis_2015]. An unbiased estimator of the natural logarithm of a population standard deviation ($\textrm{ln}\sigma$), and its sampling variance, $s^2_{\textrm{ln}\sigma}$ is given by:

\begin{equation}
\textrm{ln}\hat\sigma=\textrm{ln}s+\frac{1}{2(n-1)}
(\#eq:lnsigma-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}\hat\sigma}=\frac{1}{2(n-1)}
(\#eq:lnsigmavar-eq)
\end{equation}

where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$. Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups such as C and E, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:

\begin{equation}
\textrm{ln}VR=\textrm{ln}(\frac{s_{E}}{s_{C}}+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnVR-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}VR}=\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnVRvar-eq)
\end{equation}

However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparisongroup (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$). In this case it is likely that the standard deviation will be larger in the group with the larger mean (e.g., $s_{E}$ is larger than $s_{C}$). This mean-variance relationship is common for many variables and datasets[^4] and we will provide examples of this below. They also assume constant measurement error over the range of values for the mean which can impact their utility for examining response variation [@tenan_comment_2020].

## Log ratio of coefficient of variation ($\textrm{ln}CVR$)
The coefficient of variation is the ration of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present. The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups. Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), the $\textrm{ln}CVR$ is given by:

\begin{equation}
\textrm{ln}CVR=\textrm{ln}(\frac{CV_{E}}{CV_{C}}+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnCVR-eq)
\end{equation}

where $CV_{E}$ and $CV_{C}$ are ${s_{E}}/{\overline{x}_{E}}$ and ${s_{C}}/{\overline{x}_{C}}$ respectively. Nakagawa et al. [-@nakagawa_meta-analysis_2015] derived the sampling variance, s^2_{\textrm{ln}CVR}, as:

\begin{equation}
\begin{split}
s^2_{\textrm{ln}CVR}=\frac{s^2_{C}}{n_{C}\overline{x}_{C}}+\frac{1}{2(n_{C}-1)}-2\rho_{\textrm{ln}\overline{x}_{C},\textrm{ln}{s}_{C}}\sqrt{\frac{s^2_{C}}{n_{C}\overline{x}_{C}}+\frac{1}{2(n_{C}-1)}} \\
+ \frac{s^2_{E}}{n_{E}\overline{x}_{E}}+\frac{1}{2(n_{E}-1)}-2\rho_{\textrm{ln}\overline{x}_{E},\textrm{ln}{s}_{E}}\sqrt{\frac{s^2_{E}}{n_{E}\overline{x}_{E}}+\frac{1}{2(n_{E}-1)}}
\end{split}
(\#eq:lnCVRvar-eq)
\end{equation}

where $\rho_{\textrm{ln}\overline{x}_{C},\textrm{ln}{s}_{C}}$ and $\rho_{\textrm{ln}\overline{x}_{E},\textrm{ln}{s}_{E}}$ are the correlatios between the means and the standard deviation in the C and E groups respectively on the log scale across studies.

# Examples using resistance training studies
As noted, the examples presented used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021]. Here we have used their list of included studies and re-extracted data from 111 of these[^5]. All analysis examples were performed in using R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauer_conducting_2010]. NOTE - ADD LINK TO ANALYSIS SCRIPT AND DATA

Polito et al. [-@polito_moderators_2021] conducted a systematic review and meta-analysis of randomised trials including a RT intervention where a non-training control group comparison group was included. Their analysis focused upon the SMD between the RT interventions and control groups from the studies included, with both overall effect estimate and moderator analyses performed. Given that Polito et al. [-@polito_moderators_2021] included only studies with non-training control groups, their study selection enables the examination of interindividual response variation by means of comparing the variances in change scores between the RT intervention and control groups. Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies. Table \@ref(tab:summary-characteristics-tab).

```{r sample-size-tab, message=FALSE, warning=FALSE, echo=FALSE}

load(here("models/sample_sizes"))

knitr::kable(
  sample_sizes,
  caption = 'Sample sizes for resistance training (RT) and non training control (CON) groups for dataset.'
)

```
```{r summary-characteristics-tab, message=FALSE, warning=FALSE, echo=FALSE}

load(here("models/summary_table"))

knitr::kable(
  summary_table,
  caption = 'Summary of study and participant characteristics (median [IQR], number of groups [%]).'
)

```
## Detecting the presence of interindividual response variation to resistance training intervention
First we conduct a traditional SMD based effect size meta-analysis to explore the effects of RT compared to control for strength outcomes and hypertrophy (i.e., muscle mass/size) outcomes. Polito et al. [-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple effects within groups within studies) and so a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group[^6]. We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in in responses differed between RT interventions and controls. A positive SMD would indicate that RT interventions increase outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses compared to controls (i.e., suggests the presence of interindividual response variation).

```{r forest-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes"}

# "Caterpillar plots of SMD effect sizes for strength and hypertrophy outcomes. Note, each estimate and 95% confidence interval is a single observed effect size within each group within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"

load(here("plots/forest_strength"))
load(here("plots/forest_hypertrophy"))
library(patchwork)
(forest_strength / forest_hypertrophy) +
  plot_annotation(tag_levels = "A")

```

The pattern of results from our models examining SMDs (figure \@ref(fig:forest-plot)) were similar to those reported by Polito et al. [-@polito_moderators_2021] albeit with slightly lower estimates for both outcome types; this possibly being due to our use of a multilevel mixed-effects meta-analysis model which allowed for each effect size included to be more appropriately weighted. As might be expected, in comparison to non-training controls the RT interventions produced increases in strength (SMD = `r SMD_SDir_logVR$Estimate[1]` [95%CI: `r SMD_SDir_logVR$Lower[1]` to `r SMD_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[1]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[1]`) and hypertrophy outcomes (SMD = `r SMD_SDir_logVR$Estimate[4]` [95%CI: `r SMD_SDir_logVR$Lower[4]` to `r SMD_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[4]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[4]`). Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-plot)) were fairly wide and relative heterogeneity was fairly high coming mostly from between study variance[^7].

In addition to the SMD results, both the $SD_{ir}$ and $\textrm{ln}VR$ were also positive for both strength ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[2]` [95%CI: `r SMD_SDir_logVR$Lower[2]` to `r SMD_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[2]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[2]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[3]` [95%CI: `r SMD_SDir_logVR$Lower[3]` to `r SMD_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[3]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[3]`) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[5]` [95%CI: `r SMD_SDir_logVR$Lower[5]` to `r SMD_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[5]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[5]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[6]` [95%CI: `r SMD_SDir_logVR$Lower[6]` to `r SMD_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[6]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[6]`) suggesting that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of of interindividual response variation. This might support the previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* apart from the random error). However, as noted both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean. As we have seen from the SMD analysis RT interventions increase mean scores. Thus, if there is a mean-variance relationship in the data an increase in the mean alone may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present. The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic. First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a model that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$.

## Mean-variance relationships in muscular strength and hypertrophy
With meta-analytic models of variation we are not limited to solely exploring variation in responses to interventions. We can explore the relationships between variance in a number of outcomes and the impact of certain predictors on this. For example, as noted one possible predictor of variance is the mean itself. As such, we can model variance as the response itself. The standard deviation is however bounded at zero and so in many cases it may not conform to assumptions of normality. Therefore, we instead can use $\textrm{ln}\hat\sigma$ which is unbounded. In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al. [-@polito_moderators_2021] using a multilevel mixed-effects model.

```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes"}

load(here("plots/mean_variance_pre_plots"))

mean_variance_pre_plots

```

As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) & (B), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$). This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylor_aggregation_1961].

\begin{equation}
s^2=a\overline{x}^b
(\#eq:taylors-eq)
\end{equation}

where $a$ and $b$ are some constants. When this relationship (equation \@ref(eq:taylors-eq)) holds, under most circumstances the standard deviation is not proportional to the mean. However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear:

\begin{equation}
2\textrm{ln}s=\textrm{ln}a+b\textrm{ln}\overline{x}
(\#eq:taylorsln-eq)
\end{equation}

Figure \@ref(fig:mean-variance-pre-plot)(C) & (D) shows that the relationship between the mean and variance on the log scale better meets the assumption of normality. Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from outcomes over multiple groups and studies we can also estimate this relationship using the following model:

\begin{equation}
\textrm{ln}\hat\sigma_{ij}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j})+\beta_{1}\textrm{ln}\overline{x}_{ij}+\epsilon_{ij}+m_{ij} 
(\#eq:lnm-lns-model-eq)
\end{equation}

where $\textrm{ln}\hat\sigma_{ij}$ is the effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ij}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, and $\epsilon_{ij}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ij}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ij}}$. Additional predictor terms could be added to this model; for example, we could model a categorical variable for the outcome type and include $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome groups), and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group. In this case $\tau_{i}$ and $\varphi_{i}$, and $\tau_{j}$ and $\varphi_{j}$ are assumed to have multivariate normal distributions with the following variance-covariance structure:

\begin{equation}
\binom{\tau}{\varphi}=(\binom{0}{0},
\left(\begin{array}{cc} 
\sigma^2_{\tau} & \rho\sigma_{\tau}\sigma_{\varphi}\\
\rho\sigma_{\tau}\sigma_{\varphi} & \sigma^2_{\varphi}
\end{array}\right))
(\#eq:lnCVR-model-varcov-eq)
\end{equation}

Figure \@ref(fig:model-mean-variance-pre-plot) shows this model fit visually. Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight systematically greater degree of variance in strength measures compared to hypertrophy for a given mean score.

```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores"}

# "Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores. Note, fitted lines are the regression slopes for strength and hypertophy outcomes each with their 95% confidence interval band."

load(here("plots/model_mean_variance_pre_plots"))

model_mean_variance_pre_plots

```

The presence of Taylor's law type relationships should be examined in datasets prior to deciding on which variance effect size statistic should be employed. Returning to the context of interindividual response variation to interventions, the presence of a mean-variance relationship in the data would imply that we cannot rely on absolute comparisons of variance (i.e., $SD_{ir}$ or $\textrm{ln}VR$) to determine whether interindividual response variation is actually present. So we should also explore this for the change-scores in the RT and control groups and determing the appropriate effects to explore.

## Reanalysis of interindividual response variation using $lnCVR$ and the random slope mixed effects model
As can be seen in figure \@ref(fig:mean-variance-delta-plot) there is a mean-variance relationship in the change score data whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.

```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes"}

load(here("plots/mean_variance_delta_plots"))

mean_variance_delta_plots

```

The $\textrm{ln}CVR$ can be used to overcome this issue though. Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group and before using the $\textrm{ln}CVR$ as the response variable leads to different conclusions compared to absolute variance comparisons. The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[88]`).

There is however a potential limitation for the $\textrm{ln}CVR$ also that may need to be considered. Firstly, it is limited to the use of ratio scale data (which is not the case for the $\textrm{ln}\hat\sigma$ or $\textrm{ln}VR$). But secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean. Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative approach that is equivalent may be more appropriate.

The multilevel mixed-effects meta-analysis model using $\textrm{ln}CVR$ as used above can be written as follows:

\begin{equation}
\textrm{ln}CVR_{ij}=\mu+\tau_{i}+\tau_{j}+m_{ij} 
(\#eq:lnCVR-model-eq)
\end{equation}

where $\textrm{ln}CVR_{ij}$ is each effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ij}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ij}}$. Instead, we can use a version of the random slope model described above (equation \@ref(eq:lnm-lns-model-eq) and following paragraph) to compare the variability between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$. In this case, the categorical variable for the outcome type is instead swapped for the group and the new model term included becomes $(\beta_{2} + \varphi_{i} + \varphi_{j})Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group.

```{r, message=FALSE,warning=FALSE,echo=FALSE}
# load the summary dataframes to the environment
load(here("models/RobuEstMultiLevelModel_log_mean_mod_strength"))
load(here("models/RobuEstMultiLevelModel_log_mean_mod_hypertrophy"))

```

Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)) we fit this model to the dataset. The results were however largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.ub[3],2)`]). See figure \@ref(fig:model-mean-variance-delta-plot).

```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores"}

# "Meta-analytic scatter plot of the log mean and log standard deviation of change scores. Note, fitted lines are the regression slopes for control group (CON) and resistance training intervention (RT) groups each with their 95% confidence interval band."

load(here("plots/model_mean_variance_delta_plots"))

model_mean_variance_delta_plots

```

The reason for this apparent reduction in variation after introduction of an RT intervention is not necessarily discernible from this analysis. Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; [@halliday_resistance_2017]), or a ceiling effect on change (i.e., plateau in response; [@steele_long-term_2022]) has a constraining effect [@cortes_martinez_constant_2021]. Hopefully it is clear from the regression model here where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework. Models can be fit to explore not only how study and group level characteristics moderate effect size estimates when considering not only SMDs, but also when considering the variance based effect size statistics and models employed in this article[^8].

# Discussion
Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis. The examples we presented used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021], and so hopefully are more relatable for researchers in sport and exercise sciences.

Given the considerable interest in *precision* or *personalised* exercise prescription in particular, it is of particular interest to note the different conclusions drawn dependent upon the approach taken to determining from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present. Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely increased variance above random error, suggesting the presence of inter-individual response variation. In the case of RT interventions there is evidently an average intervention effect which is positive and as such this result might merely imply that while all likely benefit, some benefit much moreso than others. Exploring variation however even in the absence of average intervention effects might still be important as those with large enough variance could imply that the intervention is at least beneficial to some [@usui_meta-analysis_2021]. Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of RT interventions to maximise strength and hypertrophy outcomes from our example.

However, much like that seen in cross-sectional pre-intervention scores here and most physical and biological systems, change scores demonstrated a mean-variance relationship in addition to heteroskedasticity. The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ in addition to the random slope model of $\textrm{ln}\hat\sigma$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance implying that there is likely little to no inter-individual response variation to explain. Interventions, such as RT interventions explored here, which induce both meaningful SMDs and also show little evidence suggestive of interindividual variation are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usui_meta-analysis_2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortes_martinez_constant_2021]. Interventions such as these are valuable for the simplification of guidelines and recommendations. For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to recommendations [@steele_higher_2017; @steele_long-term_2022].

# Conclusion
Embracing variability and focusing on more than merely the mean differences between groups or conditions, such as intervention and control comparisons, has the potential to inform experimental design and lead to changes in both the approach and direction of follow-up studies. Where considerable differences in variation are present, this suggests a meaningful line of research should be aimed at identifying subgroup- or participant-by-intervention interactions. Where variance differences are limited this suggests that translational work towards generalisable implementation might be most meaningful. Future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utlising the established approaches [@nakagawa_meta-analysis_2015; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @usui_meta-analysis_2021; @mills_detecting_2021] presented here and echoing the efforts of other recent work [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].

# References

<div id="refs"></div>

[^1]: Though notably not all meta-analyses use *magnitude based* effect sizes. Indeed some explicitly use what Caldwell and Vigotsky term *signal-to-noise* effect sizes (e.g., @heidel_machines_2022).

[^2]: We will refer to both merely as the SMD throughout the manuscript for simplicity.

[^3]: Exploration of methodological approaches and their impact on heterogeneity have also been explored in preclinical research [@usui_meta-analysis_2021].

[^4]: For one clear example, see figure 1A in Vigostky et al. [-@vigotsky_improbable_2020] who show that the mean and standard deviation for baseline strength values typically scale with one another across most studies.

[^5]: The authors of the meta-analysis did not make their extracted data openly available, not respond to our request for the extracted data. Further, their original analysis included 119 studies however we were unable to extract data for our analyses from 8 of these for a variety of reasons (e.g., only percentage change data was reported, no standard deviations for control groups reported).

[^6]: Data was coded such that study and group had explicit nesting.

[^7]: Note, we also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases. There did not appear to be any obvious small study bias in the dataset (ADD LINK TO SUPPLEMENTARY)

[^8]: See supplementary materials (ADD LINK TO SUPPLEMENTARY) for model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continous predictors for both strength and hypertrophy outcomes. There were no obvious moderators of $\textrm{ln}CVR$ in particular.

### 2022-08-12 RT Response Heterogeneity.R

@@ -1,824 +1,0 @@
-### Refit everything in brms aswell ###
-
-
-### To save trying to get hold of all the data for these studies, conduct a sensitivity analysis
-# those where it was possible to directly calculate the change score standard deviation vs those where it is estimated
-
-# Need to sort out SMD calculations for those with pre-post, and those with pre-change
-
-# also need to now add in paired logVR and logCVR calculations
-
-            ### All models to run
-            
-            # Main treatment effects - both as d/g and as RR (on % improvement scale)
-            # logVR for delta sd
-            # But mean-variance relationship can be an issue
-            # Show example using the reps to failure dataset
-            # Show relationships between m/sd and log(m)/log(sd) for deltas
-            # logCVR for delta sd
-            # Model log(sd) with log(mean)+group as moderators and random slopes by log(mean) with random intercepts for study/group/es and condition
-            # add size to points on meta-scatter
-            # logVR and logCVR for post_m (assume randomisation deals with baseline imbalance, but maybe check with pre-scores as moderator)
-            # Introduce meta-regression of intervention characteristics on variance
-
-library(metafor)
-library(scales)
-library(tidyverse)
-library(patchwork)
-library(rms)
-library(lspline)
-
-# Read csv as data frame into environment - Note: change source address
-Data <- read.csv("C:/Users/James/Dropbox/Research/Response Heterogeneity - RT/RT Extracted Data.csv")
-
-# Calculate pre-post SDs from SEs
-Data$RT_pre_sd <- ifelse(is.na(Data$RT_pre_se), Data$RT_pre_sd, Data$RT_pre_se * sqrt(Data$RT_n))
-Data$CON_pre_sd <- ifelse(is.na(Data$CON_pre_se), Data$CON_pre_sd, Data$CON_pre_se * sqrt(Data$CON_n))
-Data$RT_post_sd <- ifelse(is.na(Data$RT_post_se), Data$RT_post_sd, Data$RT_post_se * sqrt(Data$RT_n))
[…]

### 2022-08-12 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Push update to syn to OSF for links

@@
 library(here)
+library(kableExtra)
 ```
@@
 One area of sport and exercise science which has comprised a considerable proportion of that growth appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)).
+Given such growth, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
 ```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976"}
@@
 ```
-As with many other fields [@nakagawa_meta-analysis_2015; @usui_meta-analysis_2021; @mills_detecting_2021] likely the most familiar aim with the use of meta-analysis, and indeed primary empirical research too, in sport and exercise science is to make comparisons between the means of measurements taken across different categorical grouping variables; for example, the comparison of an intervention group and a control group, or the comparison of one intervention group to another.
-Indeed, a recent umbrella review [@bernardez-vazquez_resistance_2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables on hypertrophy outcomes all of which focused on the comparison of mean changes between different intervention groups.
+As with many other fields [@nakagawa_meta-analysis_2015; @usui_meta-analysis_2021; @mills_detecting_2021] likely the most familiar aim with the use of meta-analysis, and indeed primary empirical research too, in sport and exercise science is to make comparisons between the means of measurements taken across different categorical grouping variables; for example, the comparison of an intervention group and a control group, the comparison of one intervention group to another, or comparison between non-manipulated categories such as biological sex.
+Indeed, a recent umbrella review [@bernardez-vazquez_resistance_2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes all of which focused on the comparison of mean changes between different intervention groups.
 
 Most commonly a magnitude based[^1] effect size statistic [@caldwell_case_2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
-This is usually Cohen's *d* [@cohen_statistical_1988], or it's bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007].
-The SMD[^2], and its sampling variance, $s^2_{SMD}$ are given by:
+This is usually Cohen's *d* [@cohen_statistical_1988], or it's bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007] [^2].
+The SMD, and its sampling variance, $s^2_{SMD}$ are given by:
 
@@
 \begin{equation}
-J=1-\frac{3}{4(n_{C}+n_{E})-2)-1}J
+J=1-\frac{3}{4(n_{C}+n_{E})-2)-1}
 (\#eq:biasJ-eq)
@@
 \begin{equation}
-s_{pooled}=\sqrt{\frac{(n_{C}-1)s^2_{C}(n_{E}-1)s^2_{E}+}{n_{C}+n_{E}-2}}
+s_{pooled}=\sqrt{\frac{(n_{C}-1)s^2_{C}+(n_{E}-1)s^2_{E}}{n_{C}+n_{E}-2}}
 (\#eq:s-pooled-eq)
@@
 The natural logarithm of the ratio of two means ($\textrm{ln}RR$) is also another effect size statistic that can be used [@curtis_meta-analysis_1998; @hedges_meta-analysis_1999; @lajeunesse_bias_2015].
-The lnRR, and its sampling variance, $s^2_{lnRR}$ are given by:
+The lnRR, and its sampling variance, $s^2_{\textrm{ln}RR}$ are given by:
 
@@
 
-In comparison to the SMD, which due to its calculation is affected not only by the difference in means of the two groups C and E, but is also affected by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$.
-The $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), and only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
+Due to its calculation the SMD is affected not only by the difference in means of the two groups, but is also affected by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$.
+Contrastingly, the $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), which only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
 Despite this, use of $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@deb_quantifying_2018] and to our knowledge only one meta-analysis in RT has used this effect size (Swinton et al., in press).
 
-Though the focus in sport and exercise science among other fields has been in estimating the average effects of interventions for both primary research and synthesis through meta-analysis, the field has been aware for some time that responses to interventions do vary potentially on a subgroup or even individual basis.
+Though the focus in sport and exercise science among other fields has been in estimating the average effects of interventions for both primary research and synthesis through meta-analysis, the field has been aware for some time that responses to certain interventions do vary potentially on a subgroup or even individual basis.
 The increased interested in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@hecksteden_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @ross_precision_2019; @swinton_statistical_2018; @hopkins_individual_2015; @kelley_precision_2022; @hrubeniuk_directions_2022; @pickering_non-responders_2019].
@@
 
-Comparison of heterogeneity in responses such as post-scores or change scores to longitudinal interventions are however not the only possible use of statistical methods for comparing variances.
+Comparison of heterogeneity in responses such as post-scores or change scores to interventions are however not the only possible use of statistical methods for comparing variances.
 For example, in other fields such as ecology there have been calls to shift focus onto the exploration of dispersion of traits between groups in non-experimental or interventional designs [@nakagawa_mean_2012].
@@
 ## Standard deviation for individual responses ($SD_{ir}$)
-In the context of *precision* or *personalised* approaches to exercise prescription comparison of variation between two groups, C and E, the $SD_{ir}$ has been proposed as an approach to determining the extent to which individual responses manifest [@hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019].
-The standard deviation of changes scores (post-intervention scores minus pre-intervention scores) within E reflects the gross combination of a number of sources of variation including actual subject-by-intervention interactions (i.e., actual individual responsiveness), within-subject variability in intervention response (i.e., variability in response to the same intervention administered to the same subject), and random error (i.e., from pre and post measurements; @hecksteden_individual_2015).
-The standard deviation of changes scores from the C group by contrast only reflects random error [@hecksteden_individual_2015].
+In the context of *precision* or *personalised* approaches to exercise prescription comparison of variation between two groups, control (C) and intervention (E), the $SD_{ir}$ has been proposed as an approach to determining the extent to which individual responses manifest [@hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019].
+The standard deviation of changes scores (post-intervention scores minus pre-intervention scores) within the intervention group reflects the gross combination of a number of sources of variation including participant-by-intervention interactions (i.e., actual individual responsiveness or 'trainability'), within-participant variability in intervention response (i.e., variability in response to the same intervention administered to the same participant), and random error (i.e., from pre and post measurements; @hecksteden_individual_2015).
+The standard deviation of change scores from the control group (assuming it is a non-intervention control group and not something like a 'usual-care' group) by contrast only reflects random error [@hecksteden_individual_2015].
 As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention that might reflect individual responses.
@@
 where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$.
-Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups such as C and E, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
+Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups, such as a control (C) and intervention (E), the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
 
 \begin{equation}
-\textrm{ln}VR=\textrm{ln}(\frac{s_{E}}{s_{C}}+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
+\textrm{ln}VR=\textrm{ln}(\frac{s_{E}}{s_{C}})+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
 (\#eq:lnVR-eq)
@@
 \begin{equation}
-s^2_{\textrm{ln}VR}=\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
+s^2_{\textrm{ln}VR}=\frac{1}{2(n_{E}-1)}+\frac{1}{2(n_{C}-1)}
 (\#eq:lnVRvar-eq)
@@
 
-However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparisongroup (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$).
+However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparison group (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$).
 In this case it is likely that the standard deviation will be larger in the group with the larger mean (e.g., $s_{E}$ is larger than $s_{C}$).
@@
 ## Log ratio of coefficient of variation ($\textrm{ln}CVR$)
-The coefficient of variation is the ration of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present.
+The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present.
 The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
@@
 \begin{equation}
-\textrm{ln}CVR=\textrm{ln}(\frac{CV_{E}}{CV_{C}}+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
+\textrm{ln}CVR=\textrm{ln}(\frac{CV_{E}}{CV_{C}})+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
 (\#eq:lnCVR-eq)
@@
 Nakagawa et al.
-[-@nakagawa_meta-analysis_2015] derived the sampling variance, s^2_{\textrm{ln}CVR}, as:
+[-@nakagawa_meta-analysis_2015] derived the sampling variance, $s^2_{\textrm{ln}CVR}$, as:
 
@@
 All analysis examples were performed in using R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauer_conducting_2010].
-NOTE - ADD LINK TO ANALYSIS SCRIPT AND DATA
-
-Polito et al.
-[-@polito_moderators_2021] conducted a systematic review and meta-analysis of randomised trials including a RT intervention where a non-training control group comparison group was included.
-Their analysis focused upon the SMD between the RT interventions and control groups from the studies included, with both overall effect estimate and moderator analyses performed.
-Given that Polito et al.
-[-@polito_moderators_2021] included only studies with non-training control groups, their study selection enables the examination of interindividual response variation by means of comparing the variances in change scores between the RT intervention and control groups.
-Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies.
-Table \@ref(tab:summary-characteristics-tab).
+The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are all available on the Open Science Framework (https://osf.io/2h9ma/).
 
@@
   sample_sizes,
-  caption = 'Sample sizes for resistance training (RT) and non training control (CON) groups for dataset.'
-)
+  caption = 'Sample sizes for resistance training and non training control groups for dataset.'
+  ) %>%
+  pack_rows("RT", 1, 4) %>%
+  pack_rows("CON", 5, 8) %>%
+  footnote(general = c("RT = resistance training", "CON = non-training control")
+           ) %>%
+  row_spec(0, bold = TRUE) %>%
+  kable_classic(full_width = FALSE)
 
 ```
+
+Polito et al.
+[-@polito_moderators_2021] conducted a systematic review and meta-analysis of randomised trials including a RT intervention where a non-training control comparison group was included.
+Their analysis focused upon the SMD between the RT interventions and control groups from the studies included, with both overall effect estimate and moderator analyses (i.e., meta-regressions) were performed.
+Given that Polito et al.
+[-@polito_moderators_2021] included only studies with non-training control groups, their study selection offers a unqiue context to examine variation of interindividual responses specifically by means of comparing the variances in change scores between the RT intervention and control groups.
+Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies.
+Table \@ref(tab:summary-characteristics-tab) shows the study and participant characteristics.
+
 ```{r summary-characteristics-tab, message=FALSE, warning=FALSE, echo=FALSE}
@@
   summary_table,
-  caption = 'Summary of study and participant characteristics (median [IQR], number of groups [%]).'
-)
+  caption = 'Summary of study and participant characteristics.',
+  align = c("l","c") 
+  ) %>%
+  pack_rows("Training Status",6, 7, bold = FALSE) %>%
+  pack_rows("Sample Type", 8, 9, bold = FALSE) %>%
+  pack_rows("RT + Adjuvant Intervention?", 10, 11, bold = FALSE) %>%
+  pack_rows("Task Failure?", 18, 19, bold = FALSE) %>%
+  footnote(general = c("RT = resistance training; ", "Continuous variables are median (IQR); ", "Categorical variables are count (%); ", "Not all studies reported full descriptive data (see dataset; https://osf.io/kg2z4)")
+           ) %>%
+  row_spec(0, bold = TRUE) %>%
+  kable_classic(full_width = FALSE) %>%
+    kable_styling()
 
 ```
+
 ## Detecting the presence of interindividual response variation to resistance training intervention
@@
 Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
+Indeed, for a field such as sport and exercise science where samples sizes are typically small, meta-analysis becomes even more valuable as such small sample in primary studies have even lower statistical power to detect differences in variation as compared to means [@yang_low_2022].
 The examples we presented used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021], and so hopefully are more relatable for researchers in sport and exercise sciences.

### 2022-08-12 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -279,7 +279,7 @@ save(forest_plots, file = "plots/forest_plots")
 
 forest_plots
 
-ggsave("SMD_plots.tiff", width = 10, height = 10, device = "tiff", dpi = 300)
+ggsave("plots/SMD_plots.tiff", width = 10, height = 10, device = "tiff", dpi = 300)
 
 
 ###### Comparison of variance of change between RT and CON ######
@@ -529,7 +529,7 @@ save(mean_variance_pre_plots, file = "plots/mean_variance_pre_plots")
 
 mean_variance_pre_plots
 
-ggsave("mean_variance_pre_plots.tiff", width = 10, height = 7.5, device = "tiff", dpi = 300)
+ggsave("plots/mean_variance_pre_plots.tiff", width = 10, height = 7.5, device = "tiff", dpi = 300)
 
 
 ### Fitting a random slope mixed effects model for mean-variance
@@ -572,7 +572,7 @@ model_mean_variance_pre_plots <- ggplot(Data_long_pre, aes(x=log(mean), y=SD_log
   geom_line(aes(x=log(mean), y=pred, color = outcome)) +
   scale_fill_manual("Outcome", values = alpha(c("#009E73", "#D55E00"),0.5)) +
   scale_color_manual("Outcome", values = alpha(c("#009E73", "#D55E00"),0.5)) +
-  labs(x = "Log Mean Change Score", y = "Log Standard Deviation of the Change Score", color = "Outcome", shape = "", fill = "") +
+  labs(x = "Log Mean", y = "Log Standard Deviation", color = "Outcome", shape = "", fill = "") +
   theme_classic() +
   guides(size = "none", fill = "none")
 
@@ -580,7 +580,7 @@ save(model_mean_variance_pre_plots, file = "plots/model_mean_variance_pre_plots"
 
 model_mean_variance_pre_plots
 
-ggsave("model_mean_variance_pre_plots.tiff", width = 7.5, height = 5, device = "tiff", dpi = 300)
+ggsave("plots/model_mean_variance_pre_plots.tiff", width = 7.5, height = 5, device = "tiff", dpi = 300)
 
 ###### Section - Reanalysis of interindividual response variation using $lnCVR$ and the random slope mixed effects model
 
@@ -692,7 +692,7 @@ save(mean_variance_delta_plots, file = "plots/mean_variance_delta_plots")
 
 mean_variance_delta_plots
 
[…]

### 2022-08-12 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Draft to share with everyone

@@
 Given such growth, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
-```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976"}
+```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976."}
 
@@
 ## Detecting the presence of interindividual response variation to resistance training intervention
-First we conduct a traditional SMD based effect size meta-analysis to explore the effects of RT compared to control for strength outcomes and hypertrophy (i.e., muscle mass/size) outcomes.
+First we conduct a traditional SMD based effect size meta-analysis to explore the effects of RT compared to control for strength outcomes and hypertrophy (i.e., muscle mass/size) outcomes[^6].
 Polito et al.
-[-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple effects within groups within studies) and so a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group[^6].
+[-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple effects within groups within studies) and so a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group[^7].
 We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in in responses differed between RT interventions and controls.
@@
 
-```{r forest-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes"}
+```{r forest-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes."}
 
@@
 As might be expected, in comparison to non-training controls the RT interventions produced increases in strength (SMD = `r SMD_SDir_logVR$Estimate[1]` [95%CI: `r SMD_SDir_logVR$Lower[1]` to `r SMD_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[1]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[1]`) and hypertrophy outcomes (SMD = `r SMD_SDir_logVR$Estimate[4]` [95%CI: `r SMD_SDir_logVR$Lower[4]` to `r SMD_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[4]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[4]`).
-Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-plot)) were fairly wide and relative heterogeneity was fairly high coming mostly from between study variance[^7].
+Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-plot)) were fairly wide and relative heterogeneity was fairly high coming mostly from between study variance.
 
-In addition to the SMD results, both the $SD_{ir}$ and $\textrm{ln}VR$ were also positive for both strength ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[2]` [95%CI: `r SMD_SDir_logVR$Lower[2]` to `r SMD_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[2]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[2]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[3]` [95%CI: `r SMD_SDir_logVR$Lower[3]` to `r SMD_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[3]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[3]`) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[5]` [95%CI: `r SMD_SDir_logVR$Lower[5]` to `r SMD_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[5]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[5]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[6]` [95%CI: `r SMD_SDir_logVR$Lower[6]` to `r SMD_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[6]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[6]`) suggesting that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of of interindividual response variation.
-This might support the previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* apart from the random error).
+In addition to the SMD results, both the $SD_{ir}$ and $\textrm{ln}VR$ were also positive for both strength ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[2]` [95%CI: `r SMD_SDir_logVR$Lower[2]` to `r SMD_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[2]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[2]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[3]` [95%CI: `r SMD_SDir_logVR$Lower[3]` to `r SMD_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[3]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[3]`) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[5]` [95%CI: `r SMD_SDir_logVR$Lower[5]` to `r SMD_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[5]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[5]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[6]` [95%CI: `r SMD_SDir_logVR$Lower[6]` to `r SMD_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[6]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[6]`) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation.
+This might support previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* apart from the random error).
 However, as noted both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean.
@@
 
-```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes"}
+```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes."}
 
@@
 \begin{equation}
-\textrm{ln}\hat\sigma_{ij}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j})+\beta_{1}\textrm{ln}\overline{x}_{ij}+\epsilon_{ij}+m_{ij} 
+\textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
 (\#eq:lnm-lns-model-eq)
@@
 
-where $\textrm{ln}\hat\sigma_{ij}$ is the effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ij}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, and $\epsilon_{ij}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ij}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ij}}$.
+where $\textrm{ln}\hat\sigma_{ij}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, and $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
 Additional predictor terms could be added to this model; for example, we could model a categorical variable for the outcome type and include $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome groups), and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group.
@@
 
-```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores"}
+```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores."}
 
@@
 
-## Reanalysis of interindividual response variation using $lnCVR$ and the random slope mixed effects model
+## Reanalysis of interindividual response variation using $\textrm{ln}CVR$ and the random slope mixed effects model
 As can be seen in figure \@ref(fig:mean-variance-delta-plot) there is a mean-variance relationship in the change score data whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
@@
 
-```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes"}
+```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes."}
 
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue though.
-Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group and before using the $\textrm{ln}CVR$ as the response variable leads to different conclusions compared to absolute variance comparisons.
-The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[88]`).
+Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study and group as before using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons.
+The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[88]`) and further there is lower relative heterogeneity between studies in the effect estimates.
 
@@
 \begin{equation}
-\textrm{ln}CVR_{ij}=\mu+\tau_{i}+\tau_{j}+m_{ij} 
+\textrm{ln}CVR_{ijk}=\mu+\tau_{i}+\tau_{j}+m_{ij} 
 (\#eq:lnCVR-model-eq)
@@
 
-where $\textrm{ln}CVR_{ij}$ is each effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ij}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ij}}$.
+where $\textrm{ln}CVR_{ij}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ijk}}$.
 Instead, we can use a version of the random slope model described above (equation \@ref(eq:lnm-lns-model-eq) and following paragraph) to compare the variability between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
@@
 Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)) we fit this model to the dataset.
-The results were however largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
+The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
 See figure \@ref(fig:model-mean-variance-delta-plot).
 
-```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores"}
+```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores."}
 
@@
 
-The reason for this apparent reduction in variation after introduction of an RT intervention is not necessarily discernible from this analysis.
-Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; [@halliday_resistance_2017]), or a ceiling effect on change (i.e., plateau in response; [@steele_long-term_2022]) has a constraining effect [@cortes_martinez_constant_2021].
 Hopefully it is clear from the regression model here where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework.
@@
 For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to recommendations [@steele_higher_2017; @steele_long-term_2022].
+
+The reason for this apparent reduction in variation after introduction of an RT intervention is not necessarily discernible from this analysis.
+Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @halliday_resistance_2017), or a ceiling effect on change (i.e., plateau in response; @steele_long-term_2022) has a constraining effect [@cortes_martinez_constant_2021].
+This potentially represents another interesting area of future study regarding variation however; how to produce interventions that actually reduce variance in an outcome.
+In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
 
@@
 
-[^6]: Data was coded such that study and group had explicit nesting.
+[^6]: Note, we also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases.
+There did not appear to be any obvious small study bias in the dataset (see https://osf.io/stqr3).
 
-[^7]: Note, we also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases.
-There did not appear to be any obvious small study bias in the dataset (ADD LINK TO SUPPLEMENTARY)
+[^7]: Data was coded such that study and group had explicit nesting.
 
-[^8]: See supplementary materials (ADD LINK TO SUPPLEMENTARY) for model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continous predictors for both strength and hypertrophy outcomes.
+[^8]: See supplementary materials (https://osf.io/e6vpr) for model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continous predictors for both strength and hypertrophy outcomes.
 There were no obvious moderators of $\textrm{ln}CVR$ in particular.

### 2022-08-13 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Update search to be title only in europepmc

@@
 Since that time use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@hagger_meta-analysis_2022].
-One area of sport and exercise science which has comprised a considerable proportion of that growth appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)).
+One area of sport and exercise science which has comprised a reasonable proportion of that growth in recent years appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)).
 Given such growth, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.

### 2022-08-13 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -19,17 +19,17 @@ library(europepmc)
 
 ##### Trend of Meta-Analysis in Resistance Training
 
-SES_meta_trend <- epmc_hits_trend(query = "(sport OR exercise) AND meta analysis",
+SES_meta_trend <- epmc_hits_trend(query = 'TITLE: (sport OR exercise) AND "meta analysis"',
                                   period = 1976:2022) %>%
   mutate(`Search string` = "(sport OR exercise) AND meta analysis")
 
-RT_meta_trend <- epmc_hits_trend(query = "(resistance training OR strength training) AND meta analysis",
+RT_meta_trend <- epmc_hits_trend(query = 'TITLE: ("resistance training" OR "strength training") AND "meta analysis"',
                                  period = 1976:2022) %>%
   mutate(`Search string` = "(resistance training OR strength training) AND meta analysis")
 
 trends <- rbind(SES_meta_trend, RT_meta_trend) %>%
-  mutate(`Search string` = factor(`Search string`, levels = c("(sport OR exercise) AND meta analysis",
-                                                              "(resistance training OR strength training) AND meta analysis")))
+  mutate(`Search string` = factor(`Search string`, levels = c('(sport OR exercise) AND meta analysis',
+                                                              '(resistance training OR strength training) AND meta analysis')))
 
 trends_plot <- trends %>%
   ggplot(aes(x = year, y = (query_hits / all_hits)*100, linetype = `Search string`)) +

### 2022-09-22 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Update code and manuscript

--- 2022-08-12 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
@@
 author: 
-  - James Steele
-  - James Fisher
-  - Dave Smith
-  - Andrew Vigotsky
-  - Brad Schoenfeld
-  - Yefeng Yang
-  - Shinichi Nakagawa
-date: "2022-08-04"
-abstract: |
-  TO COMPLETE.
+  - James Steele$^a$
+  - James Fisher$^a$
+  - Dave Smith$^b$
+  - Brad Schoenfeld$^c$
+  - Yefeng Yang$^d$
+  - Shinichi Nakagawa$^d$
+date: "22nd September 2022"
+abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
+However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance.
+It is well known that there is considerable variability in response to interventions.
+Hence, interest has increased in exploring *precision* or *personalised* exercise approaches.
+Not only is the mean often affected by interventions, but variances may also be impacted.
+Exploration of variances in studies such as randomised controlled trials (RCTs) can yield insight into the existence of interindividual heterogeneity in response to interventions and help determine the degree of generalisability of effects.
+Yet, larger samples sizes than those used for typical mean effects are required when probing variances.
+Thus, for a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
+Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet its lack of application in research synthesis by way of meta-analysis, we introduce and discuss effect size approaches and models for meta-analysis of variation using relatable examples from resistance training RCTs."
+thanks: |
+  Preprint, submitted to Journal of Sport Sciences. 
+  Address for correspondence: james.steele@solent.ac.uk
+  Affiliations:
+  $^a$Faculty of Sport, Health, and Social Sciences, Solent University, UK;
+  $^b$Research Centre for Musculoskeletal and Sports Medicine, Manchester Metropolitan University, Manchester, UK;
+  $^c$Health Sciences Department, CUNY Lehman College, Bronx, New York, USA;
+  $^d$Evolution & Ecology Research Centre and School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, New South Wales, Australia
 output: bookdown::pdf_document2
 bibliography: references.bib
+csl: apa.csl
 toc: false
+header-includes:
+   - \usepackage{caption}
+   - \captionsetup[figure]{font=small}
 ---
@@
 # load the raw data to the environment
-Data <- read.csv(here::here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))
+Data <- read.csv(here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))
 
 # load the summary dataframes to the environment
-load(here("models/SMD_SDir_logVR"))
+load(here("models/SMD_RR_SDir_logVR"))
 load(here("models/mods_SMD_logCVR"))
@@
 # Introduction
-Though the quantitative synthesis of results across studies has existed since the 17th century [@plackett_studies_1958], the modern day term "meta-analysis" as it is now referred to was coined by Gene Glass in [-@glass_primary_1976].
-Since that time use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@hagger_meta-analysis_2022].
-One area of sport and exercise science which has comprised a considerable proportion of that growth appears to have been in the study of resistance training (RT; figure \@ref(fig:trend-plot)).
-Given such growth, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
+Although the quantitative synthesis of results across studies has existed since the 17th century [@plackett_studies_1958], the modern-day term "meta-analysis" was coined by Gene Glass in [-@glass_primary_1976].
+Since that time, the use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@hagger_meta-analysis_2022], with resistance training (RT) accounting for a considerable proportion of this growth (figure \@ref(fig:trend-plot)).
+Accordingly, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
 ```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976."}
@@
 ```
-As with many other fields [@nakagawa_meta-analysis_2015; @usui_meta-analysis_2021; @mills_detecting_2021] likely the most familiar aim with the use of meta-analysis, and indeed primary empirical research too, in sport and exercise science is to make comparisons between the means of measurements taken across different categorical grouping variables; for example, the comparison of an intervention group and a control group, the comparison of one intervention group to another, or comparison between non-manipulated categories such as biological sex.
-Indeed, a recent umbrella review [@bernardez-vazquez_resistance_2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes all of which focused on the comparison of mean changes between different intervention groups.
+As with many other fields [@nakagawa_meta-analysis_2015; @usui_meta-analysis_2021; @mills_detecting_2021] likely the most familiar aim with the use of meta-analysis, and indeed primary empirical research too, in sport and exercise science is to make comparisons between the means of measurements taken across different categorical grouping variables; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
+Indeed, a recent umbrella review [@bernardez-vazquez_resistance_2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes, all of which focused on the comparison of mean changes between different intervention groups.
 
-Most commonly a magnitude based[^1] effect size statistic [@caldwell_case_2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
-This is usually Cohen's *d* [@cohen_statistical_1988], or it's bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007] [^2].
+Most commonly, a magnitude based[^1] effect size statistic [@caldwell_case_2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
+This statistic is usually Cohen's *d* [@cohen_statistical_1988], or it's bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007] [^2].
 The SMD, and its sampling variance, $s^2_{SMD}$ are given by:
@@
 \begin{equation}
-s^2_{\textrm{ln}RR}=\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}+\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}}
+s^2_{\textrm{ln}RR}=\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}+\frac{s^4_{C}}{2n_{C}\overline{x}^2_{C}\overline{x}^4_{C}}+\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}}+\frac{s^4_{E}}{2n_{E}\overline{x}^2_{E}\overline{x}^4_{E}}
 (\#eq:lnRRvar-eq)
@@
 
-Due to its calculation the SMD is affected not only by the difference in means of the two groups, but is also affected by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$.
-Contrastingly, the $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), which only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
-Despite this, use of $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@deb_quantifying_2018] and to our knowledge only one meta-analysis in RT has used this effect size (Swinton et al., in press).
+Due to its calculation the SMD is affected not only by the difference in means of the two groups, but also by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$.
+In contrast, the $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), which only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
+Despite this, the use of effects sizes like the $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@deb_quantifying_2018; @nuzzo_how_2022] and to our knowledge only one meta-analysis in RT has used this kind of effect size [@swinton_interpreting_2022].
 
-Though the focus in sport and exercise science among other fields has been in estimating the average effects of interventions for both primary research and synthesis through meta-analysis, the field has been aware for some time that responses to certain interventions do vary potentially on a subgroup or even individual basis.
-The increased interested in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@hecksteden_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @ross_precision_2019; @swinton_statistical_2018; @hopkins_individual_2015; @kelley_precision_2022; @hrubeniuk_directions_2022; @pickering_non-responders_2019].
-However, despite the availability of approaches to compare variances between groups, in sports and exercise science this is rarely explored in primary research [@bonafiglia_interindividual_2022] and, though there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
+Although researchers in sport and exercise science among other fields have focused on estimating the average effects of interventions using randomised trial designs for both primary research and synthesis through meta-analysis, evidence suggests that responses to certain interventions may vary on a subgroup or even individual basis.
+The increased interest in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@hecksteden_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @ross_precision_2019; @swinton_statistical_2018; @hopkins_individual_2015; @kelley_precision_2022; @hrubeniuk_directions_2022; @pickering_non-responders_2019].
+However, despite the availability of approaches to compare variances between groups, in sport and exercise science this is rarely explored in primary research [@bonafiglia_interindividual_2022] and, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
 Examination of interindividual heterogeneity in response to interventions presents considerable value to researchers and practitioners in sport and exercise science; interventions with low interindividual variation are likely to be widely generalisable, whilst an intervention with high interindividual variation is likely to have effects that are either subgroup or individual specific.
-The former kind of intervention might be widely applicable across individuals, whilst the latter kind of intervention requires specific research to tease apart subgroup- or participant-by-intervention interactions to enable successful practical application.
+The former kind of intervention might be widely applicable across individuals, whilst the latter kind of intervention requires specific research, typically with large samples [@hecksteden_individual_2015], to tease out subgroup- or participant-by-intervention interactions to facilitate successful practical application.
 
 Comparison of heterogeneity in responses such as post-scores or change scores to interventions are however not the only possible use of statistical methods for comparing variances.
-For example, in other fields such as ecology there have been calls to shift focus onto the exploration of dispersion of traits between groups in non-experimental or interventional designs [@nakagawa_mean_2012].
-Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between participant acute response variation for the purposes of identifying methods[^3] to reduce RT stimulus heterogeneity [@exner_does_2022] as well as meta-analysis exploring between participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperin_accuracy_2022].
+For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawa_mean_2012].
+Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between participant acute response variation for the purposes of identifying methods[^3] to reduce RT stimulus heterogeneity [@exner_does_2022] as well as meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperin_accuracy_2022].
 
 Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet the lack of application in research synthesis by way of meta-analysis, we present and discuss effect size approaches and models for meta-analysis of variation.
-We provide examples of the approaches presented using data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021].
+Indeed, meta-analysis presents an incredibly valuable method for exploring variation in a field such as sport and exercise science due to the typically small samples in primary studies.
+Such small samples have even lower statistical power to detect differences in variation as compared to means [@yang_low_2022].
 
@@
 Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework.
-However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; @hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015).
+However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; -@hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015).
 
 ## Standard deviation for individual responses ($SD_{ir}$)
-In the context of *precision* or *personalised* approaches to exercise prescription comparison of variation between two groups, control (C) and intervention (E), the $SD_{ir}$ has been proposed as an approach to determining the extent to which individual responses manifest [@hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019].
-The standard deviation of changes scores (post-intervention scores minus pre-intervention scores) within the intervention group reflects the gross combination of a number of sources of variation including participant-by-intervention interactions (i.e., actual individual responsiveness or 'trainability'), within-participant variability in intervention response (i.e., variability in response to the same intervention administered to the same participant), and random error (i.e., from pre and post measurements; @hecksteden_individual_2015).
-The standard deviation of change scores from the control group (assuming it is a non-intervention control group and not something like a 'usual-care' group) by contrast only reflects random error [@hecksteden_individual_2015].
-As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention that might reflect individual responses.
+In the context of *precision* or *personalised* approaches to exercise prescription the $SD_{ir}$ has been proposed as an approach to determine the extent to which individual responses manifest by comparison of variation between two groups; control and intervention [@hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019].
+The standard deviation of change scores (post-intervention scores minus pre-intervention scores) within the intervention group reflects the gross combination of a number of sources of variation including: participant-by-intervention interactions (i.e., actual individual responsiveness or 'trainability'), within-participant variability in intervention response (i.e., variability in response to the same intervention administered to the same participant), and random error (i.e., from pre and post measurements; @hecksteden_individual_2015).
+The standard deviation of change scores from the control group (assuming it is a non-intervention control group and not something like a 'usual-care' group) by contrast is assumed to only reflect random error[^4] [@hecksteden_individual_2015].
+As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention and which might reflect individual responses.
+Whilst the $SD_{ir}$ has been proposed and used primarily in the context of individual response variation to interventions, it should be noted that this kind of absolute comparison of variance between groups or conditions is not limited to such applications.
+
 The $SD_{ir}$, and its sampling variance, $s^2_{SD_{ir}}$ are given by:
@@
 \begin{equation}
-s^2_{SD_{ir}}=2(\frac{s^4_{E}}{n_{E}-1}+\frac{s^4_{C}}{n_{C}-1})
+s^2_{SD_{ir}}=2\biggl(\frac{s^4_{E}}{n_{E}-1}+\frac{s^4_{C}}{n_{C}-1}\biggl)
 (\#eq:sdIRvar-eq)
@@
 
-Thus, the $SD_{ir}$ reflects a comparison of the absolute variance in change scores between C and E.
-Whilst the $SD_{ir}$ has been proposed and used primarily in the context of individual response variation to interventions, it should be noted that this kind of absolute comparison of variance between groups or conditions is not limited to such applications.
+Thus, the $SD_{ir}$ reflects a comparison of the absolute variance in change scores between control and intervention groups.
+However, a potential concern with the $SD_{ir}$ is it's potential to violate assumptions of normality which is not the case for other effect size statistics such as $\textrm{ln}VR$ and $\textrm{ln}CVR$.
 
 ## Log ratio of standard deviations ($\textrm{ln}VR$)
-A similar effect size statistic for the comparison of absolute variance between groups, and one which has had wide applications in more than just intervention response variability within fields like ecology and evolution, is the $\textrm{ln}VR$ [@nakagawa_meta-analysis_2015].
+A similar effect size statistic for the comparison of absolute variance between groups, and one that has had wide applications in more than just intervention response variability within fields like ecology and evolution, is the $\textrm{ln}VR$ [@nakagawa_meta-analysis_2015; @senior_revisiting_2020].
 An unbiased estimator of the natural logarithm of a population standard deviation ($\textrm{ln}\sigma$), and its sampling variance, $s^2_{\textrm{ln}\sigma}$ is given by:
@@
 where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$.
-Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups, such as a control (C) and intervention (E), the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
+Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
 
 \begin{equation}
-\textrm{ln}VR=\textrm{ln}(\frac{s_{E}}{s_{C}})+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
+\textrm{ln}VR=\textrm{ln}\biggl(\frac{s_{E}}{s_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
 (\#eq:lnVR-eq)
@@
 \begin{equation}
-s^2_{\textrm{ln}VR}=\frac{1}{2(n_{E}-1)}+\frac{1}{2(n_{C}-1)}
+s^2_{\textrm{ln}VR}=\frac{1}{2}\biggl(\frac{n_{C}}{(n_{C}-1)^2}+\frac{n_{E}}{(n_{E}-1)^2}\biggl)
 (\#eq:lnVRvar-eq)
@@
 However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparison group (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$).
-In this case it is likely that the standard deviation will be larger in the group with the larger mean (e.g., $s_{E}$ is larger than $s_{C}$).
-This mean-variance relationship is common for many variables and datasets[^4] and we will provide examples of this below.
-They also assume constant measurement error over the range of values for the mean which can impact their utility for examining response variation [@tenan_comment_2020].
+In this case, it is likely that the standard deviation will be larger in the group with the larger mean (e.g., $s_{E}$ is larger than $s_{C}$).
+This mean-variance relationship is common for many variables and datasets[^5] and we provide examples of this below.
+They also assume constant measurement error over the range of values for the mean, which can impact their utility for examining response variation [@tenan_comment_2020].
 
@@
 \begin{equation}
-\textrm{ln}CVR=\textrm{ln}(\frac{CV_{E}}{CV_{C}})+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
+\textrm{ln}CVR=\textrm{ln}\biggl(\frac{CV_{E}}{CV_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
 (\#eq:lnCVR-eq)
@@
 where $CV_{E}$ and $CV_{C}$ are ${s_{E}}/{\overline{x}_{E}}$ and ${s_{C}}/{\overline{x}_{C}}$ respectively.
-Nakagawa et al.
-[-@nakagawa_meta-analysis_2015] derived the sampling variance, $s^2_{\textrm{ln}CVR}$, as:
+Senior et al.
+[-@senior_revisiting_2020] derived the sampling variance, $s^2_{\textrm{ln}CVR}$, as:
 
@@
 \begin{split}
-s^2_{\textrm{ln}CVR}=\frac{s^2_{C}}{n_{C}\overline{x}_{C}}+\frac{1}{2(n_{C}-1)}-2\rho_{\textrm{ln}\overline{x}_{C},\textrm{ln}{s}_{C}}\sqrt{\frac{s^2_{C}}{n_{C}\overline{x}_{C}}+\frac{1}{2(n_{C}-1)}} \\
-+ \frac{s^2_{E}}{n_{E}\overline{x}_{E}}+\frac{1}{2(n_{E}-1)}-2\rho_{\textrm{ln}\overline{x}_{E},\textrm{ln}{s}_{E}}\sqrt{\frac{s^2_{E}}{n_{E}\overline{x}_{E}}+\frac{1}{2(n_{E}-1)}}
+s^2_{\textrm{ln}CVR}=\frac{s_{C}^2}{n_{C}\overline{x}_{C}^2}+\frac{s_{C}^4}{2n_{C}^2\overline{x}_{C}^4}+\frac{n_{C}}{(n_{C}-1)^2} \\
++\frac{s_{E}^2}{n_{E}\overline{x}_{E}^2}+\frac{s_{E}^4}{2n_{E}^2\overline{x}_{E}^4}+\frac{n_{E}}{(n_{E}-1)^2}
 \end{split}
@@
 
-where $\rho_{\textrm{ln}\overline{x}_{C},\textrm{ln}{s}_{C}}$ and $\rho_{\textrm{ln}\overline{x}_{E},\textrm{ln}{s}_{E}}$ are the correlatios between the means and the standard deviation in the C and E groups respectively on the log scale across studies.
-
 # Examples using resistance training studies
-As noted, the examples presented used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021].
-Here we have used their list of included studies and re-extracted data from 111 of these[^5].
-All analysis examples were performed in using R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauer_conducting_2010].
-The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are all available on the Open Science Framework (https://osf.io/2h9ma/).
+As noted, to facilitate understanding for those new to examination of variation, we provide primary examples of the approaches presented using data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021].
+Here we have used their list of included studies and re-extracted data from 111 of these[^6].
+All analysis examples were performed in R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauer_conducting_2010].
+The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/).
 
@@
 Polito et al.
-[-@polito_moderators_2021] conducted a systematic review and meta-analysis of randomised trials including a RT intervention where a non-training control comparison group was included.
-Their analysis focused upon the SMD between the RT interventions and control groups from the studies included, with both overall effect estimate and moderator analyses (i.e., meta-regressions) were performed.
+[-@polito_moderators_2021] conducted a systematic review and meta-analysis of randomised trials that included a RT intervention group(s) and a non-training control comparison group.
+Their analysis focused upon the SMD between the RT intervention group(s) and the control group from the studies included, with both overall effect estimate and moderator analyses (i.e., meta-regressions) performed.
 Given that Polito et al.
-[-@polito_moderators_2021] included only studies with non-training control groups, their study selection offers a unqiue context to examine variation of interindividual responses specifically by means of comparing the variances in change scores between the RT intervention and control groups.
+[-@polito_moderators_2021] included only studies with a non-training control group, their study selection offers a unique context to examine variation of interindividual responses specifically by means of comparing the variances in change scores between the RT intervention groups(s) and control group.
 Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies.
+Indeed, this highlights the typically small samples used in sport and exercise science, and thus low power to detect difference in both means and variance [@yang_low_2022], emphasising the value of meta-analysis to explore variation.
 Table \@ref(tab:summary-characteristics-tab) shows the study and participant characteristics.
@@
 ## Detecting the presence of interindividual response variation to resistance training intervention
-First we conduct a traditional SMD based effect size meta-analysis to explore the effects of RT compared to control for strength outcomes and hypertrophy (i.e., muscle mass/size) outcomes[^6].
+First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^7] meta-analyses to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^8].
 Polito et al.
-[-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple effects within groups within studies) and so a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and group[^7].
-We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in in responses differed between RT interventions and controls.
-A positive SMD would indicate that RT interventions increase outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses compared to controls (i.e., suggests the presence of interindividual response variation).
+[-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple outcomes measures within each arm, i.e., intervention group(s) and control group, within each study) and thus a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and arm[^9].
+We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
+A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produce greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).
 
-```{r forest-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes."}
+```{r forest-SMD-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes."}
 
 # "Caterpillar plots of SMD effect sizes for strength and hypertrophy outcomes.
-Note, each estimate and 95% confidence interval is a single observed effect size within each group within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
 
-load(here("plots/forest_strength"))
-load(here("plots/forest_hypertrophy"))
+load(here("plots/forest_SMD_strength"))
+load(here("plots/forest_SMD_hypertrophy"))
 library(patchwork)
-(forest_strength / forest_hypertrophy) +
+(forest_SMD_strength / forest_SMD_hypertrophy) +
   plot_annotation(tag_levels = "A")
@@
 
-The pattern of results from our models examining SMDs (figure \@ref(fig:forest-plot)) were similar to those reported by Polito et al.
-[-@polito_moderators_2021] albeit with slightly lower estimates for both outcome types; this possibly being due to our use of a multilevel mixed-effects meta-analysis model which allowed for each effect size included to be more appropriately weighted.
-As might be expected, in comparison to non-training controls the RT interventions produced increases in strength (SMD = `r SMD_SDir_logVR$Estimate[1]` [95%CI: `r SMD_SDir_logVR$Lower[1]` to `r SMD_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[1]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[1]`) and hypertrophy outcomes (SMD = `r SMD_SDir_logVR$Estimate[4]` [95%CI: `r SMD_SDir_logVR$Lower[4]` to `r SMD_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[4]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[4]`).
-Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-plot)) were fairly wide and relative heterogeneity was fairly high coming mostly from between study variance.
+The pattern of results from our models examining SMDs (figure \@ref(fig:forest-SMD-plot)) were similar to those reported by Polito et al.
+[-@polito_moderators_2021] albeit with slightly lower estimates for both outcome types; possibly due to our use of a multilevel mixed-effects meta-analysis model that allowed for each individual effect size to be more appropriately weighted.
+As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[4]`%).
+Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between study variance.
 
-In addition to the SMD results, both the $SD_{ir}$ and $\textrm{ln}VR$ were also positive for both strength ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[2]` [95%CI: `r SMD_SDir_logVR$Lower[2]` to `r SMD_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[2]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[2]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[3]` [95%CI: `r SMD_SDir_logVR$Lower[3]` to `r SMD_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[3]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[3]`) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_SDir_logVR$Estimate[5]` [95%CI: `r SMD_SDir_logVR$Lower[5]` to `r SMD_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[5]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[5]`; $\textrm{ln}VR$ = `r SMD_SDir_logVR$Estimate[6]` [95%CI: `r SMD_SDir_logVR$Lower[6]` to `r SMD_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_SDir_logVR$I2.study[6]`, $I^2_{group}$ = `r SMD_SDir_logVR$I2.group[6]`) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation.
-This might support previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* apart from the random error).
-However, as noted both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean.
+```{r forest-RR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of exponentiated RR effect sizes for strength (A) and hypertrophy (B) outcomes."}
+
+# "Caterpillar plots of exponentiated RR effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+
+load(here("plots/forest_RR_strength"))
+load(here("plots/forest_RR_hypertrophy"))
+library(patchwork)
+(forest_RR_strength / forest_RR_hypertrophy) +
+  plot_annotation(tag_levels = "A")
+
+```
+
+For the $\textrm{ln}RR$ results we exponentiated them and converted to percentages to be more interpretable.
+These were similar with greater proportional increases in strength compared with hypertrophy (figure \@ref(fig:forest-RR-plot)).
+Increases were seen for both strength ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[2]` [95%CI: `r SMD_RR_SDir_logVR$Lower[2]` to `r SMD_RR_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[2]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[2]`%) and hypertrophy outcomes ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[5]`%).
+Confidence intervals were precise for both outcomes again, though notably there was essentially no between study or arm variance in estimates.
+
+In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[8]`%) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation.
+
+This additional variance might support previous perspectives [@carpinelli_interindividual_2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* independent of the random error).
+However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean.
 As we have seen from the SMD analysis RT interventions increase mean scores.
-Thus, if there is a mean-variance relationship in the data an increase in the mean alone may be fully responsible for any apparent increase in variation.
+Thus, if there is a mean-variance relationship in the data, an increase in the mean alone may be fully responsible for any apparent increase in variation.
 As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
@@
 First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a model that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$.
+
+```{r forest-SDir-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SDir effect sizes for strength (A) and hypertrophy (B) outcomes."}
+
+# "Caterpillar plots of SDir effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+
+load(here("plots/forest_SDir_strength"))
+load(here("plots/forest_SDir_hypertrophy"))
+library(patchwork)
+(forest_SDir_strength / forest_SDir_hypertrophy) +
+  plot_annotation(tag_levels = "A")
+
+```
+```{r forest-logVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$VR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}
+
+# "Caterpillar plots of $\textrm{ln}RR$ effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+
+load(here("plots/forest_logVR_strength"))
+load(here("plots/forest_logVR_hypertrophy"))
+library(patchwork)
+(forest_logVR_strength / forest_logVR_hypertrophy) +
+  plot_annotation(tag_levels = "A")
+
+```
 
@@
 In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al.
-[-@polito_moderators_2021] using a multilevel mixed-effects model.
+[-@polito_moderators_2021].
 
@@
 
-As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) & (B), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
+As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
 This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylor_aggregation_1961].
@@
 where $a$ and $b$ are some constants.
-When this relationship (equation \@ref(eq:taylors-eq)) holds, under most circumstances the standard deviation is not proportional to the mean.
+When this relationship holds, under most circumstances the standard deviation is not proportional to the mean.
 However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear:
@@
 
-Figure \@ref(fig:mean-variance-pre-plot)(C) & (D) shows that the relationship between the mean and variance on the log scale better meets the assumption of normality.
-Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from outcomes over multiple groups and studies we can also estimate this relationship using the following model:
+Figure \@ref(fig:mean-variance-pre-plot)(B) and (D) shows that the relationship between the mean and variance on the log scale better meets the assumption of normality.
+Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from multiple outcomes within multiple arms within studies we can also estimate this relationship using a multilevel mixed-effects meta-regression model.
+For example, the following model specifies $\textrm{ln}\overline{x}$ as a fixed effect with random intercepts for study and arm:
 
@@
 
-where $\textrm{ln}\hat\sigma_{ij}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, and $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
-Additional predictor terms could be added to this model; for example, we could model a categorical variable for the outcome type and include $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome groups), and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group.
-In this case $\tau_{i}$ and $\varphi_{i}$, and $\tau_{j}$ and $\varphi_{j}$ are assumed to have multivariate normal distributions with the following variance-covariance structure:
+where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^10]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, and $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
 
-\begin{equation}
-\binom{\tau}{\varphi}=(\binom{0}{0},
-\left(\begin{array}{cc} 
-\sigma^2_{\tau} & \rho\sigma_{\tau}\sigma_{\varphi}\\
-\rho\sigma_{\tau}\sigma_{\varphi} & \sigma^2_{\varphi}
-\end{array}\right))
-(\#eq:lnCVR-model-varcov-eq)
-\end{equation}
+Additional predictor terms could be added to this model; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^11].
 
@@
 
-## Reanalysis of interindividual response variation using $\textrm{ln}CVR$ and the random slope mixed effects model
-As can be seen in figure \@ref(fig:mean-variance-delta-plot) there is a mean-variance relationship in the change score data whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
+## Reanalysis of interindividual response variation using $\textrm{ln}CVR$ and the meta-regression of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$
+As can be seen in figure \@ref(fig:mean-variance-delta-plot)(A) and (C) there is a mean-variance relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
+Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figure \@ref(fig:mean-variance-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seens with the pre-intervention scores.
 As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue though.
-Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study and group as before using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons.
-The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{group}$ = `r mods_SMD_logCVR$I2.group[88]`) and further there is lower relative heterogeneity between studies in the effect estimates.
+Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study and arm as before using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons.
+The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I2.arm[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{arm}$ = `r mods_SMD_logCVR$I2.arm[88]`) and further there is lower relative heterogeneity between studies in the effect estimates (figure \@ref(fig:forest-logCVR-plot).
+
+```{r forest-logCVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$CVR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}
+
+# "Caterpillar plots of $\textrm{ln}CVR$ effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+
+load(here("plots/forest_logCVR_strength"))
+load(here("plots/forest_logCVR_hypertrophy"))
+library(patchwork)
+(forest_logCVR_strength / forest_logCVR_hypertrophy) +
+  plot_annotation(tag_levels = "A")
+
+```
 
 There is however a potential limitation for the $\textrm{ln}CVR$ also that may need to be considered.
-Firstly, it is limited to the use of ratio scale data (which is not the case for the $\textrm{ln}\hat\sigma$ or $\textrm{ln}VR$).
-But secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
-Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative approach that is equivalent may be more appropriate.
+Firstly, it is limited to the use of ratio scale data (which is not the case for the $\textrm{ln}\hat\sigma$ or $\textrm{ln}VR$); hence the need to transform the change scores to be positively signed.
+Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
+Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)), an alternative approach that is equivalent may be more appropriate.
 
@@
 
-where $\textrm{ln}CVR_{ij}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ group ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of groups) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ group, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ijk}}$.
-Instead, we can use a version of the random slope model described above (equation \@ref(eq:lnm-lns-model-eq) and following paragraph) to compare the variability between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
-In this case, the categorical variable for the outcome type is instead swapped for the group and the new model term included becomes $(\beta_{2} + \varphi_{i} + \varphi_{j})Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ group.
+where $\textrm{ln}CVR_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ijk}}$.
+
+Instead, we can use a version of the model described above (equation \@ref(eq:lnm-lns-model-eq) and the paragraph which follows it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
+In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
 
@@
 # load the summary dataframes to the environment
-load(here("models/RobuEstMultiLevelModel_log_mean_mod_strength"))
-load(here("models/RobuEstMultiLevelModel_log_mean_mod_hypertrophy"))
+load(here("models/MultiLevelModel_ri__log_mean_mod_strength"))
+load(here("models/MultiLevelModel_ri__log_mean_mod_hypertrophy"))
 
@@
 
-Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)) we fit this model to the dataset.
-The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
+Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model to the dataset[^12].
+The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_strength$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.ub[3],2)`]).
 See figure \@ref(fig:model-mean-variance-delta-plot).
@@
 Hopefully it is clear from the regression model here where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework.
-Models can be fit to explore not only how study and group level characteristics moderate effect size estimates when considering not only SMDs, but also when considering the variance based effect size statistics and models employed in this article[^8].
+Models can be fit to explore not only how study and arm level characteristics moderate effect size estimates when considering not only effects sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance based effect size statistics and models employed in this article[^13].
 
 # Discussion
-Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
-Indeed, for a field such as sport and exercise science where samples sizes are typically small, meta-analysis becomes even more valuable as such small sample in primary studies have even lower statistical power to detect differences in variation as compared to means [@yang_low_2022].
-The examples we presented used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021], and so hopefully are more relatable for researchers in sport and exercise sciences.
+Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
+Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^14] [@yang_low_2022].
+The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@polito_moderators_2021], which hopefully makes the findings more relatable for researchers in sport and exercise sciences.
 
-Given the considerable interest in *precision* or *personalised* exercise prescription in particular, it is of particular interest to note the different conclusions drawn dependent upon the approach taken to determining from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
+It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
 Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely increased variance above random error, suggesting the presence of inter-individual response variation.
-In the case of RT interventions there is evidently an average intervention effect which is positive and as such this result might merely imply that while all likely benefit, some benefit much moreso than others.
-Exploring variation however even in the absence of average intervention effects might still be important as those with large enough variance could imply that the intervention is at least beneficial to some [@usui_meta-analysis_2021].
-Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of RT interventions to maximise strength and hypertrophy outcomes from our example.
+In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive, yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
+Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usui_meta-analysis_2021].
+Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.
 
-However, much like that seen in cross-sectional pre-intervention scores here and most physical and biological systems, change scores demonstrated a mean-variance relationship in addition to heteroskedasticity.
-The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ in addition to the random slope model of $\textrm{ln}\hat\sigma$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance implying that there is likely little to no inter-individual response variation to explain.
-Interventions, such as RT interventions explored here, which induce both meaningful SMDs and also show little evidence suggestive of interindividual variation are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usui_meta-analysis_2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortes_martinez_constant_2021].
+However, similar to the cross-sectional pre-intervention scores reported here (and indeed most physical and biological variables), change scores demonstrated a mean-variance relationship in addition to heteroskedasticity.
+The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ in addition to the random slope model of $\textrm{ln}\hat\sigma$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance implying that there is likely little to no interindividual response variation to explain.
+Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usui_meta-analysis_2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortes_martinez_constant_2021].
 Interventions such as these are valuable for the simplification of guidelines and recommendations.
@@
 
-The reason for this apparent reduction in variation after introduction of an RT intervention is not necessarily discernible from this analysis.
+The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis.
 Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @halliday_resistance_2017), or a ceiling effect on change (i.e., plateau in response; @steele_long-term_2022) has a constraining effect [@cortes_martinez_constant_2021].
-This potentially represents another interesting area of future study regarding variation however; how to produce interventions that actually reduce variance in an outcome.
+However, this potentially represents another interesting area of future study regarding variation; specifically, how to produce interventions that actually reduce variance in an outcome.
 In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
@@
 Embracing variability and focusing on more than merely the mean differences between groups or conditions, such as intervention and control comparisons, has the potential to inform experimental design and lead to changes in both the approach and direction of follow-up studies.
-Where considerable differences in variation are present, this suggests a meaningful line of research should be aimed at identifying subgroup- or participant-by-intervention interactions.
-Where variance differences are limited this suggests that translational work towards generalisable implementation might be most meaningful.
-Future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utlising the established approaches [@nakagawa_meta-analysis_2015; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @usui_meta-analysis_2021; @mills_detecting_2021] presented here and echoing the efforts of other recent work [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
+Whether there is evidence of meaningful average intervention effects or not, where considerable variance effects are present it suggests that a meaningful line of research would be to aim at identifying subgroup- or participant-by-intervention interactions using appropriate study designs [@hecksteden_individual_2015].
+Where variance effects are limited this instead suggests that translational work towards generalisable implementation might be most the most meaningful line of future research.
+Finally, there may be cases where it is in fact desirable to identify interventions that actually reduce variance; for example, improvements in methodological approaches to enhance research [@usui_meta-analysis_2021], or interventions to reduce variation in sport performances.
+Thus, researchers in sport and exercise science should consider exploring variance more systematically, and indeed utilise the meta-analytic framework to support this.
+This could include the re-analysis of past meta-analysis as we have done so here, and indeed researchers conducted future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawa_meta-analysis_2015; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @usui_meta-analysis_2021; @mills_detecting_2021] presented here and echoing the efforts of other recent work [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
 
@@
 
-[^2]: We will refer to both merely as the SMD throughout the manuscript for simplicity.
+[^2]: We will refer to both merely as the SMD throughout the manuscript for simplicity and note that throughout when reporting a 'SMD' we are reporting the bias-corrected version.
 
@@
 
-[^4]: For one clear example, see figure 1A in Vigostky et al.
+[^4]: Though notably, in the case of health behaviour studies it may be the case that if someone volunteers for a study it could conceivably motivate them to alter various habits even when they are assigned to a control group thus influencing change scores.
+
+[^5]: For one clear example, see figure 1A in Vigostky et al.
 [-@vigotsky_improbable_2020] who show that the mean and standard deviation for baseline strength values typically scale with one another across most studies.
 
-[^5]: The authors of the meta-analysis did not make their extracted data openly available, not respond to our request for the extracted data.
+[^6]: The authors of the meta-analysis did not make their extracted data openly available, nor did they respond to our request for the extracted data.
 Further, their original analysis included 119 studies however we were unable to extract data for our analyses from 8 of these for a variety of reasons (e.g., only percentage change data was reported, no standard deviations for control groups reported).
 
-[^6]: Note, we also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases.
+[^7]: It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental intervention, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure).
+We presented the SMD and $\textrm{ln}RR$ effect sizes in equations \@ref(eq:SMD-eq) and \@ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., -@gurevitch_interaction_2000; Morris et al., -@morris_direct_2007; Morris -@morris_estimating_2008) and these are the effect sizes used in the meta-analyses referred to here.
+
+[^8]: We also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases.
 There did not appear to be any obvious small study bias in the dataset (see https://osf.io/stqr3).
 
-[^7]: Data was coded such that study and group had explicit nesting.
+[^9]: We use the term *arm* to refer to an intervention group-control group contrast to accomodate studies including multiple intervention groups.
+This is so as to not confuse the reader with the use of *group* to designate either the RT intervention group(s) or control group separately.
+Thus, in the instances of models using effect sizes relating to comparisons between an intervention group and control group (i.e., SMD, $\textrm{ln}RR$, $SD_{ir}$, $\textrm{ln}VR$, and $\textrm{ln}CVR$) we calculate comparisons between each intervention group (i.e., arm) and the control group.
+Thus, where a study had for example two RT interventions and a control, two separate arms would be coded (RT intervention 1 compared to control, and RT intervention 2 compared to control).
+Data was coded such that study and arm had explicit nesting.
 
-[^8]: See supplementary materials (https://osf.io/e6vpr) for model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continous predictors for both strength and hypertrophy outcomes.
+[^10]: In contrast to the models examining effect sizes relating to comparisons between an intervention group and control group, in the models examining $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ as a predictor the term *arm* refers to both intervention group(s) and control group.
+Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2n and control).
+Data was again coded such that study and arm had explicit nesting.
+
+[^11]: We do not have to limit ourselves to only fixed effect predictor terms as we have here.
+Indeed, for mixed effect models generally some argue that models should use a maximal random effects structure including both random intercepts and slopes to enhance generalisability of inferences [@barr_random_2013].
+We could model a categorical variable for the outcome type and using random effects and include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
+Indeed, we fit additional models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
+The comparison of these models is included in the supplementary materials (ADD LINK).
+Note, the addition of both random slopes for study, and for arm, improved model fit significantly, though we limit presentation in the main text to the simpler model.
+
+[^12]: Note, as with the models examinine outcome upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
+The comparison of these models is included in the supplementary materials (ADD LINK).
+In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy the addition of both random slopes improved model fit significantly.
+Though again we limit presentation in the main text to the simpler model as substantively conclusions were the same.
+
+[^13]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$ (used as results for $\textrm{ln}CVR$ and the random slope model were similar) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
 There were no obvious moderators of $\textrm{ln}CVR$ in particular.
+
+[^14]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study estimate have very large sampling errors.

### 2022-09-22 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Update rmd with links and update trackdown

@@
 Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework.
-However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; -@hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015).
+However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; -@hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015; @senior_revisiting_2020).
+We present the independent groups versions due to use of randomised controlled trials in our examples below, but note that dependent versions also exist for $\textrm{ln}VR$ and $\textrm{ln}CVR$.
 
@@
 Indeed, we fit additional models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
-The comparison of these models is included in the supplementary materials (ADD LINK).
+The comparison of these models is included in the supplementary materials (https://osf.io/4xrcg).
 Note, the addition of both random slopes for study, and for arm, improved model fit significantly, though we limit presentation in the main text to the simpler model.
@@
 [^12]: Note, as with the models examinine outcome upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
-The comparison of these models is included in the supplementary materials (ADD LINK).
+The comparison of these models is included in the supplementary materials (strength - https://osf.io/n4wgk; hypertrophy - https://osf.io/hf8dx).
 In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy the addition of both random slopes improved model fit significantly.

### 2022-09-24 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R
using t-distribution
using t-distribution to reduce the Type I error rate and research nominal coverage rate of confidence intervals

@@ -87,7 +87,7 @@ Data <- escalc(measure = "ZCOR", ri = RT_ri, ni = RT_n, data = Data)
 Meta_RT_ri <- rma.mv(yi, V=vi, data=Data,
                      slab=paste(label),
                      random = list(~ 1 | study, ~ 1 | arm), method="REML",
-                     control=list(optimizer="optim", optmethod="Nelder-Mead"))
+                     control=list(optimizer="optim", optmethod="Nelder-Mead")) # probably need to use improved inference methods - I meant to use t-distrubution rather than z-distrubution to test the significance of individual model cefficients
 
 RobuEstMeta_RT_ri <- robust(Meta_RT_ri, Data$study)
 

### 2022-10-18 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Addressing Yefeng's comments

@@
   - Shinichi Nakagawa$^d$
-date: "22nd September 2022"
+date: "18th October 2022"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
@@
 Most commonly, a magnitude based[^1] effect size statistic [@caldwell_case_2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
-This statistic is usually Cohen's *d* [@cohen_statistical_1988], or it's bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007] [^2].
+This statistic is usually Cohen's *d* [@cohen_statistical_1988], or its bias-corrected metric referred to as Hedges' *g* [@hedges_statistical_2014; @borenstein_introduction_2021; @nakagawa_effect_2007] [^2].
 The SMD, and its sampling variance, $s^2_{SMD}$ are given by:
@@
 
-Although researchers in sport and exercise science among other fields have focused on estimating the average effects of interventions using randomised trial designs for both primary research and synthesis through meta-analysis, evidence suggests that responses to certain interventions may vary on a subgroup or even individual basis.
+Although researchers in sport and exercise science, among other fields, have focused on estimating the average effects of interventions using randomised trial designs for both primary research and synthesis through meta-analysis, responses to certain interventions may vary on a subgroup or even individual basis.
 The increased interest in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@hecksteden_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @ross_precision_2019; @swinton_statistical_2018; @hopkins_individual_2015; @kelley_precision_2022; @hrubeniuk_directions_2022; @pickering_non-responders_2019].
-However, despite the availability of approaches to compare variances between groups, in sport and exercise science this is rarely explored in primary research [@bonafiglia_interindividual_2022] and, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
+However, despite the availability of approaches to compare variances between groups, in sport and exercise science this is rarely explored in primary research [@bonafiglia_interindividual_2022].
+Moreover, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
 Examination of interindividual heterogeneity in response to interventions presents considerable value to researchers and practitioners in sport and exercise science; interventions with low interindividual variation are likely to be widely generalisable, whilst an intervention with high interindividual variation is likely to have effects that are either subgroup or individual specific.
@@
 
-Comparison of heterogeneity in responses such as post-scores or change scores to interventions are however not the only possible use of statistical methods for comparing variances.
+Comparison of heterogeneity in responses, such as post-scores or change scores to interventions, are not the only possible use of statistical methods for comparing variances.
 For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawa_mean_2012].
-Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between participant acute response variation for the purposes of identifying methods[^3] to reduce RT stimulus heterogeneity [@exner_does_2022] as well as meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperin_accuracy_2022].
+Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^3] to reduce RT stimulus heterogeneity, [@exner_does_2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperin_accuracy_2022].
 
 Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet the lack of application in research synthesis by way of meta-analysis, we present and discuss effect size approaches and models for meta-analysis of variation.
-Indeed, meta-analysis presents an incredibly valuable method for exploring variation in a field such as sport and exercise science due to the typically small samples in primary studies.
+Indeed, meta-analysis presents a very valuable method for exploring variation in a field such as sport and exercise science due to the typically small samples in primary studies.
 Such small samples have even lower statistical power to detect differences in variation as compared to means [@yang_low_2022].
@@
 Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework.
-However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; -@hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015; @senior_revisiting_2020).
+However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; @hedges_sex_1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawa_meta-analysis_2015; @senior_revisiting_2020).
 We present the independent groups versions due to use of randomised controlled trials in our examples below, but note that dependent versions also exist for $\textrm{ln}VR$ and $\textrm{ln}CVR$.
@@
 The standard deviation of change scores from the control group (assuming it is a non-intervention control group and not something like a 'usual-care' group) by contrast is assumed to only reflect random error[^4] [@hecksteden_individual_2015].
-As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention and which might reflect individual responses.
+As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention and that might reflect individual responses.
 Whilst the $SD_{ir}$ has been proposed and used primarily in the context of individual response variation to interventions, it should be noted that this kind of absolute comparison of variance between groups or conditions is not limited to such applications.
@@
 Thus, the $SD_{ir}$ reflects a comparison of the absolute variance in change scores between control and intervention groups.
-However, a potential concern with the $SD_{ir}$ is it's potential to violate assumptions of normality which is not the case for other effect size statistics such as $\textrm{ln}VR$ and $\textrm{ln}CVR$.
+However, a potential concern with the $SD_{ir}$ is its potential to violate assumptions of normality, which is not the case for other effect size statistics such as $\textrm{ln}VR$ and $\textrm{ln}CVR$.
 
 ## Log ratio of standard deviations ($\textrm{ln}VR$)
-A similar effect size statistic for the comparison of absolute variance between groups, and one that has had wide applications in more than just intervention response variability within fields like ecology and evolution, is the $\textrm{ln}VR$ [@nakagawa_meta-analysis_2015; @senior_revisiting_2020].
+A similar effect size statistic for the comparison of absolute variance between groups, and one that has had wide applications in more than just intervention response variability within fields such as ecology and evolution, is the $\textrm{ln}VR$ [@nakagawa_meta-analysis_2015; @senior_revisiting_2020].
 An unbiased estimator of the natural logarithm of a population standard deviation ($\textrm{ln}\sigma$), and its sampling variance, $s^2_{\textrm{ln}\sigma}$ is given by:
@@
 ## Detecting the presence of interindividual response variation to resistance training intervention
-First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^7] meta-analyses to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^8].
+First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^7] meta-analysis to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^8].
 Polito et al.
-[-@polito_moderators_2021] originally used a normal random-effects meta-analysis, however the data we extracted were hierarchical in nature (multiple outcomes measures within each arm, i.e., intervention group(s) and control group, within each study) and thus a multilevel mixed-effects meta-analysis model with cluster-robust variance estimation was used with random intercepts for study and arm[^9].
-We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
-A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produce greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).
+[-@polito_moderators_2021] originally used a normal random-effects meta-analysis.
+However, the data we extracted were hierarchical in nature (multiple outcomes measures within each arm, i.e., intervention group(s) and control group, within each study) and thus a multilevel mixed-effects meta-analysis model [@van_den_noortgate_three-level_2013] with cluster-robust variance estimation [@hedges_robust_2010] was used with random intercepts for study, arm[^9], and effect.
+We then fitted the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
+A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produced greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).
 
@@
 The pattern of results from our models examining SMDs (figure \@ref(fig:forest-SMD-plot)) were similar to those reported by Polito et al.
-[-@polito_moderators_2021] albeit with slightly lower estimates for both outcome types; possibly due to our use of a multilevel mixed-effects meta-analysis model that allowed for each individual effect size to be more appropriately weighted.
-As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[4]`%).
-Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between study variance.
+[-@polito_moderators_2021], albeit with slightly lower estimates for both outcome types; possibly due to our use of a multilevel mixed-effects meta-analysis model that allowed for each individual effect size to be more appropriately weighted.
+As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%).
+Confidence intervals were precise for both outcomes, though prediction intervals for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance.
 
@@
 For the $\textrm{ln}RR$ results we exponentiated them and converted to percentages to be more interpretable.
-These were similar with greater proportional increases in strength compared with hypertrophy (figure \@ref(fig:forest-RR-plot)).
-Increases were seen for both strength ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[2]` [95%CI: `r SMD_RR_SDir_logVR$Lower[2]` to `r SMD_RR_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[2]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[2]`%) and hypertrophy outcomes ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[5]`%).
-Confidence intervals were precise for both outcomes again, though notably there was essentially no between study or arm variance in estimates.
+These were similar, with greater proportional increases in strength compared with hypertrophy (figure \@ref(fig:forest-RR-plot)).
+Increases were seen for both strength ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[2]` [95%CI: `r SMD_RR_SDir_logVR$Lower[2]` to `r SMD_RR_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[2]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[2]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[2]`%) and hypertrophy ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[6]` [95%CI: `r SMD_RR_SDir_logVR$Lower[6]` to `r SMD_RR_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[6]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[6]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[6]`%).
+Confidence intervals were again precise for both outcomes, and whilst relative heterogeneity was lower compared to SMD models prediction intervals were still quite wide.
 
-In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I2.arm[8]`%) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation.
+In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[3]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[4]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[7]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[8]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[8]`%) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation.
+Although, heterogeneity across the models and levels (study, arm, effect) were again relatively large and quite varied.
 
@@
 However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean.
-As we have seen from the SMD analysis RT interventions increase mean scores.
+As we have seen from the SMD and $\textrm{ln}RR$ models, RT interventions increase mean scores.
 Thus, if there is a mean-variance relationship in the data, an increase in the mean alone may be fully responsible for any apparent increase in variation.
@@
 We can explore the relationships between variance in a number of outcomes and the impact of certain predictors on this.
-For example, as noted one possible predictor of variance is the mean itself.
+For example, as noted, one possible predictor of variance is the mean itself.
 As such, we can model variance as the response itself.
-The standard deviation is however bounded at zero and so in many cases it may not conform to assumptions of normality.
-Therefore, we instead can use $\textrm{ln}\hat\sigma$ which is unbounded.
+The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality.
+Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
 In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al.
@@
 \begin{equation}
-\textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
+\textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
 (\#eq:lnm-lns-model-eq)
@@
 
-where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^10]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, and $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
+where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^10]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
 
@@
 Figure \@ref(fig:model-mean-variance-pre-plot) shows this model fit visually.
-Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight systematically greater degree of variance in strength measures compared to hypertrophy for a given mean score.
+Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight but systematically greater degree of variance in strength measures compared to hypertrophy for a given mean score.
 
@@
 Returning to the context of interindividual response variation to interventions, the presence of a mean-variance relationship in the data would imply that we cannot rely on absolute comparisons of variance (i.e., $SD_{ir}$ or $\textrm{ln}VR$) to determine whether interindividual response variation is actually present.
-So we should also explore this for the change-scores in the RT and control groups and determing the appropriate effects to explore.
+So we should also explore this for the change-scores in the RT and control groups and determining the appropriate effects to explore.
 
 ## Reanalysis of interindividual response variation using $\textrm{ln}CVR$ and the meta-regression of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$
-As can be seen in figure \@ref(fig:mean-variance-delta-plot)(A) and (C) there is a mean-variance relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
-Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figure \@ref(fig:mean-variance-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seens with the pre-intervention scores.
+As can be seen in figures \@ref(fig:mean-variance-delta-plot)(A) and (C) there is a mean-variance relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
+Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures \@ref(fig:mean-variance-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seens with the pre-intervention scores.
 As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue though.
-Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study and arm as before using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons.
-The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I2.arm[28]`) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I2.study[88]`, $I^2_{arm}$ = `r mods_SMD_logCVR$I2.arm[88]`) and further there is lower relative heterogeneity between studies in the effect estimates (figure \@ref(fig:forest-logCVR-plot).
+Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons.
+The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[28]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[28]`%) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[88]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[88]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[88]`%) and further there is lower relative heterogeneity between studies in the effect estimates (figure \@ref(fig:forest-logCVR-plot)).
 
@@
 
-There is however a potential limitation for the $\textrm{ln}CVR$ also that may need to be considered.
+There is, however, a potential limitation for the $\textrm{ln}CVR$ also that may need to be considered.
 Firstly, it is limited to the use of ratio scale data (which is not the case for the $\textrm{ln}\hat\sigma$ or $\textrm{ln}VR$); hence the need to transform the change scores to be positively signed.
@@
 \begin{equation}
-\textrm{ln}CVR_{ijk}=\mu+\tau_{i}+\tau_{j}+m_{ij} 
+\textrm{ln}CVR_{ijk}=\mu+\tau_{i}+\tau_{j}+\tau_{k}+m_{ij} 
 (\#eq:lnCVR-model-eq)
@@
 
-where $\textrm{ln}CVR_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study and $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, which are assumed to be normall distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ijk}}$.
+where $\textrm{ln}CVR_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnCVR-eq), for the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\mu$ is the intercept or overall mean, $\tau_{i}$ is the deviation from $\mu$ for the $i\textrm{th}$ study, $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, and $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect which are assumed to be normally distributed around zero with variance of $\sigma^2_{\tau}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}CVR_{ijk}}$.
 
-Instead, we can use a version of the model described above (equation \@ref(eq:lnm-lns-model-eq) and the paragraph which follows it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
+Instead, we can use a version of the model described above (see equation \@ref(eq:lnm-lns-model-eq) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
@@
 
-Hopefully it is clear from the regression model here where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework.
-Models can be fit to explore not only how study and arm level characteristics moderate effect size estimates when considering not only effects sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance based effect size statistics and models employed in this article[^13].
+Hopefully it is clear from the regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors (i.e., meta-regression) when exploring variance through a meta-analytic framework.
+Models can be fitted to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering not only effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article[^13].
 
@@
 It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
-Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely increased variance above random error, suggesting the presence of inter-individual response variation.
+Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation.
 In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive, yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
@@
 However, similar to the cross-sectional pre-intervention scores reported here (and indeed most physical and biological variables), change scores demonstrated a mean-variance relationship in addition to heteroskedasticity.
-The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ in addition to the random slope model of $\textrm{ln}\hat\sigma$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance implying that there is likely little to no interindividual response variation to explain.
+The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain.
 Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usui_meta-analysis_2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortes_martinez_constant_2021].
 Interventions such as these are valuable for the simplification of guidelines and recommendations.
-For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to recommendations [@steele_higher_2017; @steele_long-term_2022].
+For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations [@steele_higher_2017; @steele_long-term_2022].
 
@@
 Whether there is evidence of meaningful average intervention effects or not, where considerable variance effects are present it suggests that a meaningful line of research would be to aim at identifying subgroup- or participant-by-intervention interactions using appropriate study designs [@hecksteden_individual_2015].
-Where variance effects are limited this instead suggests that translational work towards generalisable implementation might be most the most meaningful line of future research.
+Where variance effects are limited this instead suggests that translational work towards generalisable implementation might be the most meaningful line of future research.
 Finally, there may be cases where it is in fact desirable to identify interventions that actually reduce variance; for example, improvements in methodological approaches to enhance research [@usui_meta-analysis_2021], or interventions to reduce variation in sport performances.
 Thus, researchers in sport and exercise science should consider exploring variance more systematically, and indeed utilise the meta-analytic framework to support this.
-This could include the re-analysis of past meta-analysis as we have done so here, and indeed researchers conducted future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawa_meta-analysis_2015; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @usui_meta-analysis_2021; @mills_detecting_2021] presented here and echoing the efforts of other recent work [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
+This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawa_meta-analysis_2015; @hopkins_individual_2015; @atkinson_true_2015; @atkinson_issues_2019; @usui_meta-analysis_2021; @mills_detecting_2021] presented here and echoing the efforts of other recent work [@kelley_are_2022; @kelley_are_2020; @esteves_individual_2021; @bonafiglia_interindividual_2022; @steele_slow_2021; @fisher_role_2022].
 
@@
 [^7]: It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental intervention, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure).
-We presented the SMD and $\textrm{ln}RR$ effect sizes in equations \@ref(eq:SMD-eq) and \@ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., -@gurevitch_interaction_2000; Morris et al., -@morris_direct_2007; Morris -@morris_estimating_2008) and these are the effect sizes used in the meta-analyses referred to here.
+We presented the SMD and $\textrm{ln}RR$ effect sizes in equations \@ref(eq:SMD-eq) and \@ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., [-@gurevitch_interaction_2000]; Morris et al., [-@morris_direct_2007]; Morris [-@morris_estimating_2008]; Lajeunesse [-@lajeunesse_bias_2015]) and these are the effect sizes used in the meta-analyses referred to here.
 
@@
 [^10]: In contrast to the models examining effect sizes relating to comparisons between an intervention group and control group, in the models examining $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ as a predictor the term *arm* refers to both intervention group(s) and control group.
-Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2n and control).
+Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2, and control).
 Data was again coded such that study and arm had explicit nesting.
@@
 Indeed, for mixed effect models generally some argue that models should use a maximal random effects structure including both random intercepts and slopes to enhance generalisability of inferences [@barr_random_2013].
-We could model a categorical variable for the outcome type and using random effects and include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
+We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
 Indeed, we fit additional models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
@@
 
-[^14]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study estimate have very large sampling errors.
+[^14]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.

### 2022-10-26 paper/Pre-print - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Add pre-print version - sportrxiv citation added

--- 2022-10-18 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
@@
 thanks: |
-  Preprint, submitted to Journal of Sport Sciences. 
+  Preprint, please cite as: Steele, J., Fisher, J., Smith, D., Schoenfeld, B., Yang, R., & Nakagawa, S.
+(2022).
+Meta-Analysis of Variation in Sport and Exercise Science: Examples of Application Within Resistance Training Research.
+$\textrm{sportr}\chi\textrm{iv}$ DOI: ????.
   Address for correspondence: james.steele@solent.ac.uk

### 2022-10-26 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
Add anonymous version for submission

--- 2022-10-18 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
@@
 subtitle: "Examples of Application Within Resistance Training Research"
-author: 
-  - James Steele$^a$
-  - James Fisher$^a$
-  - Dave Smith$^b$
-  - Brad Schoenfeld$^c$
-  - Yefeng Yang$^d$
-  - Shinichi Nakagawa$^d$
 date: "18th October 2022"
@@
 Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet its lack of application in research synthesis by way of meta-analysis, we introduce and discuss effect size approaches and models for meta-analysis of variation using relatable examples from resistance training RCTs."
-thanks: |
-  Preprint, submitted to Journal of Sport Sciences. 
-  Address for correspondence: james.steele@solent.ac.uk
-  Affiliations:
-  $^a$Faculty of Sport, Health, and Social Sciences, Solent University, UK;
-  $^b$Research Centre for Musculoskeletal and Sports Medicine, Manchester Metropolitan University, Manchester, UK;
-  $^c$Health Sciences Department, CUNY Lehman College, Bronx, New York, USA;
-  $^d$Evolution & Ecology Research Centre and School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, New South Wales, Australia
+
 output: bookdown::pdf_document2

### 2023-09-07 paper/Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
Addressing round 1 JSS revisions

@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "18th October 2022"
+date: "????"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
 However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance.
-It is well known that there is considerable variability in response to interventions.
+It is well known that \textcolor{red}{individual responses to interventions vary considerably}.
 Hence, interest has increased in exploring *precision* or *personalised* exercise approaches.
 Not only is the mean often affected by interventions, but variances may also be impacted.
-Exploration of variances in studies such as randomised controlled trials (RCTs) can yield insight into the existence of interindividual heterogeneity in response to interventions and help determine the degree of generalisability of effects.
+Exploration of variances in studies such as randomised controlled trials (RCTs) can yield insight into interindividual heterogeneity in response to interventions and help determine generalisability of effects.
 Yet, larger samples sizes than those used for typical mean effects are required when probing variances.
-Thus, for a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
-Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet its lack of application in research synthesis by way of meta-analysis, we introduce and discuss effect size approaches and models for meta-analysis of variation using relatable examples from resistance training RCTs."
-
+Thus, \textcolor{red}{in} a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
+\textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}."
 output: bookdown::pdf_document2
@@
    - \captionsetup[figure]{font=small}
+   - \usepackage{lineno}
+   - \linenumbers
 ---
@@
 ```
-```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
 # Load relevant packages
 library(here)
+library(tidyverse)
 library(kableExtra)
+library(patchwork)
+library(ggtext)
 ```
@@
 # load the raw data to the environment
-Data <- read.csv(here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))
+Data <- read.csv(here::here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))
 

### 2023-09-07 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
Addressing round 1 JSS revisions

---
title: "Meta-Analysis of Variation in Sport and Exercise Science"
subtitle: "Examples of Application Within Resistance Training Research"
date: "7th September 2023"
abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies. However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance. It is well known that \\textcolor{red}{individual responses to interventions vary considerably}. Hence, interest has increased in exploring *precision* or *personalised* exercise approaches. Not only is the mean often affected by interventions, but variances may also be impacted. Exploration of variances in studies such as randomised controlled trials (RCTs) can yield insight into interindividual heterogeneity in response to interventions and help determine generalisability of effects. Yet, larger samples sizes than those used for typical mean effects are required when probing variances. Thus, \\textcolor{red}{in} a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing. \\textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \\textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}."
output: bookdown::pdf_document2
bibliography: references.bib
csl: apa.csl
toc: false
header-includes:
   - \usepackage{caption}
   - \captionsetup[figure]{font=small}
   - \usepackage{lineno}
   - \linenumbers
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
options(knitr.kable.NA = '')
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Load relevant packages
library(here)
library(tidyverse)
library(kableExtra)
library(patchwork)
library(ggtext)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# load the raw data to the environment
Data <- read.csv(here::here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))

# load the summary dataframes to the environment
load(here("models/SMD_RR_SDir_logVR"))
load(here("models/mods_SMD_logCVR"))

```
# Introduction
Although the quantitative synthesis of results across studies has existed since the 17th century [@plackettStudiesHistoryProbability1958], the modern-day term "meta-analysis" was coined by Gene Glass [-@glassPrimarySecondaryMetaAnalysis1976]. Since that time, the use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@haggerMetaanalysis2022], with resistance training (RT) accounting for a considerable proportion of this growth (figure \@ref(fig:trend-plot)). Accordingly, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers. \textcolor{red}{However, to begin we provide a conceptual overview of meta-analyses and effect sizes as typically used within sport and exercise science.}

```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976."}

load(here("plots/trends_plot"))
trends_plot

```

\textcolor{red}{There are two popular models}[^1] \textcolor{red}{used for meta-analysis: the fixed-effect model and the random-effects model} [@borensteinBasicIntroductionFixedeffect2010]. \textcolor{red}{The fixed effect model assumes that there is one true effect size}[^2] \textcolor{red}{that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error. This essentially means that there is a single common effect which is fixed across studies and each study takes samples of individuals from the population to estimate this effect. We can express this model in the following formula:}

[^1]: \textcolor{red}{To clarify language here for those unfamiliar, the term and concept 'model' is used commonly in statistics. A statistical model essentially is specification of what we think the data generating process might be for a given situation. In the context of meta-analyses the data are usually the individual effects that we have extracted from studies i.e., the results of each study. The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.}

[^2]: \textcolor{red}{'Effect size' is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research. This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between} [@caldwellCaseDefaultEffect2020] \textcolor{red}{including, as we shall see, statistics describing variation.}

\begin{equation}
\hat\theta_{i}=\theta+m_{i} 
(\#eq:fixed-model-eq)
\end{equation}

\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$. Studies with smaller standard errors of their effect estimates have smaller sampling errors and so these higher precision estimates are given greater weight in the model. The weighting given to each study is calculated as:}

\begin{equation}
w_{i}=\frac{1}{s^2_{i}}
(\#eq:fixed-weights-eq)
\end{equation}

\textcolor{red}{Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance. It is referred to as inverse-variance weighting. Then, the overall weighted mean effect estimate from the model is then calculated as:}

\begin{equation}
\hat\theta=\frac{\sum^I_{i=1}\hat\theta_{i}w_{i}}{\sum^I_{i=1}w_{i}}
(\#eq:fixed-average-eq)
\end{equation}

\textcolor{red}{Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect. Instead it allows for the true effects that each study estimates to differ. Each study may share a common underlying effect size but, due to differences between studies in factors such as population characteristics, the manner in which outcomes are operationalised, or subtle differences in intervention and context to name a few, it is possible that the actual effect being estimated by each study differs. The assumption of the random-effects model then is that the studies included estimate effects that come from a larger population of effects determined by the inclusion criteria for studies included. So, the model assumes that the studies included reflect a random sample of all possible permutations of study and the effects they estimate from this population distribution of studies and effects. Hence, in the fixed-effect model there is one effect and it is assumed to be *fixed* across studies, whereas in the random-effects model there are many and we examine an assumed *random* sample of them. We can express this model in the following formula:}

\begin{equation}
\hat\theta_{i}=\theta_{i}+\tau_{i}+m_{i}
(\#eq:random-model-eq)
\end{equation}

\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$. Essentially this model assumes that each individual study estimates an effect and there will be some sampling error in estimating it. But, the effects each study estimates comes from an overall distribution of true effects with a mean value (i.e., $\theta_{i}$). The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated}[^3]. \textcolor{red}{This essentially describes the heterogeneity between the effects included. Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:}

[^3]: \textcolor{red}{This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform. This is beyond the scope of this paper to discuss. We note however that the models we present all utilise Restricted Maximum Likelihood estimation.}

\begin{equation}
w^*_{i}=\frac{1}{s^2_{i}+\tau^2}
(\#eq:random-weights-eq)
\end{equation}

\textcolor{red}{Then the average of the distribution of effects can be calculated as per \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.}

\textcolor{red}{Within sport and exercise science, although historically fixed effects models were more common} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays, likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above, the fixed-effect model is less commonly employed than the random-effects model} [@haggerMetaanalysis2022].

[^4]: \textcolor{red}{Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).}

As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] \textcolor{red}{likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable. The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping}; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex. Indeed, a recent umbrella review [@bernardez-vazquezResistanceTrainingVariables2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes, all of which focused on the comparison of mean changes between different intervention groups.

\textcolor{red}{Often, due to the varying operationalisations used for broad outcome concepts}[^5]\textcolor{red}{, an effect size is used which is "standardised" across studies to allow for their synthesis.} Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions. This statistic is usually a version of Cohen's *d* [@cohenStatisticalPowerAnalysis1988], or its bias-corrected[^7] metric referred to as Hedges' *g* [@hedgesStatisticalMethodsMetaAnalysis2014; @borensteinIntroductionMetaAnalysis2021; @nakagawaEffectSizeConfidence2007][^8]. The SMD, and its sampling variance, $s^2_{SMD}$ are given by:

[^5]: \textcolor{red}{For example, 'strength' might be examined in different studies using different operationalisation including one repetition maximum testing or maximum voluntary contractions. Or the same operationalisations may be employed but different exercises such as the squat or bench press.}

[^6]: Though notably not all meta-analyses use *magnitude based* effect sizes. Indeed some explicitly use what Caldwell and Vigotsky [-@caldwellCaseDefaultEffect2020] term *signal-to-noise* effect sizes (e.g., @heidelMachinesFreeWeight2022).

[^7]: \textcolor{red}{For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct. Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.}

[^8]: We will refer to both merely as the SMD throughout the manuscript for simplicity and note that throughout when reporting a 'SMD' we are reporting the bias-corrected version. \textcolor{red}{We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance;} @caldwellCaseDefaultEffect2020).

\begin{equation}
SMD=\frac{\overline{x}_E - \overline{x}_C}{s_{pooled}}J
(\#eq:SMD-eq)
\end{equation}

\begin{equation}
J=1-\frac{3}{4(n_{C}+n_{E})-2)-1}
(\#eq:biasJ-eq)
\end{equation}

\begin{equation}
s_{pooled}=\sqrt{\frac{(n_{C}-1)s^2_{C}+(n_{E}-1)s^2_{E}}{n_{C}+n_{E}-2}}
(\#eq:s-pooled-eq)
\end{equation}

\begin{equation}
s^2_{SMD}=\frac{n_{C}+n_{E}}{n_{C}n_{E}}+\frac{SMD^2}{2(n_{E}+n_{C})}
(\#eq:SMDvar-eq)
\end{equation}

where $\overline{x}_C$ and $\overline{x}_E$ are the sample means of the control group (C) and experimental (E) or intervention group respectively, $s_C$ and $s_E$ are the standard deviations of the two groups, $n_C$ and $n_E$ are the sample sizes of the two groups, and $J$ is a bias correction for small sample sizes.

The natural logarithm of the ratio of two means ($\textrm{ln}RR$) is also another effect size statistic that can be used [@curtisMetaanalysisElevatedCO21998; @hedgesMetaAnalysisResponseRatios1999; @lajeunesseMetaanalysisResponseRatios2011; @lajeunesseBiasCorrectionLog2015]. The lnRR, and its sampling variance, $s^2_{\textrm{ln}RR}$ are given by:

\begin{equation}
\textrm{ln}RR=\textrm{ln}\frac{\overline{x}_E}{\overline{x}_C}
(\#eq:lnRR-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}RR}=\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}+\frac{s^4_{C}}{2n_{C}\overline{x}^2_{C}\overline{x}^4_{C}}+\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}}+\frac{s^4_{E}}{2n_{E}\overline{x}^2_{E}\overline{x}^4_{E}}
(\#eq:lnRRvar-eq)
\end{equation}

Due to its calculation the SMD is affected not only by the difference in means of the two groups, but also by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$ in the denominator. In contrast, the $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), which only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)). Despite this, the use of effects sizes like the $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@debQuantifyingEffectsAcute2018; @nuzzoEccentricConcentricStrength2023] and to our knowledge only one meta-analysis of RT interventions has used this kind of effect size [@swintonInterpretingMagnitudeChange2022].

Although researchers in sport and exercise science, among other fields, have focused on estimating the average effects of interventions using randomised trial designs for both primary research and synthesis through meta-analysis, responses to certain interventions may vary on a subgroup or even individual basis. The increased interest in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@heckstedenIndividualResponseExercise2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @rossPrecisionExerciseMedicine2019; @swintonStatisticalFrameworkInterpret2018; @hopkinsIndividualResponsesMade2015; @kelleyPrecisionExerciseMedicine2022; @hrubeniukDirectionsExerciseTreatment2022; @pickeringNonRespondersExerciseExistand2019]. However, despite the availability of approaches to compare variances between groups, in sport and exercise science this is rarely explored in primary research [@bonafigliaInterindividualDifferencesTrainability2022]. Moreover, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022]. Examination of interindividual heterogeneity in response to interventions presents considerable value to researchers and practitioners in sport and exercise science; interventions with low interindividual variation are likely to be widely generalisable, whilst an intervention with high interindividual variation is likely to have effects that are either subgroup or individual specific. The former kind of intervention might be widely applicable across individuals, whilst the latter kind of intervention requires specific research, typically with large samples [@heckstedenIndividualResponseExercise2015], to tease out subgroup- or participant-by-intervention interactions to facilitate successful practical application.

Comparison of heterogeneity in responses, such as post-scores or change scores to interventions, are not the only possible use of statistical methods for comparing variances. For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawaMeanStrikesBack2012]. Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] \textcolor{red}{and in the number of repetitions that can be performed at different percentages of one repetition maximum} [@nuzzoMaximalNumberRepetitions2023].

[^9]: Exploration of methodological approaches and their impact on heterogeneity have also been explored in preclinical research [@usuiMetaanalysisVariationSuggests2021].

Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet the lack of application in research synthesis by way of primary research or meta-analysis, we present and discuss effect size approaches and models for meta-analysis of variation. Indeed, meta-analysis presents a very valuable method for exploring variation in a field such as sport and exercise science due to the typically small samples in primary studies. Such small samples have even lower statistical power to detect differences in variation as compared to means [@yangLowStatisticalPower2022].

# Effect size statistics for meta-analytic comparisons of variation
Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework. However, several have been proposed that we now describe: the standard deviation for individual responses ($SD_{ir}$; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019), the log ratio of standard deviations ($\textrm{ln}VR$; termed the "variability ratio"; @hedgesSexDifferencesMental1995), and the log ratio of coefficient of variation ($\textrm{ln}CVR$; termed the "coefficient of variation ratio"; @nakagawaMetaanalysisVariationEcological2015; @seniorRevisitingExpandingMetaanalysis2020). We present the independent groups versions due to use of randomised controlled trials in our examples below, but note that dependent versions (i.e., for comparing related samples) also exist for $\textrm{ln}VR$ and $\textrm{ln}CVR$ [@seniorRevisitingExpandingMetaanalysis2020].

## Standard deviation for individual responses ($SD_{ir}$)
In the context of *precision* or *personalised* approaches to exercise prescription the $SD_{ir}$ has been proposed as an approach to determine the extent to which individual responses manifest by comparison of variation between two groups; control and intervention [@hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019]. The standard deviation of change scores (post-intervention scores minus pre-intervention scores) within the intervention group reflects the gross combination of a number of sources of variation including: participant-by-intervention interactions (i.e., actual individual responsiveness or 'trainability'), within-participant variability in intervention response (i.e., variability in response to the same intervention administered to the same participant), and random error (i.e., from pre and post measurements; @heckstedenIndividualResponseExercise2015). The standard deviation of change scores from the control group (assuming it is a non-intervention control group and not something like a 'usual-care' group) by contrast is assumed to only reflect random error[^10] [@heckstedenIndividualResponseExercise2015]. As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention and that might reflect individual responses. Whilst the $SD_{ir}$ has been proposed and used primarily in the context of individual response variation to interventions, it should be noted that this kind of absolute comparison of variance between groups or conditions is not limited to such applications.

[^10]: Though notably, in the case of health behaviour studies it may be the case that if someone volunteers for a study it could conceivably motivate them to alter various habits even when they are assigned to a control group thus influencing change scores.

The $SD_{ir}$, and its sampling variance, $s^2_{SD_{ir}}$ are given by:

\begin{equation}
SD_{ir}=\sqrt{s^2_{E} - s^2_{C}}
(\#eq:SDir-eq)
\end{equation}

\begin{equation}
s^2_{SD_{ir}}=2\biggl(\frac{s^4_{E}}{n_{E}-1}+\frac{s^4_{C}}{n_{C}-1}\biggl)
(\#eq:sdIRvar-eq)
\end{equation}

Thus, the $SD_{ir}$ reflects a comparison of the absolute variance in change scores between control and intervention groups. However, a potential concern with the $SD_{ir}$ is its potential to violate assumptions of normality, which is not the case for other effect size statistics such as $\textrm{ln}VR$ and $\textrm{ln}CVR$.

## Log ratio of standard deviations ($\textrm{ln}VR$)
A similar effect size statistic for the comparison of absolute variance between groups, and one that has had wide applications in more than just intervention response variability within fields such as ecology and evolution, is the $\textrm{ln}VR$ [@hedgesSexDifferencesMental1995; @nakagawaMetaanalysisVariationEcological2015; @seniorRevisitingExpandingMetaanalysis2020]. An unbiased estimator of the natural logarithm of a population standard deviation ($\textrm{ln}\sigma$), and its sampling variance, $s^2_{\textrm{ln}\sigma}$ is given by:

\begin{equation}
\textrm{ln}\hat\sigma=\textrm{ln}s+\frac{1}{2(n-1)}
(\#eq:lnsigma-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}\hat\sigma}=\frac{1}{2(n-1)}
(\#eq:lnsigmavar-eq)
\end{equation}

where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$. Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:

\begin{equation}
\textrm{ln}VR=\textrm{ln}\biggl(\frac{s_{E}}{s_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnVR-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}VR}=\frac{1}{2}\biggl(\frac{n_{C}}{(n_{C}-1)^2}+\frac{n_{E}}{(n_{E}-1)^2}\biggl)
(\#eq:lnVRvar-eq)
\end{equation}

However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparison group (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$). In this case, it is likely that the standard deviation will be larger in the group with the larger mean (e.g., $s_{E}$ is larger than $s_{C}$). This mean-variance relationship is common for many variables and datasets[^11] and to highlight this we provide examples below. They also assume constant measurement error over the range of values for the mean, which can impact their utility for examining response variation [@tenanCommentMethodStop2020].

[^11]: For one clear example, see figure 1A in Vigostky et al. [-@vigotskyImprobableDataPatterns2020] who show that the mean and standard deviation for baseline strength values typically scale with one another across most studies.

## Log ratio of coefficient of variation ($\textrm{ln}CVR$)
The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present. \textcolor{red}{In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.} The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups. Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), the $\textrm{ln}CVR$ is given by:

\begin{equation}
\textrm{ln}CVR=\textrm{ln}\biggl(\frac{CV_{E}}{CV_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnCVR-eq)
\end{equation}

where $CV_{E}$ and $CV_{C}$ are ${s_{E}}/{\overline{x}_{E}}$ and ${s_{C}}/{\overline{x}_{C}}$ respectively. Senior et al. [-@seniorRevisitingExpandingMetaanalysis2020] derived the sampling variance, $s^2_{\textrm{ln}CVR}$, as:

\begin{equation}
\begin{split}
s^2_{\textrm{ln}CVR}=\frac{s_{C}^2}{n_{C}\overline{x}_{C}^2}+\frac{s_{C}^4}{2n_{C}^2\overline{x}_{C}^4}+\frac{n_{C}}{(n_{C}-1)^2} \\
+\frac{s_{E}^2}{n_{E}\overline{x}_{E}^2}+\frac{s_{E}^4}{2n_{E}^2\overline{x}_{E}^4}+\frac{n_{E}}{(n_{E}-1)^2}
\end{split}
(\#eq:lnCVRvar-eq)
\end{equation}

# Examples using resistance training studies
As noted, to facilitate understanding for those new to examination of variation, we provide primary examples of the approaches presented using data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021]. Here we have used their list of included studies and re-extracted data from 111 of these[^12]. All analysis examples were performed in R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauerConductingMetaAnalysesMetafor2010]. The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/) \textcolor{red}{or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git)}.

[^12]: The authors of the meta-analysis did not make their extracted data openly available, nor did they respond to our request for the extracted data. Further, their original analysis included 119 studies however we were unable to extract data for our analyses from 8 of these for a variety of reasons (e.g., only percentage change data was reported, no standard deviations for control groups reported).

```{r sample-size-tab, message=FALSE, warning=FALSE, echo=FALSE}

load(here("models/sample_sizes"))

knitr::kable(
  sample_sizes,
  caption = 'Sample sizes for resistance training and non training control groups for dataset.'
  ) %>%
  pack_rows("RT", 1, 4) %>%
  pack_rows("CON", 5, 8) %>%
  footnote(general = c("RT = resistance training", "CON = non-training control")
           ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE)

```

Polito et al. [-@politoModeratorsStrengthGains2021] conducted a systematic review and meta-analysis of randomised trials that included a RT intervention group(s) and a non-training control comparison group. Their analysis focused upon the SMD between the RT intervention group(s) and the control group from the studies included, with both overall effect estimate and moderator analyses (i.e., meta-regressions[^13]) performed. Given that Polito et al. [-@politoModeratorsStrengthGains2021] included only studies with a non-training control group, their study selection offers a unique context to examine variation of interindividual responses specifically by comparing the variances in change scores between the RT intervention groups(s) and control group. Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies. Indeed, this highlights the typically small samples used in sport and exercise science, and thus low power to detect difference in both means and variances in in individual studies [@yangLowStatisticalPower2022], emphasising the value of meta-analysis to explore variation. Table \@ref(tab:summary-characteristics-tab) shows the study and participant characteristics.

\textcolor{red}{Although ultimately we will recommend certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples using all approaches described in order to aid the reader in understanding their strengths and weaknesses. We hope this will make clear why we offer certain recommendations in our discussion As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, and also to provide an explanation of them in plain English.}

[^13]: \textcolor{red}{Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s). This can be extended to meta-analytic synthesis where the independent variables reflect characteristics associated with the effects included. For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.}

```{r summary-characteristics-tab, message=FALSE, warning=FALSE, echo=FALSE}

load(here("models/summary_table"))

knitr::kable(
  summary_table,
  caption = 'Summary of study and participant characteristics.',
  align = c("l","c") 
  ) %>%
  pack_rows("Training Status",6, 7, bold = FALSE) %>%
  pack_rows("Sample Type", 8, 9, bold = FALSE) %>%
  pack_rows("RT + Adjuvant Intervention?", 10, 11, bold = FALSE) %>%
  pack_rows("Task Failure?", 18, 19, bold = FALSE) %>%
  footnote(general = c("RT = resistance training; ", "Continuous variables are median (IQR); ", "Categorical variables are count (%); ", "Not all studies reported full descriptive data (see dataset; https://osf.io/kg2z4)")
           ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE) %>%
    kable_styling()

```

## Detecting the presence of interindividual response variation to resistance training intervention \textcolor{red}{with absolute variance statistics}
First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^14] meta-analysis to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^15]. Polito et al. [-@politoModeratorsStrengthGains2021] originally used a normal random-effects meta-analysis as described in the introduction. However, the data we extracted were hierarchical in nature\textcolor{red}{; as opposed to there being only the assumption that studies are a random effect, due to their being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect. \textcolor{red}{This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies. We can describe the overall model then as:}

[^14]: It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental intervention, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure). We presented the SMD and $\textrm{ln}RR$ effect sizes in equations \@ref(eq:SMD-eq) and \@ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., [-@gurevitchInteractionCompetitionPredation2000]; Morris et al., [-@morrisDirectInteractiveEffects2007]; Morris [-@morrisEstimatingEffectSizes2008]; Lajeunesse [-@lajeunesseMetaanalysisResponseRatios2011; -@lajeunesseBiasCorrectionLog2015]) and these are the effect sizes used in the meta-analyses referred to here.

[^15]: We also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases. There did not appear to be any obvious small study bias in the dataset (see https://osf.io/stqr3).

[^16]: We use the term *arm* to refer to an intervention group-control group contrast to accommodate studies including multiple intervention groups. This is so as to not confuse the reader with the use of *group* to designate either the RT intervention group(s) or control group separately. Thus, in the instances of models using effect sizes relating to comparisons between an intervention group and control group (i.e., SMD, $\textrm{ln}RR$, $SD_{ir}$, $\textrm{ln}VR$, and $\textrm{ln}CVR$) we calculate comparisons *between* each intervention group (i.e., arm) and the control group. Thus, where a study had for example two RT interventions and a control, two separate arms would be coded (RT intervention 1 compared to control, and RT intervention 2 compared to control). Data was coded such that study and arm had explicit nesting.

\begin{equation}
\hat\theta_{ijk}=(\theta+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})++m_{ijk}
(\#eq:ml-model-eq)
\end{equation}

\textcolor{red}{where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$. This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects}[^17]. \textcolor{red}{The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).}

[^17]: \textcolor{red}{Technically then the 'random' effects model presented earlier is also a 'mixed' effects model. It is traditionally referred to as the random-effects model though.}

We then fitted the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls. A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produced greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).

```{r forest-SMD-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes."}

# "Caterpillar plots of SMD effect sizes for strength and hypertrophy outcomes. Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"

load(here("plots/forest_SMD_strength"))
load(here("plots/forest_SMD_hypertrophy"))
library(patchwork)
(forest_SMD_strength / forest_SMD_hypertrophy) +
  plot_annotation(tag_levels = "A")

```

The pattern of results from our models examining SMDs (figure \@ref(fig:forest-SMD-plot)) were similar to those reported by Polito et al. [-@politoModeratorsStrengthGains2021], albeit with slightly lower estimates for both outcome types; possibly due to our use of a multilevel mixed-effects meta-analysis model that allowed for each individual effect size to be more appropriately weighted (the relative amount of heterogeneity between effects for each level is presented as the $I^2$ statistic). As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%). Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}

```{r forest-RR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of exponentiated RR effect sizes for strength (A) and hypertrophy (B) outcomes."}

# "Caterpillar plots of exponentiated RR effect sizes for strength and hypertrophy outcomes. Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"

load(here("plots/forest_RR_strength"))
load(here("plots/forest_RR_hypertrophy"))
library(patchwork)
(forest_RR_strength / forest_RR_hypertrophy) +
  plot_annotation(tag_levels = "A")

```

For the $\textrm{ln}RR$ results we exponentiated them and converted to percentages to be more interpretable. These were similar, with greater proportional increases in strength compared with hypertrophy (figure \@ref(fig:forest-RR-plot)). Increases were seen for both strength ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[2]` [95%CI: `r SMD_RR_SDir_logVR$Lower[2]` to `r SMD_RR_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[2]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[2]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[2]`%) and hypertrophy ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[6]` [95%CI: `r SMD_RR_SDir_logVR$Lower[6]` to `r SMD_RR_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[6]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[6]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[6]`%). Confidence intervals were again precise for both outcomes, and whilst relative heterogeneity was lower compared to SMD models prediction intervals were still quite wide.

In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[3]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[4]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[7]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[8]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[8]`%) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation. Although, heterogeneity across the models and levels (study, arm, effect) were again relatively large and quite varied.

This additional variance might support previous perspectives [@carpinelliInterindividualHeterogeneityAdaptations2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* independent of the random error). However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean \textcolor{red}{(i.e., that the variance is similar whether mean values are low or high)}. As we have seen from the SMD and $\textrm{ln}RR$ models, RT interventions increase mean scores. Thus, if there is a mean-variance relationship in the data, an increase in the mean alone may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present. The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic. First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ \textcolor{red}{and increase its flexibility to accomodate wider applications}.

```{r forest-SDir-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SDir effect sizes for strength (A) and hypertrophy (B) outcomes."}

# "Caterpillar plots of SDir effect sizes for strength and hypertrophy outcomes. Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"

load(here("plots/forest_SDir_strength"))
load(here("plots/forest_SDir_hypertrophy"))
library(patchwork)
(forest_SDir_strength / forest_SDir_hypertrophy) +
  plot_annotation(tag_levels = "A")

```
```{r forest-logVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$VR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}

# "Caterpillar plots of $\textrm{ln}RR$ effect sizes for strength and hypertrophy outcomes. Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"

load(here("plots/forest_logVR_strength"))
load(here("plots/forest_logVR_hypertrophy"))
library(patchwork)
(forest_logVR_strength / forest_logVR_hypertrophy) +
  plot_annotation(tag_levels = "A")

```

## Mean-variance relationships in muscular strength and hypertrophy
With meta-analytic models of variation we are not limited to solely exploring variation in responses to interventions (e.g., @halperinAccuracyPredictingRepetitions2022; @nuzzoMaximalNumberRepetitions2023). We can explore the relationships between variance in a number of outcomes and the impact of certain predictors on this in the form of meta-regression. For example, as noted, one possible predictor of variance is the mean itself. As such, we can model variance of each effect as the response itself with the mean of the effect as the predictor. The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models \textcolor{red}{(i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).} Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded. In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al. [-@politoModeratorsStrengthGains2021].

```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes."}

load(here("plots/mean_variance_pre_plots"))

mean_variance_pre_plots

```

As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$). \textcolor{red}{The variance in standard deviations increase with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].

\begin{equation}
s^2=a\overline{x}^b
(\#eq:taylors-eq)
\end{equation}

where $a$ and $b$ are some constants. When this relationship holds, under most circumstances the standard deviation is not proportional to the mean. However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear \textcolor{red}{based upon the product and power logarithmic rules:}

\begin{equation}
2\textrm{ln}s=\textrm{ln}a+b\textrm{ln}\overline{x}
(\#eq:taylorsln-eq)
\end{equation}

Figure \@ref(fig:mean-variance-pre-plot)(B) and (D) shows that the relationship between the mean and variance on the log scale better meets the assumption of normality. Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from multiple outcomes within multiple arms within studies we can also estimate this relationship using a multilevel mixed-effects meta-regression model similar to that applied above. In this case though we are including an addition *predictor* variable, the $\textrm{ln}\overline{x}$. For example, the following model specifies $\textrm{ln}\overline{x}$ as a fixed effect with random intercepts for study and arm:

\begin{equation}
\textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
(\#eq:lnm-lns-model-eq)
\end{equation}

where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^18]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept or overall mean of the effects, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.

\textcolor{red}{These kinds of models are incredibly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].

[^18]: In contrast to the models examining effect sizes relating to comparisons between and intervention group and control group, in the models examining $\textrm{ln}\sigma_{ijk}$ and $\textrm{ln}\overline{x}_{ijk}$ as a predictor the term *arm* refers to boh the intervention groups(s) and control group. Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2, and control). Data were again coded such that study and arm had explicit nesting.

[^19]: We do not have to limit ourselves to only fixed effect predictor terms as we have here. Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes)}, and their correlations, to enhance generalisability of inferences [@barrRandomEffectsStructure2013]. We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the overall average slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm. Indeed, we fit additional models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm. The comparison of these models is included in the supplementary materials (https://osf.io/4xrcg). Note, the addition of both random slopes for study, and for arm, improved model fit significantly, though we limit presentation in the main text to the simpler model.

Figure \@ref(fig:model-mean-variance-pre-plot) shows this model fit visually where the size of the points reflects their weight in the model. Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight but systematically greater degree of variance in strength measures compared to hypertrophy for a given mean score.

```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores."}

# "Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores. Note, fitted lines are the regression slopes for strength and hypertophy outcomes each with their 95% confidence interval band."

load(here("plots/model_mean_variance_pre_plots"))

model_mean_variance_pre_plots

```

The presence of Taylor's law type relationships should be examined in datasets prior to deciding on which variance effect size statistic should be employed. Returning to the context of interindividual response variation to interventions, the presence of a mean-variance relationship in the data would imply that we cannot rely on absolute comparisons of variance (i.e., $SD_{ir}$ or $\textrm{ln}VR$) to determine whether interindividual response variation is actually present. So we should also explore this for the change-scores in the RT and control groups and determining the appropriate effect sizes to explore.

## Reanalysis of interindividual response variation using $\textrm{ln}CVR$ 
As can be seen in figures \@ref(fig:mean-variance-delta-plot)(A) and (C) there is also a mean-variance relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation. Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures \@ref(fig:mean-variance-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seen with the pre-intervention scores above. As such, in this case, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.

```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes."}

load(here("plots/mean_variance_delta_plots"))

mean_variance_delta_plots

```

The $\textrm{ln}CVR$ can be used to overcome this issue though. Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before (see equation \@ref(eq:ml-model-eq)) using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons using $SD_{ir}$ or $\textrm{ln}VR$. The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[28]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[28]`%) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[88]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[88]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[88]`%) and further there is lower relative heterogeneity between studies in the effect estimates (figure \@ref(fig:forest-logCVR-plot)).

```{r forest-logCVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$CVR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}

# "Caterpillar plots of $\textrm{ln}CVR$ effect sizes for strength and hypertrophy outcomes. Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"

load(here("plots/forest_logCVR_strength"))
load(here("plots/forest_logCVR_hypertrophy"))
library(patchwork)
(forest_logCVR_strength / forest_logCVR_hypertrophy) +
  plot_annotation(tag_levels = "A")

```

There is, however, a potential limitation for the $\textrm{ln}CVR$ also that may need to be considered. Firstly, it is limited to the use of ratio scale data (which is not the case for the $\textrm{ln}\hat\sigma$ or $\textrm{ln}VR$); hence the need to transform the change scores to be positively signed in this specific case. Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean. Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet equivalent approach might be desirable. \textcolor{red}{Lastly, this statistic is limited to examination of pairwise comparisons of variance. In this example this poses no issue as we are comparing RT intervention group(s) to a control group. But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer an equivalent model to the $\textrm{ln}CVR$ one just applied.}

## Meta-regression of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$

Instead, we can use a version of the meta-regression model described above (see equation \@ref(eq:lnm-lns-model-eq) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$. In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$. \textcolor{red}{This model with just $Group$ as a predictor is equivalent to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups. This reflects the pairwise nature of the $\textrm{ln}CVR$.}

```{r, message=FALSE,warning=FALSE,echo=FALSE}
# load the summary dataframes to the environment
load(here("models/MultiLevelModel_ri__log_mean_mod_strength"))
load(here("models/MultiLevelModel_ri__log_mean_mod_hypertrophy"))

```

Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model to the dataset[^20]. The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{2[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_strength$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{2[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.ub[3],2)`]). See figure \@ref(fig:model-mean-variance-delta-plot).

[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm. The comparison of these models is included in the supplementary materials (strength - https://osf.io/n4wgk; hypertrophy - https://osf.io/hf8dx). In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy the addition of both random slopes improved model fit significantly. Though again we limit presentation in the main text to the simpler model as substantively conclusions were the same.

```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores."}

# "Meta-analytic scatter plot of the log mean and log standard deviation of change scores. Note, fitted lines are the regression slopes for control group (CON) and resistance training intervention (RT) groups each with their 95% confidence interval band."

load(here("plots/model_mean_variance_delta_plots"))

model_mean_variance_delta_plots

```

\textcolor{red}{Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework. Of course, the pairwise models described can be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering not only effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article}[^21]. \textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size. The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}

[^21]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderators as results for $\textrm{ln}CVR$ and the meta-regression model of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ were similar) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes. There were no obvious moderators of $\textrm{ln}CVR$ in particular.

# Discussion
Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis. Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^22] [@yangLowStatisticalPower2022].

[^22]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.

It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present. \textcolor{red}{We deliberately presented these varying approaches in order to highlight their assumptions to aid in readers understanding of their applications. Given the differences conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of more simple approaches are met or not before their application in specific situations. Where for example there is not an obvious mean-variance relationship it may be appropriate to utilise the $SD_{ir}$ or $\textrm{ln}VR$. However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate. Further, depending on the exact research questions it is worth considering balancing model complexity with its ability to provide an answer. If a simple pairwise comparison across a categorical variable is of interest then this can be explored with equivalent models using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$. But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.}

The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes the findings more relatable for researchers in sport and exercise sciences. Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$, \textcolor{red}{at least in this application,} gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation. In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive, yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others. Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021]. Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.

However, similar to the cross-sectional pre-intervention scores reported here, change scores demonstrated a mean-variance relationship in addition to heteroskedasticity. The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain. Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021]. Interventions such as these are valuable for the simplification of guidelines and recommendations. For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations [@steeleHigherEffortbasedParadigm2017; @steeleLongTermTimeCourseStrength2022].

The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis. Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021]. However, this potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome. In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.

# Conclusion
Embracing variability and focusing on more than merely the mean differences between groups or conditions, such as intervention and control comparisons, has the potential to inform experimental design and lead to changes in both the approach and direction of follow-up studies. Whether there is evidence of meaningful average intervention effects or not, where considerable variance effects are present it suggests that a meaningful line of research would be to aim at identifying subgroup- or participant-by-intervention interactions using appropriate study designs [@heckstedenIndividualResponseExercise2015]. Where variance effects are limited this instead suggests that translational work towards generalisable implementation might be the most meaningful line of future research. Finally, there may be cases where it is in fact desirable to identify interventions that actually reduce variance; for example, improvements in methodological approaches to enhance research [@usuiMetaanalysisVariationSuggests2021], or interventions to reduce variation in sport performances. Thus, researchers in sport and exercise science should consider exploring variance more systematically, and indeed utilise the meta-analytic framework to support this. This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawaMetaanalysisVariationEcological2015; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] presented here and echoing the efforts of other recent work [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].

# References

<div id="refs"></div>

### 2023-09-07 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - test.Rmd
Addressing round 1 JSS revisions

--- 2023-09-07 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "7th September 2023"
+date: "????"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
@@
 [^3]: \textcolor{red}{This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
-This is beyond the scope of this paper to discuss.
-We note however that the models we present all utilise Restricted Maximum Likelihood estimation.}
+This is beyond the scope of this paper to discuss.}
 
@@
 The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present.
-\textcolor{red}{In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.} The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
+The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
 Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), the $\textrm{ln}CVR$ is given by:
@@
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-\textcolor{red}{This model with just $Group$ as a predictor is equivalent to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
-This reflects the pairwise nature of the $\textrm{ln}CVR$.}
+This model with just $Group$ as a predictor is equivalent to the $\textrm{ln}CVR$ model in the previous section.
 
@@
 Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model to the dataset[^20].
-The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{2[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_strength$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{2[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.ub[3],2)`]).
+The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_strength$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.ub[3],2)`]).
 See figure \@ref(fig:model-mean-variance-delta-plot).
@@
 
-[^21]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderators as results for $\textrm{ln}CVR$ and the meta-regression model of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ were similar) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
+[^21] See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderators as results for $\textrm{ln}CVR$ and the meta-regression model of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ were similar) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
 There were no obvious moderators of $\textrm{ln}CVR$ in particular.

### 2023-09-07 paper/Revised Pre-print - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Addressing round 1 JSS revisions

--- 2023-09-07 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - test.Rmd
@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "????"
+author: 
+  - James Steele$^a$
+  - James Fisher$^a$
+  - Dave Smith$^b$
+  - Brad Schoenfeld$^c$
+  - Yefeng Yang$^d$
+  - Shinichi Nakagawa$^d$
+date: "7th September 2023"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
@@
 \\textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \\textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}."
+thanks: |
+  Preprint, please cite as: Steele, J., Fisher, J., Smith, D., Schoenfeld, B., Yang, R., & Nakagawa, S.
+(2022).
+Meta-Analysis of Variation in Sport and Exercise Science: Examples of Application Within Resistance Training Research.
+$\textrm{sportr}\chi\textrm{iv}$ DOI: https://doi.org/10.51224/SRXIV.214.
+  Address for correspondence: james.steele@solent.ac.uk
+  Affiliations:
+  $^a$Faculty of Sport, Health, and Social Sciences, Solent University, UK;
+  $^b$Research Centre for Musculoskeletal and Sports Medicine, Manchester Metropolitan University, Manchester, UK;
+  $^c$Health Sciences Department, CUNY Lehman College, Bronx, New York, USA;
+  $^d$Evolution & Ecology Research Centre and School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, New South Wales, Australia
 output: bookdown::pdf_document2
@@
    - \captionsetup[figure]{font=small}
-   - \usepackage{lineno}
-   - \linenumbers
 ---

### 2023-09-07 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -18,6 +18,7 @@ library(patchwork)
 library(europepmc)
 library(kableExtra)               
 library(webshot)
+library(ggtext)
 
 ##### Trend of Meta-Analysis in Resistance Training
 
@@ -171,12 +172,11 @@ I2_SMD_strength_lab <- data.frame(level = c("study", "arm", "es"),
 forest_SMD_strength <- Data_SMD_strength %>% 
   mutate(es = factor(es, levels = es[order(yi)]),
          se = sqrt(vi)) %>%
-  ggplot(aes(x=yi, y=es)) +
+  ggplot(aes(x=yi, y=fct_rev(es))) +
   geom_vline(xintercept = 0, linetype = "dashed") +
   geom_point(size=0.5, alpha=0.25) +
   scale_x_continuous(limits = c(-2,7.5), breaks = c(-1,0,1,2,3,4,5,6)) +
   geom_linerange(aes(xmin = yi - (se*1.96), xmax = yi + (se*1.96)), alpha=0.25) +
-  scale_y_discrete(limits=rev, expand = expansion(mult = c(0.075,0))) +
   geom_text(data = mutate_if(PI_SMD_strength,
                              is.numeric, round, 2),
             aes(label = glue::glue("Overall Estimate = {round(pred,2)} [95% Confidence Interval: {round(ci.lb,2)} to {round(ci.ub,2)}]"),
@@ -184,9 +184,12 @@ forest_SMD_strength <- Data_SMD_strength %>%
   geom_text(data = PI_SMD_strength,
             aes(label = glue::glue("[95% Prediction Interval: {round(pi.lb,2)} to {round(pi.ub,2)}]"),
                 x = 5, y = 90), hjust = "centre", size = 3) +
-  geom_text(data = I2_SMD_strength_lab,
-            aes(label = glue::glue("I^2 [Study = {round(study,1)}%; Arm = {round(arm,1)}%; Effect = {round(es,1)}%]"),
-                x = 5, y = 70), hjust = "centre", size = 3) +
+  geom_richtext(data = I2_SMD_strength_lab,
+            aes(label = glue::glue("*I*<sup>2</sup> [Study = {round(study,1)}%; Arm = {round(arm,1)}%; Effect = {round(es,1)}%]"),
+                x = 5, y = 70), size = 3,
+            fill = NA, label.color = NA, # remove background and outline
[…]

### 2023-09-12 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
Added edits based on the varying results for lnCVR and different MLMR models

@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "7th September 2023"
+date: "12th September 2023"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
 However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance.
-It is well known that \\textcolor{red}{individual responses to interventions vary considerably}.
+It is thought that \\textcolor{red}{individual responses to interventions vary considerably}.
 Hence, interest has increased in exploring *precision* or *personalised* exercise approaches.
@@
 \\textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \\textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}."
-output: bookdown::pdf_document2
+output: 
+  bookdown::pdf_document2:
+    extra_dependencies: ["flafter"]
 bibliography: references.bib
@@
 
-[^1]: \textcolor{red}{To clarify language here for those unfamiliar, the term and concept 'model' is used commonly in statistics.
+[^1]: \textcolor{red}{To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics.
 A statistical model essentially is specification of what we think the data generating process might be for a given situation.
@@
 
-[^2]: \textcolor{red}{'Effect size' is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
+[^2]: \textcolor{red}{*Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
 This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between} [@caldwellCaseDefaultEffect2020] \textcolor{red}{including, as we shall see, statistics describing variation.}
@@
 
-\textcolor{red}{Then the average of the distribution of effects can be calculated as per \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.}
+\textcolor{red}{Then the average of the distribution of effects can be calculated as per equation} \@ref(eq:fixed-average-eq) \textcolor{red}{substituting $w_{i}$ for $w^*_{i}$.}
 
-\textcolor{red}{Within sport and exercise science, although historically fixed effects models were more common} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays, likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above, the fixed-effect model is less commonly employed than the random-effects model} [@haggerMetaanalysis2022].
+\textcolor{red}{Although historically fixed effects models were commonplace in the field of sport and exercise science} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays the random effects model is more often employed} [@haggerMetaanalysis2022].
+\textcolor{red}{This is likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above.}
 
@@
 
-[^5]: \textcolor{red}{For example, 'strength' might be examined in different studies using different operationalisation including one repetition maximum testing or maximum voluntary contractions.
+[^5]: \textcolor{red}{For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions.
 Or the same operationalisations may be employed but different exercises such as the squat or bench press.}
@@
 In contrast, the $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), which only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
-Despite this, the use of effects sizes like the $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@debQuantifyingEffectsAcute2018; @nuzzoEccentricConcentricStrength2023] and to our knowledge only one meta-analysis of RT interventions has used this kind of effect size [@swintonInterpretingMagnitudeChange2022].
+Despite this, the use of effect sizes like the $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@debQuantifyingEffectsAcute2018; @nuzzoEccentricConcentricStrength2023] and to our knowledge only a couple of meta-analyses of RT interventions has used this kind of effect size [@swintonInterpretingMagnitudeChange2022; @wolfPartialVsFull2023].
 
@@
 Such small samples have even lower statistical power to detect differences in variation as compared to means [@yangLowStatisticalPower2022].
+
+\textcolor{red}{Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
+We hope this will make clear why we offer such recommendations in our discussion and conclusion.
+As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.}
 
@@
 
-\textcolor{red}{Although ultimately we will recommend certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
-We hope this will make clear why we offer certain recommendations in our discussion As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, and also to provide an explanation of them in plain English.}
-
 [^13]: \textcolor{red}{Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
@@
   pack_rows("Sample Type", 8, 9, bold = FALSE) %>%
-  pack_rows("RT + Adjuvant Intervention?", 10, 11, bold = FALSE) %>%
+  pack_rows("RT Intervention Only?", 10, 11, bold = FALSE) %>%
   pack_rows("Task Failure?", 18, 19, bold = FALSE) %>%
@@
 [-@politoModeratorsStrengthGains2021] originally used a normal random-effects meta-analysis as described in the introduction.
-However, the data we extracted were hierarchical in nature\textcolor{red}{; as opposed to there being only the assumption that studies are a random effect, due to their being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
+However, the data we extracted were hierarchical in nature.
+\textcolor{red}{Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
 \textcolor{red}{This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
-We can describe the overall model then as:}
+We can therefore describe the overall model as:}
 
@@
 \begin{equation}
-\hat\theta_{ijk}=(\theta+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})++m_{ijk}
+\hat\theta_{ijk}=(\theta+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+m_{ijk}
 (\#eq:ml-model-eq)
@@
 
-[^17]: \textcolor{red}{Technically then the 'random' effects model presented earlier is also a 'mixed' effects model. It is traditionally referred to as the random-effects model though.}
+[^17]: \textcolor{red}{Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.}
 
-We then fitted the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
+We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
 A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produced greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).
@@
 As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%).
-Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}
+Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure }\@ref(fig:forest-SMD-plot)) \textcolor{red}{were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}
 
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic.
-First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ \textcolor{red}{and increase its flexibility to accomodate wider applications}.
+First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures \textcolor{red}{to emphasise the need to consider this assumption,} and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ \textcolor{red}{and provide flexibility to accomodate wider applications}.
 
@@
 As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
-\textcolor{red}{The variance in standard deviations increase with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
+\textcolor{red}{The variance in standard deviations increases with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
 
@@
 Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from multiple outcomes within multiple arms within studies we can also estimate this relationship using a multilevel mixed-effects meta-regression model similar to that applied above.
-In this case though we are including an addition *predictor* variable, the $\textrm{ln}\overline{x}$.
-For example, the following model specifies $\textrm{ln}\overline{x}$ as a fixed effect with random intercepts for study and arm:
+In this case though we are including an additional *predictor* variable, the $\textrm{ln}\overline{x}$.
+For example, the following model specifies $\textrm{ln}\overline{x}$ as a fixed effect with random intercepts for study, arm, and effect:
 
@@
 
-\textcolor{red}{These kinds of models are incredibly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
+\textcolor{red}{These kinds of models are highly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
 
-[^18]: In contrast to the models examining effect sizes relating to comparisons between and intervention group and control group, in the models examining $\textrm{ln}\sigma_{ijk}$ and $\textrm{ln}\overline{x}_{ijk}$ as a predictor the term *arm* refers to boh the intervention groups(s) and control group.
+[^18]: In contrast to the models presented examining effect sizes relating to comparisons between and intervention group and control group, in the models examining $\textrm{ln}\sigma_{ijk}$ and $\textrm{ln}\overline{x}_{ijk}$ as a predictor the term *arm* refers to both the intervention groups(s) and control group.
 Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2, and control).
@@
 Returning to the context of interindividual response variation to interventions, the presence of a mean-variance relationship in the data would imply that we cannot rely on absolute comparisons of variance (i.e., $SD_{ir}$ or $\textrm{ln}VR$) to determine whether interindividual response variation is actually present.
-So we should also explore this for the change-scores in the RT and control groups and determining the appropriate effect sizes to explore.
+So we should also explore this assumption for the change-scores in the RT and control groups and determining the appropriate effect sizes to explore.
 
@@
 Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
-Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet equivalent approach might be desirable.
+Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet \textcolor{red}{comparable} approach might be desirable.
 \textcolor{red}{Lastly, this statistic is limited to examination of pairwise comparisons of variance.
-In this example this poses no issue as we are comparing RT intervention group(s) to a control group.
-But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer an equivalent model to the $\textrm{ln}CVR$ one just applied.}
+In this example this approach poses no issue as we are comparing RT intervention group(s) to a control group.
+But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a \textcolor{red}{comparable} effect of interest to the $\textrm{ln}CVR$ one just applied.}
 
@@
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-\textcolor{red}{This model with just $Group$ as a predictor is equivalent to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
-This reflects the pairwise nature of the $\textrm{ln}CVR$.}
-
-```{r, message=FALSE,warning=FALSE,echo=FALSE}
-# load the summary dataframes to the environment
-load(here("models/MultiLevelModel_ri__log_mean_mod_strength"))
-load(here("models/MultiLevelModel_ri__log_mean_mod_hypertrophy"))
-
-```
-
-Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model to the dataset[^20].
-The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{2[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_strength$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{2[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.ub[3],2)`]).
-See figure \@ref(fig:model-mean-variance-delta-plot).
+\textcolor{red}{This model with just $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
+This reflects the pairwise nature of the $\textrm{ln}CVR$.} Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other extensions[^20] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot) for the best fitting model plots).
 
@@
 The comparison of these models is included in the supplementary materials (strength - https://osf.io/n4wgk; hypertrophy - https://osf.io/hf8dx).
-In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy the addition of both random slopes improved model fit significantly.
-Though again we limit presentation in the main text to the simpler model as substantively conclusions were the same.
+In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy intercept only model fit best.
 
@@
 
+\textcolor{red}{Despite the models comparability in terms of what they are trying to estimate (i.e., the difference in variance in change scores between the RT and CON groups) we arrive at substantively different conclusions compared to the $\textrm{ln}CVR$ model and dependent upon the exact specification (see footnote 20) of the meta-regression model.
+Table} \@ref(tab:model-estimates-tab) \textcolor{red}{shows the different estimates from each model for this difference.
+Where the $\textrm{ln}CVR$ suggested that the introduction of an RT intervention actually reduced variance in change scores (i.e., the estimate was negative) we see the opposite conclusion (i.e., the estimate is positive) for all multilevel meta-regression models of strength, and for hypertrophy this appears to be dependent on whether the coefficients (i.e., slopes) for the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ are allowed to vary between studies and arms.}
+
+```{r model-estimates-tab, message=FALSE,warning=FALSE,echo=FALSE}
+# load the summary dataframes to the environment
+load(here("models/group_model_estimates"))
+
+knitr::kable(
+  model_estimates, 
+  col.names = c("Outcome", "Model", "Estimate", "Lower 95% CI", "Upper 95% CI"),
+  caption = "Comparison of estimates from model using $\\textrm{ln}CVR$ and multilevel meta-regression models of $\\textrm{ln}\\hat\\sigma$ and $\\textrm{ln}\\overline{x}$ with group (RT vs CON) as a predictor",
+) %>%
+  row_spec(0, bold = TRUE) %>%
+  pack_rows(index = c("Strength" = 4, "Hypertrophy" = 4)) %>%
+  footnote(general = c("CI = confidence interval", "MLMR = multilevel meta-regression")) %>%
+  kable_classic(full_width = FALSE) %>%
+  kable_styling()
+
+```
+
+\textcolor{red}{It is worth explaining why this might be the case to aid readers in considering the assumptions that different models make.
+The $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; In the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
+In both however this is a strong assumption.
+The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in the random slope models this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
+Evidently, the assumptions that are made about the data-generating processes can impact upon the modelling strategy chosen}[^21].
+\textcolor{red}{Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression model allows).
+It is not necessarily clear which model should be preferred here, but thought should be given to the assumptions each makes and the fit of each model to the data.}
+
+[^21]: \textcolor{red}{It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
+Given that for most effects that might be included in such models we can determine the sampling variance for $\overline{x}$ one approach to address this might be to employ models that incorporate the variance on this predictor (i.e., measurement error or errors in variables models).
+This is beyond the scope of this paper to discuss.}
+
+## Moderators of variance effects
 \textcolor{red}{Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework.
-Of course, the pairwise models described can be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering not only effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article}[^21].
+Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)}[^22].
 \textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size.
@@
 
-[^21]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderators as results for $\textrm{ln}CVR$ and the meta-regression model of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ were similar) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
+[^22]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderator analysis results) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
 There were no obvious moderators of $\textrm{ln}CVR$ in particular.
@@
 Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
-Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^22] [@yangLowStatisticalPower2022].
+Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^23] [@yangLowStatisticalPower2022].
 
-[^22]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.
+[^23]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.
 
 It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
-\textcolor{red}{We deliberately presented these varying approaches in order to highlight their assumptions to aid in readers understanding of their applications.
-Given the differences conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of more simple approaches are met or not before their application in specific situations.
+\textcolor{red}{We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
+Given the different conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of simpler approaches are met, or indeed seem reasonable, before their application in specific situations.
 Where for example there is not an obvious mean-variance relationship it may be appropriate to utilise the $SD_{ir}$ or $\textrm{ln}VR$.
-However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate.
+However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate, though assumptions regarding whether this relationship is proportional and fixed should also be considered.
+The structure of the data also impacts the specific modelling approach to be employed and whether or not assumptions that the mean-variance relationship varies within study or arm levels can be incorporated.
 Further, depending on the exact research questions it is worth considering balancing model complexity with its ability to provide an answer.
-If a simple pairwise comparison across a categorical variable is of interest then this can be explored with equivalent models using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$.
-But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.}
+If a simple pairwise comparison across a categorical variable is of interest then this can be explored with comparable models using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$.
+But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.
+Lastly, given the possible sensitivity of substantive conclusions to the exact model specifications used it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
+With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions} [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
 
-The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes the findings more relatable for researchers in sport and exercise sciences.
-Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$, \textcolor{red}{at least in this application,} gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation.
-In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive, yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
-Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
+Whether or not there is clear evidence for average intervention effects, research aimed at exploring variation is important.
+Even in the absence of an average intervention effect it may be that a large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
 Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.
+Interventions, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
+Interventions such as these are valuable for the simplification of guidelines and recommendations.
+This also potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome.
+In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
 
-However, similar to the cross-sectional pre-intervention scores reported here, change scores demonstrated a mean-variance relationship in addition to heteroskedasticity.
-The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain.
-Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
-Interventions such as these are valuable for the simplification of guidelines and recommendations.
-For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations [@steeleHigherEffortbasedParadigm2017; @steeleLongTermTimeCourseStrength2022].
-
-The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis.
-Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021].
-However, this potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome.
-In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
+The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
+\textcolor{red}{Given the varied results, this dataset does not necessarily provide clear answers as to the question of whether true interindividual response variation is present and so this remains an open area of research.
+However, the examples provided highlight the importance of considering carefully the assumptions inherent in the modelling approaches chosen to explore such a substantive topic.
+We hope they also prompt researchers to further explore variation.}
 
@@
 This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawaMetaanalysisVariationEcological2015; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] presented here and echoing the efforts of other recent work [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
+\textcolor{red}{Researchers should be aware though that the meta-analytic modelling of variance, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make.
+Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.}
 

### 2023-09-12 paper/Revised Pre-print - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Added edits based on the varying results for lnCVR and different MLMR models

--- 2023-09-07 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - test.Rmd
@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "????"
+author: 
+  - James Steele$^a$
+  - James Fisher$^a$
+  - Dave Smith$^b$
+  - Brad Schoenfeld$^c$
+  - Yefeng Yang$^d$
+  - Shinichi Nakagawa$^d$
+date: "12th September 2023"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
 However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance.
-It is well known that \\textcolor{red}{individual responses to interventions vary considerably}.
+It is thought that individual responses to interventions vary considerably.
 Hence, interest has increased in exploring *precision* or *personalised* exercise approaches.
@@
 Yet, larger samples sizes than those used for typical mean effects are required when probing variances.
-Thus, \\textcolor{red}{in} a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
-\\textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \\textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}."
-output: bookdown::pdf_document2
+Thus, in a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
+Despite the value of embracing and exploring variation alongside mean effects in sport and exercise science it is rarely applied to research synthesis through meta-analysis.We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs."
+thanks: |
+  Preprint, please cite as: Steele, J., Fisher, J., Smith, D., Schoenfeld, B., Yang, R., & Nakagawa, S.
+(2022).
+Meta-Analysis of Variation in Sport and Exercise Science: Examples of Application Within Resistance Training Research.
+$\textrm{sportr}\chi\textrm{iv}$ DOI: https://doi.org/10.51224/SRXIV.214.
+  Address for correspondence: james.steele@solent.ac.uk
+  Affiliations:
+  $^a$Faculty of Sport, Health, and Social Sciences, Solent University, UK;
+  $^b$Research Centre for Musculoskeletal and Sports Medicine, Manchester Metropolitan University, Manchester, UK;
+  $^c$Health Sciences Department, CUNY Lehman College, Bronx, New York, USA;
+  $^d$Evolution & Ecology Research Centre and School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, New South Wales, Australia
+output: 
+  bookdown::pdf_document2:
+    extra_dependencies: ["flafter"]
 bibliography: references.bib
@@
 Accordingly, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
-\textcolor{red}{However, to begin we provide a conceptual overview of meta-analyses and effect sizes as typically used within sport and exercise science.}
+However, to begin we provide a conceptual overview of meta-analyses and effect sizes as typically used within sport and exercise science.
 
@@
 
-\textcolor{red}{There are two popular models}[^1] \textcolor{red}{used for meta-analysis: the fixed-effect model and the random-effects model} [@borensteinBasicIntroductionFixedeffect2010].
-\textcolor{red}{The fixed effect model assumes that there is one true effect size}[^2] \textcolor{red}{that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error.
+There are two popular models[^1] used for meta-analysis: the fixed-effect model and the random-effects model [@borensteinBasicIntroductionFixedeffect2010].
+The fixed effect model assumes that there is one true effect size[^2] that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error.
 This essentially means that there is a single common effect which is fixed across studies and each study takes samples of individuals from the population to estimate this effect.
-We can express this model in the following formula:}
+We can express this model in the following formula:
 
-[^1]: \textcolor{red}{To clarify language here for those unfamiliar, the term and concept 'model' is used commonly in statistics.
+[^1]: To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics.
 A statistical model essentially is specification of what we think the data generating process might be for a given situation.
 In the context of meta-analyses the data are usually the individual effects that we have extracted from studies i.e., the results of each study.
-The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.}
+The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.
 
-[^2]: \textcolor{red}{'Effect size' is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
-This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between} [@caldwellCaseDefaultEffect2020] \textcolor{red}{including, as we shall see, statistics describing variation.}
+[^2]: *Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
+This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between [@caldwellCaseDefaultEffect2020] including, as we shall see, statistics describing variation.
 
@@
 
-\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
+where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
 Studies with smaller standard errors of their effect estimates have smaller sampling errors and so these higher precision estimates are given greater weight in the model.
-The weighting given to each study is calculated as:}
+The weighting given to each study is calculated as:
 
@@
 
-\textcolor{red}{Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
+Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
 It is referred to as inverse-variance weighting.
-Then, the overall weighted mean effect estimate from the model is then calculated as:}
+Then, the overall weighted mean effect estimate from the model is then calculated as:
 
@@
 
-\textcolor{red}{Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
+Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
 Instead it allows for the true effects that each study estimates to differ.
@@
 Hence, in the fixed-effect model there is one effect and it is assumed to be *fixed* across studies, whereas in the random-effects model there are many and we examine an assumed *random* sample of them.
-We can express this model in the following formula:}
+We can express this model in the following formula:
 
@@
 
-\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
+where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
 Essentially this model assumes that each individual study estimates an effect and there will be some sampling error in estimating it.
 But, the effects each study estimates comes from an overall distribution of true effects with a mean value (i.e., $\theta_{i}$).
-The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated}[^3].
-\textcolor{red}{This essentially describes the heterogeneity between the effects included.
-Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:}
+The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated[^3].
+This essentially describes the heterogeneity between the effects included.
+Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:
 
-[^3]: \textcolor{red}{This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
-This is beyond the scope of this paper to discuss.}
+[^3]: This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
+This is beyond the scope of this paper to discuss.
+We note however that the models we present all utilise Restricted Maximum Likelihood estimation.
 
@@
 
-\textcolor{red}{Then the average of the distribution of effects can be calculated as per \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.}
+Then the average of the distribution of effects can be calculated as per equation \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.
 
-\textcolor{red}{Within sport and exercise science, although historically fixed effects models were more common} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays, likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above, the fixed-effect model is less commonly employed than the random-effects model} [@haggerMetaanalysis2022].
+Although historically fixed effects models were commonplace in the field of sport and exercise science [@haggerMetaanalysisSportExercise2006] nowadays the random effects model is more often employed [@haggerMetaanalysis2022].
+This is likely due to the fact that direct (i.e., exact) replication of studies is rare[^4] and instead studies often explore similar effects across varying moderating factors such as those noted above.
 
-[^4]: \textcolor{red}{Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).}
+[^4]: Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).
 
-As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] \textcolor{red}{likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable.
-The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping}; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
+As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable.
+The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
 Indeed, a recent umbrella review [@bernardez-vazquezResistanceTrainingVariables2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes, all of which focused on the comparison of mean changes between different intervention groups.
 
-\textcolor{red}{Often, due to the varying operationalisations used for broad outcome concepts}[^5]\textcolor{red}{, an effect size is used which is "standardised" across studies to allow for their synthesis.} Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
+Often, due to the varying operationalisations used for broad outcome concepts[^5], an effect size is used which is "standardised" across studies to allow for their synthesis.
+Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
 This statistic is usually a version of Cohen's *d* [@cohenStatisticalPowerAnalysis1988], or its bias-corrected[^7] metric referred to as Hedges' *g* [@hedgesStatisticalMethodsMetaAnalysis2014; @borensteinIntroductionMetaAnalysis2021; @nakagawaEffectSizeConfidence2007][^8].
@@
 
-[^5]: \textcolor{red}{For example, 'strength' might be examined in different studies using different operationalisation including one repetition maximum testing or maximum voluntary contractions.
-Or the same operationalisations may be employed but different exercises such as the squat or bench press.}
+[^5]: For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions.
+Or the same operationalisations may be employed but different exercises such as the squat or bench press.
 
@@
 
-[^7]: \textcolor{red}{For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct.
-Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.}
+[^7]: For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct.
+Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.
 
 [^8]: We will refer to both merely as the SMD throughout the manuscript for simplicity and note that throughout when reporting a 'SMD' we are reporting the bias-corrected version.
-\textcolor{red}{We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance;} @caldwellCaseDefaultEffect2020).
+We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance; @caldwellCaseDefaultEffect2020).
 
@@
 For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawaMeanStrikesBack2012].
-Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] \textcolor{red}{and in the number of repetitions that can be performed at different percentages of one repetition maximum} [@nuzzoMaximalNumberRepetitions2023].
+Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] and in the number of repetitions that can be performed at different percentages of one repetition maximum [@nuzzoMaximalNumberRepetitions2023].
 
@@
 Such small samples have even lower statistical power to detect differences in variation as compared to means [@yangLowStatisticalPower2022].
+
+Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
+We hope this will make clear why we offer such recommendations in our discussion and conclusion.
+As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.
 
@@
 The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present.
+In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.
 The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
@@
 All analysis examples were performed in R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauerConductingMetaAnalysesMetafor2010].
-The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/) \textcolor{red}{or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git)}.
+The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/) or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git).
 
@@
 
-\textcolor{red}{Although ultimately we will recommend certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
-We hope this will make clear why we offer certain recommendations in our discussion As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, and also to provide an explanation of them in plain English.}
-
-[^13]: \textcolor{red}{Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
+[^13]: Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
 This can be extended to meta-analytic synthesis where the independent variables reflect characteristics associated with the effects included.
-For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.}
+For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.
 
@@
 
-## Detecting the presence of interindividual response variation to resistance training intervention \textcolor{red}{with absolute variance statistics}
+## Detecting the presence of interindividual response variation to resistance training intervention with absolute variance statistics
 First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^14] meta-analysis to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^15].
@@
 [-@politoModeratorsStrengthGains2021] originally used a normal random-effects meta-analysis as described in the introduction.
-However, the data we extracted were hierarchical in nature\textcolor{red}{; as opposed to there being only the assumption that studies are a random effect, due to their being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
-\textcolor{red}{This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
-We can describe the overall model then as:}
+However, the data we extracted were hierarchical in nature.
+Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.
+Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
+This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
+We can therefore describe the overall model as:
 
@@
 
-\textcolor{red}{where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
-This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects}[^17].
-\textcolor{red}{The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).}
+where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
+This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects[^17].
+The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).
 
-[^17]: \textcolor{red}{Technically then the 'random' effects model presented earlier is also a 'mixed' effects model. It is traditionally referred to as the random-effects model though.}
+[^17]: Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.
 
-We then fitted the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
+We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
 A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produced greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).
@@
 As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%).
-Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}
+Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).
 
@@
 This additional variance might support previous perspectives [@carpinelliInterindividualHeterogeneityAdaptations2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* independent of the random error).
-However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean \textcolor{red}{(i.e., that the variance is similar whether mean values are low or high)}.
+However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean (i.e., that the variance is similar whether mean values are low or high).
 As we have seen from the SMD and $\textrm{ln}RR$ models, RT interventions increase mean scores.
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic.
-First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ \textcolor{red}{and increase its flexibility to accomodate wider applications}.
+First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures to emphasise the need to consider this assumption, and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ and provide flexibility to accomodate wider applications.
 
@@
 As such, we can model variance of each effect as the response itself with the mean of the effect as the predictor.
-The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models \textcolor{red}{(i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).} Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
+The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models (i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).
+Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
 In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al.
@@
 As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
-\textcolor{red}{The variance in standard deviations increase with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
+The variance in standard deviations increases with higher mean values.
+This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
 
@@
 When this relationship holds, under most circumstances the standard deviation is not proportional to the mean.
-However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear \textcolor{red}{based upon the product and power logarithmic rules:}
+However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear based upon the product and power logarithmic rules:
 
@@
 
-\textcolor{red}{These kinds of models are incredibly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
+These kinds of models are highly flexible.
+Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
 
-[^18]: In contrast to the models examining effect sizes relating to comparisons between and intervention group and control group, in the models examining $\textrm{ln}\sigma_{ijk}$ and $\textrm{ln}\overline{x}_{ijk}$ as a predictor the term *arm* refers to boh the intervention groups(s) and control group.
+[^18]: In contrast to the models presented examining effect sizes relating to comparisons between and intervention group and control group, in the models examining $\textrm{ln}\sigma_{ijk}$ and $\textrm{ln}\overline{x}_{ijk}$ as a predictor the term *arm* refers to both the intervention groups(s) and control group.
 Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2, and control).
@@
 [^19]: We do not have to limit ourselves to only fixed effect predictor terms as we have here.
-Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes)}, and their correlations, to enhance generalisability of inferences [@barrRandomEffectsStructure2013].
+Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes (i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences [@barrRandomEffectsStructure2013].
 We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the overall average slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
@@
 Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
-Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet equivalent approach might be desirable.
-\textcolor{red}{Lastly, this statistic is limited to examination of pairwise comparisons of variance.
-In this example this poses no issue as we are comparing RT intervention group(s) to a control group.
-But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer an equivalent model to the $\textrm{ln}CVR$ one just applied.}
+Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet comparable approach might be desirable.
+Lastly, this statistic is limited to examination of pairwise comparisons of variance.
+In this example this approach poses no issue as we are comparing RT intervention group(s) to a control group.
+But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a comparable effect of interest to the $\textrm{ln}CVR$ one just applied.
 
@@
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-This model with just $Group$ as a predictor is equivalent to the $\textrm{ln}CVR$ model in the previous section.
-
-```{r, message=FALSE,warning=FALSE,echo=FALSE}
-# load the summary dataframes to the environment
-load(here("models/MultiLevelModel_ri__log_mean_mod_strength"))
-load(here("models/MultiLevelModel_ri__log_mean_mod_hypertrophy"))
-
-```
-
-Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model to the dataset[^20].
-The results were largely similar, albeit slightly attenuated, to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_strength$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(MultiLevelModel_ri__log_mean_mod_hypertrophy$ci.ub[3],2)`]).
-See figure \@ref(fig:model-mean-variance-delta-plot).
+This model with just $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
+This reflects the pairwise nature of the $\textrm{ln}CVR$.
+Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other extensions[^20] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot) for the best fitting model plots).
 
@@
 
-\textcolor{red}{Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework.
-Of course, the pairwise models described can be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering not only effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article}[^21].
-\textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size.
-The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}
+Despite the models comparability in terms of what they are trying to estimate (i.e., the difference in variance in change scores between the RT and CON groups) we arrive at substantively different conclusions compared to the $\textrm{ln}CVR$ model and dependent upon the exact specification (see footnote 20) of the meta-regression model.
+Table \@ref(tab:model-estimates-tab) shows the different estimates from each model for this difference.
+Where the $\textrm{ln}CVR$ suggested that the introduction of an RT intervention actually reduced variance in change scores (i.e., the estimate was negative) we see the opposite conclusion (i.e., the estimate is positive) for all multilevel meta-regression models of strength, and for hypertrophy this appears to be dependent on whether the coefficients (i.e., slopes) for the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ are allowed to vary between studies and arms.
 
-[^21] See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderators as results for $\textrm{ln}CVR$ and the meta-regression model of $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ were similar) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
+```{r model-estimates-tab, message=FALSE,warning=FALSE,echo=FALSE}
+# load the summary dataframes to the environment
+load(here("models/group_model_estimates"))
+
+knitr::kable(
+  model_estimates, 
+  col.names = c("Outcome", "Model", "Estimate", "Lower 95% CI", "Upper 95% CI"),
+  caption = "Comparison of estimates from model using $\\textrm{ln}CVR$ and multilevel meta-regression models of $\\textrm{ln}\\hat\\sigma$ and $\\textrm{ln}\\overline{x}$ with group (RT vs CON) as a predictor",
+) %>%
+  row_spec(0, bold = TRUE) %>%
+  pack_rows(index = c("Strength" = 4, "Hypertrophy" = 4)) %>%
+  footnote(general = c("CI = confidence interval", "MLMR = multilevel meta-regression")) %>%
+  kable_classic(full_width = FALSE) %>%
+  kable_styling()
+
+```
+
+It is worth explaining why this might be the case to aid readers in considering the assumptions that different models make.
+The $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; In the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
+In both however this is a strong assumption.
+The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in the random slope models this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
+Evidently, the assumptions that are made about the data-generating processes can impact upon the modelling strategy chosen[^21].
+Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression model allows).
+It is not necessarily clear which model should be preferred here, but thought should be given to the assumptions each makes and the fit of each model to the data.
+
+[^21]: It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
+Given that for most effects that might be included in such models we can determine the sampling variance for $\overline{x}$ one approach to address this might be to employ models that incorporate the variance on this predictor (i.e., measurement error or errors in variables models).
+This is beyond the scope of this paper to discuss.
+
+## Moderators of variance effects
+Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework.
+Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)[^22].
+But these are limited to the pairwise comparisons of a categorical variable as the effect size.
+The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories including any number of arbitrary predictors, fixed and random and assumptions about their correlations.
+
+[^22]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderator analysis results) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
 There were no obvious moderators of $\textrm{ln}CVR$ in particular.
@@
 Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
-Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^22] [@yangLowStatisticalPower2022].
+Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^23] [@yangLowStatisticalPower2022].
 
-[^22]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.
+[^23]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.
 
 It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
-\textcolor{red}{We deliberately presented these varying approaches in order to highlight their assumptions to aid in readers understanding of their applications.
-Given the differences conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of more simple approaches are met or not before their application in specific situations.
+We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
+Given the different conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of simpler approaches are met, or indeed seem reasonable, before their application in specific situations.
 Where for example there is not an obvious mean-variance relationship it may be appropriate to utilise the $SD_{ir}$ or $\textrm{ln}VR$.
-However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate.
+However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate, though assumptions regarding whether this relationship is proportional and fixed should also be considered.
+The structure of the data also impacts the specific modelling approach to be employed and whether or not assumptions that the mean-variance relationship varies within study or arm levels can be incorporated.
 Further, depending on the exact research questions it is worth considering balancing model complexity with its ability to provide an answer.
-If a simple pairwise comparison across a categorical variable is of interest then this can be explored with equivalent models using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$.
-But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.}
+If a simple pairwise comparison across a categorical variable is of interest then this can be explored with comparable models using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$.
+But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.
+Lastly, given the possible sensitivity of substantive conclusions to the exact model specifications used it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
+With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
 
-The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes the findings more relatable for researchers in sport and exercise sciences.
-Using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$, \textcolor{red}{at least in this application,} gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation.
-In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive, yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
-Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
+Whether or not there is clear evidence for average intervention effects, research aimed at exploring variation is important.
+Even in the absence of an average intervention effect it may be that a large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
 Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.
+Interventions, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
+Interventions such as these are valuable for the simplification of guidelines and recommendations.
+This also potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome.
+In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
 
-However, similar to the cross-sectional pre-intervention scores reported here, change scores demonstrated a mean-variance relationship in addition to heteroskedasticity.
-The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain.
-Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
-Interventions such as these are valuable for the simplification of guidelines and recommendations.
-For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations [@steeleHigherEffortbasedParadigm2017; @steeleLongTermTimeCourseStrength2022].
-
-The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis.
-Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021].
-However, this potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome.
-In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
+The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
+Given the varied results, this dataset does not necessarily provide clear answers as to the question of whether true interindividual response variation is present and so this remains an open area of research.
+However, the examples provided highlight the importance of considering carefully the assumptions inherent in the modelling approaches chosen to explore such a substantive topic.
+We hope they also prompt researchers to further explore variation.
 
@@
 This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawaMetaanalysisVariationEcological2015; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] presented here and echoing the efforts of other recent work [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
+Researchers should be aware though that the meta-analytic modelling of variance, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make.
+Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.
 

### 2023-09-12 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -1445,7 +1445,7 @@ webshot("models/compare_group_models_strength.html", "models/compare_group_model
 ### Meta-analytic scatter plot
 
 # get the predicted log values
-Data_long_strength_log <- cbind(Data_long_strength, predict(RobuEstMultiLevelModel_ri__log_mean_mod_strength )) %>%
+Data_long_strength_log <- cbind(Data_long_strength, predict(RobuEstMultiLevelModel_rs_study__log_mean_mod_strength )) %>%
     mutate(wi = 1/sqrt(SD_log_vi),
            size = 0.5 + 3.0 * (wi - min(wi))/(max(wi) - min(wi))) 
 
@@ -1595,6 +1595,51 @@ save(model_mean_variance_delta_plots, file = "plots/model_mean_variance_delta_pl
 model_mean_variance_delta_plots
 
 ggsave("plots/model_mean_variance_delta_plots.tiff", width = 10, height = 5, device = "tiff", dpi = 300)
+
+
+### Compare estimates from log CVR and each meta-regression model for strength and hypertrophy
+model_estimates <- data.frame(Outcome = "",
+                 Model = c("lnCVR", "MLMR: Random Intercept", "MLMR: + Random Slope (Study)", "MLMR: + Random Slope (Study & Arm)",
+                           "lnCVR", "MLMR: Random Intercept", "MLMR: + Random Slope (Study)", "MLMR: + Random Slope (Study & Arm)"),
+                 Estimate = c(RobuEstMultiLevelModel_logCVR_strength$b, 
+                              RobuEstMultiLevelModel_ri__log_mean_mod_strength$b[3], 
+                              RobuEstMultiLevelModel_rs_study__log_mean_mod_strength$b[3], 
+                              RobuEstMultiLevelModel_rs_both__log_mean_mod_strength$b[3],
+                              RobuEstMultiLevelModel_logCVR_hypertrophy$b, 
+                              RobuEstMultiLevelModel_ri__log_mean_mod_hypertrophy$b[3], 
+                              RobuEstMultiLevelModel_rs_study__log_mean_mod_hypertrophy$b[3],
+                              RobuEstMultiLevelModel_rs_both__log_mean_mod_hypertrophy$b[3]),
+                 `Lower 95% CI` = c(RobuEstMultiLevelModel_logCVR_strength$ci.lb, 
+                           RobuEstMultiLevelModel_ri__log_mean_mod_strength$ci.lb[3], 
[…]

### 2023-09-14 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R
Fixed error in es categorical coding after reshaping to long, refit all relevant models

@@ -14,6 +14,7 @@
 library(metafor)
 library(tidyverse)
 library(orchaRd)
+library(bayestestR)
 library(patchwork)
 library(europepmc)
 library(kableExtra)               
@@ -868,17 +869,21 @@ Data_long_pre <- cbind(Data_long_pre_m, sd = Data_long_pre_sd$sd)
 Data_long_pre_RT <- Data_long_pre %>%
   filter(group == "RT") %>%
   mutate(n = RT_n,
-         arm = as.factor(unclass(factor(unlist(arm))))) %>%
+         arm = as.factor(unclass(factor(unlist(arm)))),
+         es = as.factor(unclass(factor(unlist(es))))) %>%
   select(study, arm, es, outcome, group, mean, sd, n)
 
 Data_long_pre_CON <- Data_long_pre %>%
   filter(group == "CON") %>%
-  distinct() %>%
+  distinct(study, outcome, group, mean, sd, .keep_all = TRUE) %>%
   mutate(n = CON_n,
-         arm = as.factor(unclass(factor(unlist(arm)))+length(unique(Data_long_pre_RT$arm)))) %>%
+         arm = as.factor(unclass(factor(unlist(arm)))+length(unique(Data_long_pre_RT$arm))),
+         es = as.factor(unclass(factor(unlist(es)))+length(unique(Data_long_pre_RT$es)))) %>%
   select(study, arm, es, outcome, group, mean, sd, n)
 
-Data_long_pre <- rbind(Data_long_pre_RT, Data_long_pre_CON)
+Data_long_pre <- rbind(Data_long_pre_RT, Data_long_pre_CON) %>%
+  filter(!is.na(mean) &
+           !is.na(sd))
 
 # Calculate log SD and variance of log SD
 Data_long_pre$SD_log <- log(Data_long_pre$sd) + (1/(2*(Data_long_pre$n-1)))
@@ -936,97 +941,211 @@ ggsave("plots/mean_variance_pre_plots.tiff", width = 10, height = 7.5, device = 
 
 ### Comparing a range of models
 
-### Fitting a random intercept only mixed effects model for mean-variance
-# We will include the outcome type as a fixed moderator with random intercepts for both study and arm 
+# Random intercepts only
 
-MultiLevelModel_ri__log_mean_variance <- rma.mv(SD_log, V=SD_log_vi, data=Data_long_pre,
+MultiLevelModel_ri_only_log_mean_variance <- rma.mv(SD_log, V=SD_log_vi, data=Data_long_pre,
[…]

### 2023-09-14 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
Re-revise manuscript, update pre-print

@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "12th September 2023"
+date: "14th September 2023"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
@@
 
-\textcolor{red}{Although historically fixed effects models were commonplace in the field of sport and exercise science} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays the random effects model is more often employed} [@haggerMetaanalysis2022].
+\textcolor{red}{Although historically fixed effect models were commonplace in the field of sport and exercise science} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays the random effects model is more often employed} [@haggerMetaanalysis2022].
 \textcolor{red}{This is likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above.}
@@
 [^19]: We do not have to limit ourselves to only fixed effect predictor terms as we have here.
-Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes)}, and their correlations, to enhance generalisability of inferences [@barrRandomEffectsStructure2013].
+Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences} [@barrRandomEffectsStructure2013].
 We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the overall average slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
-Indeed, we fit additional models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm.
-The comparison of these models is included in the supplementary materials (https://osf.io/4xrcg).
-Note, the addition of both random slopes for study, and for arm, improved model fit significantly, though we limit presentation in the main text to the simpler model.
+\textcolor{red}{These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms.
+We could also do the same and include random slopes for $(\beta_{1}$ on $\textrm{ln}\overline{x}$ thus allowing for the strength of the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ to also vary between studies and arms.
+Indeed, we fit a range of models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Outcome$ by study, (5) the inclusion of correlated random slopes for $Outcome$ by study and arm, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study, (7) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and arm.
+The comparison of these models using $2\times\textrm{log}BF$} [@kassBayesFactors1995] \textcolor{red}{from approximate Bayesian information criterion} [@wagenmakersPracticalSolutionPervasive2007] \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (https://osf.io/3tv6x).
+There was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.}
 
@@
 
+```{r, message=FALSE,warning=FALSE,echo=FALSE}
+# load the summary dataframes to the environment
+load(here("models/RobuEstMultiLevelModel_ri_only_log_mean_mod_strength"))
+load(here("models/RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy"))
+
+```
+
 Instead, we can use a version of the meta-regression model described above (see equation \@ref(eq:lnm-lns-model-eq) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-\textcolor{red}{This model with just $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
-This reflects the pairwise nature of the $\textrm{ln}CVR$.} Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other extensions[^20] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot) for the best fitting model plots).
-
-[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
-The comparison of these models is included in the supplementary materials (strength - https://osf.io/n4wgk; hypertrophy - https://osf.io/hf8dx).
-In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy intercept only model fit best.
+\textcolor{red}{This kind of meta-regression model with  $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
+This reflects the pairwise nature of the $\textrm{ln}CVR$.}
 
@@
 
-\textcolor{red}{Despite the models comparability in terms of what they are trying to estimate (i.e., the difference in variance in change scores between the RT and CON groups) we arrive at substantively different conclusions compared to the $\textrm{ln}CVR$ model and dependent upon the exact specification (see footnote 20) of the meta-regression model.
-Table} \@ref(tab:model-estimates-tab) \textcolor{red}{shows the different estimates from each model for this difference.
-Where the $\textrm{ln}CVR$ suggested that the introduction of an RT intervention actually reduced variance in change scores (i.e., the estimate was negative) we see the opposite conclusion (i.e., the estimate is positive) for all multilevel meta-regression models of strength, and for hypertrophy this appears to be dependent on whether the coefficients (i.e., slopes) for the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ are allowed to vary between studies and arms.}
+Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other model specifications[^20]\textsuperscript{,}[^21] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot)).
+The results were largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
+Both models suggest that the introduction of the RT interventions *reduced* variance in change scores.
 
-```{r model-estimates-tab, message=FALSE,warning=FALSE,echo=FALSE}
-# load the summary dataframes to the environment
-load(here("models/group_model_estimates"))
+[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with \textcolor{red}{(1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).
+The comparison of these models using $2\times\textrm{log}BF$} [@kassBayesFactors1995] \textcolor{red}{from approximate Bayesian information criterion} [@wagenmakersPracticalSolutionPervasive2007] \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (see https://osf.io/b5deh for strength and https://osf.io/5ektd for hypertrophy).
+Similar to the models including $Outcome$ there was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.
+All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (https://osf.io/5g7ce) all of which revealed similar conclusions.}
 
-knitr::kable(
-  model_estimates, 
-  col.names = c("Outcome", "Model", "Estimate", "Lower 95% CI", "Upper 95% CI"),
-  caption = "Comparison of estimates from model using $\\textrm{ln}CVR$ and multilevel meta-regression models of $\\textrm{ln}\\hat\\sigma$ and $\\textrm{ln}\\overline{x}$ with group (RT vs CON) as a predictor",
-) %>%
-  row_spec(0, bold = TRUE) %>%
-  pack_rows(index = c("Strength" = 4, "Hypertrophy" = 4)) %>%
-  footnote(general = c("CI = confidence interval", "MLMR = multilevel meta-regression")) %>%
-  kable_classic(full_width = FALSE) %>%
-  kable_styling()
-
-```
-
-\textcolor{red}{It is worth explaining why this might be the case to aid readers in considering the assumptions that different models make.
-The $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; In the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
-In both however this is a strong assumption.
-The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in the random slope models this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
-Evidently, the assumptions that are made about the data-generating processes can impact upon the modelling strategy chosen}[^21].
-\textcolor{red}{Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression model allows).
-It is not necessarily clear which model should be preferred here, but thought should be given to the assumptions each makes and the fit of each model to the data.}
-
-[^21]: \textcolor{red}{It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
+[^21]: \textcolor{red}{It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variance relationship.
+For example, the $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; in the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
+In both however this is a *strong* assumption.
+The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in model where random slopes are included this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
+Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression models allow).
+It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
 Given that for most effects that might be included in such models we can determine the sampling variance for $\overline{x}$ one approach to address this might be to employ models that incorporate the variance on this predictor (i.e., measurement error or errors in variables models).
-This is beyond the scope of this paper to discuss.}
+This is beyond the scope of this paper to discuss.
+It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.}
 
@@
 \textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size.
-The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}
+The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}
 
@@
 
-It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
+It is of particular interest to note the different conclusions drawn here are dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
 \textcolor{red}{We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
@@
 But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.
-Lastly, given the possible sensitivity of substantive conclusions to the exact model specifications used it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
+Lastly, it is possible that substantive conclusions might be sensitive to the exact model specifications used.
+As such, it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
 With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions} [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
 
-Whether or not there is clear evidence for average intervention effects, research aimed at exploring variation is important.
-Even in the absence of an average intervention effect it may be that a large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
+The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
+Examining this data using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation.
+In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive (as seen in the SMD and $\textrm{ln}RR$ models), yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
+Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
 Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.
-Interventions, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
+
+However, similar to the cross-sectional pre-intervention score models reported here (and indeed most physical and biological variables), change scores also demonstrated a mean-variance relationship in addition to heteroskedasticity.
+The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite substantive direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain.
+Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
 Interventions such as these are valuable for the simplification of guidelines and recommendations.
-This also potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome.
+For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations [@steeleHigherEffortbasedParadigm2017; @steeleLongTermTimeCourseStrength2022].
+
+The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis.
+Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021].
+However, this potentially represents another interesting area of future study regarding variation; specifically, how to produce interventions that actually reduce variance in an outcome.
 In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
-
-The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
-\textcolor{red}{Given the varied results, this dataset does not necessarily provide clear answers as to the question of whether true interindividual response variation is present and so this remains an open area of research.
-However, the examples provided highlight the importance of considering carefully the assumptions inherent in the modelling approaches chosen to explore such a substantive topic.
-We hope they also prompt researchers to further explore variation.}
 

### 2023-09-14 paper/Revised Pre-print - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Re-revise manuscript, update pre-print

@@
   - Shinichi Nakagawa$^d$
-date: "12th September 2023"
+date: "14th September 2023"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
@@
    - \captionsetup[figure]{font=small}
+---
+
 ---
@@
 
-Although historically fixed effects models were commonplace in the field of sport and exercise science [@haggerMetaanalysisSportExercise2006] nowadays the random effects model is more often employed [@haggerMetaanalysis2022].
+Although historically fixed effect models were commonplace in the field of sport and exercise science [@haggerMetaanalysisSportExercise2006] nowadays the random effects model is more often employed [@haggerMetaanalysis2022].
 This is likely due to the fact that direct (i.e., exact) replication of studies is rare[^4] and instead studies often explore similar effects across varying moderating factors such as those noted above.
@@
 We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the overall average slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
-Indeed, we fit additional models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm.
-The comparison of these models is included in the supplementary materials (https://osf.io/4xrcg).
-Note, the addition of both random slopes for study, and for arm, improved model fit significantly, though we limit presentation in the main text to the simpler model.
+These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms.
+We could also do the same and include random slopes for $(\beta_{1}$ on $\textrm{ln}\overline{x}$ thus allowing for the strength of the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ to also vary between studies and arms.
+Indeed, we fit a range of models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Outcome$ by study, (5) the inclusion of correlated random slopes for $Outcome$ by study and arm, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study, (7) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and arm.
+The comparison of these models using $2\times\textrm{log}BF$ [@kassBayesFactors1995] from approximate Bayesian information criterion [@wagenmakersPracticalSolutionPervasive2007] to determine under which is the observed data most likely is included in the supplementary materials (https://osf.io/3tv6x).
+There was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.
 
@@
 
+```{r, message=FALSE,warning=FALSE,echo=FALSE}
+# load the summary dataframes to the environment
+load(here("models/RobuEstMultiLevelModel_ri_only_log_mean_mod_strength"))
+load(here("models/RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy"))
+
+```
+
 Instead, we can use a version of the meta-regression model described above (see equation \@ref(eq:lnm-lns-model-eq) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-This model with just $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
+This kind of meta-regression model with  $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
 This reflects the pairwise nature of the $\textrm{ln}CVR$.
-Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other extensions[^20] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot) for the best fitting model plots).
-
-[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts for study and arm only, (2) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study, and (3) the inclusion of random slopes for $\textrm{ln}\overline{x}$ by study and arm.
-The comparison of these models is included in the supplementary materials (strength - https://osf.io/n4wgk; hypertrophy - https://osf.io/hf8dx).
-In this case, for strength the addition of random slopes for study, but not for arm, improved model fit significantly, and for hypertrophy intercept only model fit best.
 
@@
 
-Despite the models comparability in terms of what they are trying to estimate (i.e., the difference in variance in change scores between the RT and CON groups) we arrive at substantively different conclusions compared to the $\textrm{ln}CVR$ model and dependent upon the exact specification (see footnote 20) of the meta-regression model.
-Table \@ref(tab:model-estimates-tab) shows the different estimates from each model for this difference.
-Where the $\textrm{ln}CVR$ suggested that the introduction of an RT intervention actually reduced variance in change scores (i.e., the estimate was negative) we see the opposite conclusion (i.e., the estimate is positive) for all multilevel meta-regression models of strength, and for hypertrophy this appears to be dependent on whether the coefficients (i.e., slopes) for the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ are allowed to vary between studies and arms.
+Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other model specifications[^20]\textsuperscript{,}[^21] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot)).
+The results were largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
+Both models suggest that the introduction of the RT interventions *reduced* variance in change scores.
 
-```{r model-estimates-tab, message=FALSE,warning=FALSE,echo=FALSE}
-# load the summary dataframes to the environment
-load(here("models/group_model_estimates"))
+[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).
+The comparison of these models using $2\times\textrm{log}BF$ [@kassBayesFactors1995] from approximate Bayesian information criterion [@wagenmakersPracticalSolutionPervasive2007] to determine under which is the observed data most likely is included in the supplementary materials (see https://osf.io/b5deh for strength and https://osf.io/5ektd for hypertrophy).
+Similar to the models including $Outcome$ there was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.
+All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (https://osf.io/5g7ce) all of which revealed similar conclusions.
 
-knitr::kable(
-  model_estimates, 
-  col.names = c("Outcome", "Model", "Estimate", "Lower 95% CI", "Upper 95% CI"),
-  caption = "Comparison of estimates from model using $\\textrm{ln}CVR$ and multilevel meta-regression models of $\\textrm{ln}\\hat\\sigma$ and $\\textrm{ln}\\overline{x}$ with group (RT vs CON) as a predictor",
-) %>%
-  row_spec(0, bold = TRUE) %>%
-  pack_rows(index = c("Strength" = 4, "Hypertrophy" = 4)) %>%
-  footnote(general = c("CI = confidence interval", "MLMR = multilevel meta-regression")) %>%
-  kable_classic(full_width = FALSE) %>%
-  kable_styling()
-
-```
-
-It is worth explaining why this might be the case to aid readers in considering the assumptions that different models make.
-The $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; In the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
-In both however this is a strong assumption.
-The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in the random slope models this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
-Evidently, the assumptions that are made about the data-generating processes can impact upon the modelling strategy chosen[^21].
-Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression model allows).
-It is not necessarily clear which model should be preferred here, but thought should be given to the assumptions each makes and the fit of each model to the data.
-
-[^21]: It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
+[^21]: It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variance relationship.
+For example, the $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; in the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
+In both however this is a *strong* assumption.
+The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in model where random slopes are included this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
+Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression models allow).
+It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
 Given that for most effects that might be included in such models we can determine the sampling variance for $\overline{x}$ one approach to address this might be to employ models that incorporate the variance on this predictor (i.e., measurement error or errors in variables models).
 This is beyond the scope of this paper to discuss.
+It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.
 
@@
 But these are limited to the pairwise comparisons of a categorical variable as the effect size.
-The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories including any number of arbitrary predictors, fixed and random and assumptions about their correlations.
+The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.
 
@@
 
-It is of particular interest to note the different conclusions drawn here dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
+It is of particular interest to note the different conclusions drawn here are dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
 We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
@@
 But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore.
-Lastly, given the possible sensitivity of substantive conclusions to the exact model specifications used it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
+Lastly, it is possible that substantive conclusions might be sensitive to the exact model specifications used.
+As such, it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
 With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
 
-Whether or not there is clear evidence for average intervention effects, research aimed at exploring variation is important.
-Even in the absence of an average intervention effect it may be that a large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
+The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
+Examining this data using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation.
+In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive (as seen in the SMD and $\textrm{ln}RR$ models), yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
+Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
 Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.
-Interventions, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
+
+However, similar to the cross-sectional pre-intervention score models reported here (and indeed most physical and biological variables), change scores also demonstrated a mean-variance relationship in addition to heteroskedasticity.
+The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite substantive direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain.
+Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
 Interventions such as these are valuable for the simplification of guidelines and recommendations.
-This also potentially represents another interesting area of future study regarding variation opposite to the usual search for individual variation; specifically, how to produce interventions that actually reduce variance in an outcome.
+For example, muscle strengthening interventions such as RT are recommended for *everyone* in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations [@steeleHigherEffortbasedParadigm2017; @steeleLongTermTimeCourseStrength2022].
+
+The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis.
+Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021].
+However, this potentially represents another interesting area of future study regarding variation; specifically, how to produce interventions that actually reduce variance in an outcome.
 In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.
-
-The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
-Given the varied results, this dataset does not necessarily provide clear answers as to the question of whether true interindividual response variation is present and so this remains an open area of research.
-However, the examples provided highlight the importance of considering carefully the assumptions inherent in the modelling approaches chosen to explore such a substantive topic.
-We hope they also prompt researchers to further explore variation.
 

### 2023-09-14 paper/Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous.tex
Re-revise manuscript, update pre-print

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\captionsetup[figure]{font=small}
\usepackage{lineno}
\linenumbers
\usepackage{flafter}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Meta-Analysis of Variation in Sport and Exercise Science},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Meta-Analysis of Variation in Sport and Exercise Science}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Examples of Application Within Resistance Training Research}
\author{}
\date{14th September 2023}

\begin{document}
\maketitle
\begin{abstract}
Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies. However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance. It is thought that \textcolor{red}{individual responses to interventions vary considerably}. Hence, interest has increased in exploring \emph{precision} or \emph{personalised} exercise approaches. Not only is the mean often affected by interventions, but variances may also be impacted. Exploration of variances in studies such as randomised controlled trials (RCTs) can yield insight into interindividual heterogeneity in response to interventions and help determine generalisability of effects. Yet, larger samples sizes than those used for typical mean effects are required when probing variances. Thus, \textcolor{red}{in} a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing. \textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Although the quantitative synthesis of results across studies has existed since the 17th century (Plackett, 1958), the modern-day term ``meta-analysis'' was coined by Gene Glass (1976). Since that time, the use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably (Hagger, 2022), with resistance training (RT) accounting for a considerable proportion of this growth (figure @ref(fig:trend-plot)). Accordingly, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers. \textcolor{red}{However, to begin we provide a conceptual overview of meta-analyses and effect sizes as typically used within sport and exercise science.}

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/trend-plot-1}

}

\caption{Trends in meta-analyses published in sport and exercise science since 1976.}(\#fig:trend-plot)
\end{figure}

\textcolor{red}{There are two popular models}\footnote{\textcolor{red}{To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics. A statistical model essentially is specification of what we think the data generating process might be for a given situation. In the context of meta-analyses the data are usually the individual effects that we have extracted from studies i.e., the results of each study. The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.}} \textcolor{red}{used for meta-analysis: the fixed-effect model and the random-effects model} (Borenstein et al., 2010). \textcolor{red}{The fixed effect model assumes that there is one true effect size}\footnote{\textcolor{red}{*Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research. This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between} (Caldwell \& Vigotsky, 2020) \textcolor{red}{including, as we shall see, statistics describing variation.}} \textcolor{red}{that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error. This essentially means that there is a single common effect which is fixed across studies and each study takes samples of individuals from the population to estimate this effect. We can express this model in the following formula:}

\begin{equation}
\hat\theta_{i}=\theta+m_{i} 
(\#eq:fixed-model-eq)
\end{equation}

\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$. Studies with smaller standard errors of their effect estimates have smaller sampling errors and so these higher precision estimates are given greater weight in the model. The weighting given to each study is calculated as:}

\begin{equation}
w_{i}=\frac{1}{s^2_{i}}
(\#eq:fixed-weights-eq)
\end{equation}

\textcolor{red}{Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance. It is referred to as inverse-variance weighting. Then, the overall weighted mean effect estimate from the model is then calculated as:}

\begin{equation}
\hat\theta=\frac{\sum^I_{i=1}\hat\theta_{i}w_{i}}{\sum^I_{i=1}w_{i}}
(\#eq:fixed-average-eq)
\end{equation}

\textcolor{red}{Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect. Instead it allows for the true effects that each study estimates to differ. Each study may share a common underlying effect size but, due to differences between studies in factors such as population characteristics, the manner in which outcomes are operationalised, or subtle differences in intervention and context to name a few, it is possible that the actual effect being estimated by each study differs. The assumption of the random-effects model then is that the studies included estimate effects that come from a larger population of effects determined by the inclusion criteria for studies included. So, the model assumes that the studies included reflect a random sample of all possible permutations of study and the effects they estimate from this population distribution of studies and effects. Hence, in the fixed-effect model there is one effect and it is assumed to be *fixed* across studies, whereas in the random-effects model there are many and we examine an assumed *random* sample of them. We can express this model in the following formula:}

\begin{equation}
\hat\theta_{i}=\theta_{i}+\tau_{i}+m_{i}
(\#eq:random-model-eq)
\end{equation}

\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$. Essentially this model assumes that each individual study estimates an effect and there will be some sampling error in estimating it. But, the effects each study estimates comes from an overall distribution of true effects with a mean value (i.e., $\theta_{i}$). The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated}\footnote{\textcolor{red}{This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform. This is beyond the scope of this paper to discuss. We note however that the models we present all utilise Restricted Maximum Likelihood estimation.}}. \textcolor{red}{This essentially describes the heterogeneity between the effects included. Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:}

\begin{equation}
w^*_{i}=\frac{1}{s^2_{i}+\tau^2}
(\#eq:random-weights-eq)
\end{equation}

\textcolor{red}{Then the average of the distribution of effects can be calculated as per equation} @ref(eq:fixed-average-eq) \textcolor{red}{substituting $w_{i}$ for $w^*_{i}$.}

\textcolor{red}{Although historically fixed effect models were commonplace in the field of sport and exercise science} (Hagger, 2006) \textcolor{red}{nowadays the random effects model is more often employed} (Hagger, 2022). \textcolor{red}{This is likely due to the fact that direct (i.e., exact) replication of studies is rare}\footnote{\textcolor{red}{Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).}} \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above.}

As with many other fields (Mills et al., 2021; Nakagawa et al., 2015; Usui et al., 2021) \textcolor{red}{likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable. The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping}; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex. Indeed, a recent umbrella review (Bernárdez-Vázquez et al., 2022) of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes, all of which focused on the comparison of mean changes between different intervention groups.

\textcolor{red}{Often, due to the varying operationalisations used for broad outcome concepts}\footnote{\textcolor{red}{For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions. Or the same operationalisations may be employed but different exercises such as the squat or bench press.}}\textcolor{red}{, an effect size is used which is "standardised" across studies to allow for their synthesis.} Most commonly, a magnitude based\footnote{Though notably not all meta-analyses use \emph{magnitude based} effect sizes. Indeed some explicitly use what Caldwell and Vigotsky (2020) term \emph{signal-to-noise} effect sizes (e.g., Heidel et al. (2022)).} effect size statistic (Caldwell \& Vigotsky, 2020), the standardised mean difference (SMD), is used to compare means between groups or conditions. This statistic is usually a version of Cohen's \emph{d} (Cohen, 1988), or its bias-corrected\footnote{\textcolor{red}{For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct. Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.}} metric referred to as Hedges' \emph{g} {[}Hedges \& Olkin (2014); Borenstein et al. (2021); Nakagawa \& Cuthill (2007){]}\footnote{We will refer to both merely as the SMD throughout the manuscript for simplicity and note that throughout when reporting a `SMD' we are reporting the bias-corrected version. \textcolor{red}{We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance;} Caldwell \& Vigotsky (2020)).}. The SMD, and its sampling variance, \(s^2_{SMD}\) are given by:

\begin{equation}
SMD=\frac{\overline{x}_E - \overline{x}_C}{s_{pooled}}J
(\#eq:SMD-eq)
\end{equation}

\begin{equation}
J=1-\frac{3}{4(n_{C}+n_{E})-2)-1}
(\#eq:biasJ-eq)
\end{equation}

\begin{equation}
s_{pooled}=\sqrt{\frac{(n_{C}-1)s^2_{C}+(n_{E}-1)s^2_{E}}{n_{C}+n_{E}-2}}
(\#eq:s-pooled-eq)
\end{equation}

\begin{equation}
s^2_{SMD}=\frac{n_{C}+n_{E}}{n_{C}n_{E}}+\frac{SMD^2}{2(n_{E}+n_{C})}
(\#eq:SMDvar-eq)
\end{equation}

where \(\overline{x}_C\) and \(\overline{x}_E\) are the sample means of the control group (C) and experimental (E) or intervention group respectively, \(s_C\) and \(s_E\) are the standard deviations of the two groups, \(n_C\) and \(n_E\) are the sample sizes of the two groups, and \(J\) is a bias correction for small sample sizes.

The natural logarithm of the ratio of two means (\(\textrm{ln}RR\)) is also another effect size statistic that can be used (Curtis \& Wang, 1998; Hedges et al., 1999; Lajeunesse, 2011, 2015). The lnRR, and its sampling variance, \(s^2_{\textrm{ln}RR}\) are given by:

\begin{equation}
\textrm{ln}RR=\textrm{ln}\frac{\overline{x}_E}{\overline{x}_C}
(\#eq:lnRR-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}RR}=\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}+\frac{s^4_{C}}{2n_{C}\overline{x}^2_{C}\overline{x}^4_{C}}+\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}}+\frac{s^4_{E}}{2n_{E}\overline{x}^2_{E}\overline{x}^4_{E}}
(\#eq:lnRRvar-eq)
\end{equation}

Due to its calculation the SMD is affected not only by the difference in means of the two groups, but also by the standard deviations of both groups due to the standardisation of the effect size by \(s_{pooled}\) in the denominator. In contrast, the \(\textrm{ln}RR\) is uninfluenced by the standard deviations in either groups (see equation @ref(eq:lnRR-eq)), which only affects the sampling variance (see equation @ref(eq:lnRRvar-eq)). Despite this, the use of effect sizes like the \(\textrm{ln}RR\) has been limited in previous meta-analyses in sport and exercise science (Deb et al., 2018; Nuzzo et al., 2023a) and to our knowledge only a couple of meta-analyses of RT interventions has used this kind of effect size (Swinton et al., 2022; Wolf et al., 2023).

Although researchers in sport and exercise science, among other fields, have focused on estimating the average effects of interventions using randomised trial designs for both primary research and synthesis through meta-analysis, responses to certain interventions may vary on a subgroup or even individual basis. The increased interest in \emph{precision} or \emph{personalised} approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions (Atkinson et al., 2019; Atkinson \& Batterham, 2015; Hecksteden et al., 2015; Hopkins, 2015; Hrubeniuk et al., 2022; Kelley, 2022; Pickering \& Kiely, 2019; Ross et al., 2019; Swinton et al., 2018). However, despite the availability of approaches to compare variances between groups, in sport and exercise science this is rarely explored in primary research (Bonafiglia et al., 2022). Moreover, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter (Bonafiglia et al., 2022; Esteves et al., 2021; Fisher et al., 2022; Kelley et al., 2020, 2022; Steele et al., 2021). Examination of interindividual heterogeneity in response to interventions presents considerable value to researchers and practitioners in sport and exercise science; interventions with low interindividual variation are likely to be widely generalisable, whilst an intervention with high interindividual variation is likely to have effects that are either subgroup or individual specific. The former kind of intervention might be widely applicable across individuals, whilst the latter kind of intervention requires specific research, typically with large samples (Hecksteden et al., 2015), to tease out subgroup- or participant-by-intervention interactions to facilitate successful practical application.

Comparison of heterogeneity in responses, such as post-scores or change scores to interventions, are not the only possible use of statistical methods for comparing variances. For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs (Nakagawa \& Schielzeth, 2012). Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods\footnote{Exploration of methodological approaches and their impact on heterogeneity have also been explored in preclinical research (Usui et al., 2021).} to reduce RT stimulus heterogeneity, (Exner et al., 2022) as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT (Halperin et al., 2022) \textcolor{red}{and in the number of repetitions that can be performed at different percentages of one repetition maximum} (Nuzzo et al., 2023b).

Given the value of embracing and exploring variation alongside mean effects in sport and exercise science, yet the lack of application in research synthesis by way of primary research or meta-analysis, we present and discuss effect size approaches and models for meta-analysis of variation. Indeed, meta-analysis presents a very valuable method for exploring variation in a field such as sport and exercise science due to the typically small samples in primary studies. Such small samples have even lower statistical power to detect differences in variation as compared to means (Yang et al., 2022).

\textcolor{red}{Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses. We hope this will make clear why we offer such recommendations in our discussion and conclusion. As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.}

\hypertarget{effect-size-statistics-for-meta-analytic-comparisons-of-variation}{%
\section{Effect size statistics for meta-analytic comparisons of variation}\label{effect-size-statistics-for-meta-analytic-comparisons-of-variation}}

Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework. However, several have been proposed that we now describe: the standard deviation for individual responses (\(SD_{ir}\); Hopkins (2015); Atkinson \& Batterham (2015); Atkinson et al. (2019)), the log ratio of standard deviations (\(\textrm{ln}VR\); termed the ``variability ratio''; Hedges \& Nowell (1995)), and the log ratio of coefficient of variation (\(\textrm{ln}CVR\); termed the ``coefficient of variation ratio''; Nakagawa et al. (2015); Senior et al. (2020)). We present the independent groups versions due to use of randomised controlled trials in our examples below, but note that dependent versions (i.e., for comparing related samples) also exist for \(\textrm{ln}VR\) and \(\textrm{ln}CVR\) (Senior et al., 2020).

\hypertarget{standard-deviation-for-individual-responses-sd_ir}{%
\subsection{\texorpdfstring{Standard deviation for individual responses (\(SD_{ir}\))}{Standard deviation for individual responses (SD\_\{ir\})}}\label{standard-deviation-for-individual-responses-sd_ir}}

In the context of \emph{precision} or \emph{personalised} approaches to exercise prescription the \(SD_{ir}\) has been proposed as an approach to determine the extent to which individual responses manifest by comparison of variation between two groups; control and intervention (Atkinson et al., 2019; Atkinson \& Batterham, 2015; Hopkins, 2015). The standard deviation of change scores (post-intervention scores minus pre-intervention scores) within the intervention group reflects the gross combination of a number of sources of variation including: participant-by-intervention interactions (i.e., actual individual responsiveness or `trainability'), within-participant variability in intervention response (i.e., variability in response to the same intervention administered to the same participant), and random error (i.e., from pre and post measurements; Hecksteden et al. (2015)). The standard deviation of change scores from the control group (assuming it is a non-intervention control group and not something like a `usual-care' group) by contrast is assumed to only reflect random error\footnote{Though notably, in the case of health behaviour studies it may be the case that if someone volunteers for a study it could conceivably motivate them to alter various habits even when they are assigned to a control group thus influencing change scores.} (Hecksteden et al., 2015). As such, the difference in these standard deviations can be used to determine the extent to which additional variation has been introduced by the intervention and that might reflect individual responses. Whilst the \(SD_{ir}\) has been proposed and used primarily in the context of individual response variation to interventions, it should be noted that this kind of absolute comparison of variance between groups or conditions is not limited to such applications.

The \(SD_{ir}\), and its sampling variance, \(s^2_{SD_{ir}}\) are given by:

\begin{equation}
SD_{ir}=\sqrt{s^2_{E} - s^2_{C}}
(\#eq:SDir-eq)
\end{equation}

\begin{equation}
s^2_{SD_{ir}}=2\biggl(\frac{s^4_{E}}{n_{E}-1}+\frac{s^4_{C}}{n_{C}-1}\biggl)
(\#eq:sdIRvar-eq)
\end{equation}

Thus, the \(SD_{ir}\) reflects a comparison of the absolute variance in change scores between control and intervention groups. However, a potential concern with the \(SD_{ir}\) is its potential to violate assumptions of normality, which is not the case for other effect size statistics such as \(\textrm{ln}VR\) and \(\textrm{ln}CVR\).

\hypertarget{log-ratio-of-standard-deviations-textrmlnvr}{%
\subsection{\texorpdfstring{Log ratio of standard deviations (\(\textrm{ln}VR\))}{Log ratio of standard deviations (\textbackslash textrm\{ln\}VR)}}\label{log-ratio-of-standard-deviations-textrmlnvr}}

A similar effect size statistic for the comparison of absolute variance between groups, and one that has had wide applications in more than just intervention response variability within fields such as ecology and evolution, is the \(\textrm{ln}VR\) (Hedges \& Nowell, 1995; Nakagawa et al., 2015; Senior et al., 2020). An unbiased estimator of the natural logarithm of a population standard deviation (\(\textrm{ln}\sigma\)), and its sampling variance, \(s^2_{\textrm{ln}\sigma}\) is given by:

\begin{equation}
\textrm{ln}\hat\sigma=\textrm{ln}s+\frac{1}{2(n-1)}
(\#eq:lnsigma-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}\hat\sigma}=\frac{1}{2(n-1)}
(\#eq:lnsigmavar-eq)
\end{equation}

where \(\textrm{ln}\hat\sigma\) is an estimate of \(\textrm{ln}\sigma\), and it is assumed with sufficiently large sample size and value of \(\sigma\) that \(\textrm{ln}\sigma\) is normally distributed with variance \(s^2_{\textrm{ln}\sigma}\). Given equations @ref(eq:lnsigma-eq) and @ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the \(\textrm{ln}VR\), and its sampling variance, \(s^2_{\textrm{ln}VR}\) is given by:

\begin{equation}
\textrm{ln}VR=\textrm{ln}\biggl(\frac{s_{E}}{s_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnVR-eq)
\end{equation}

\begin{equation}
s^2_{\textrm{ln}VR}=\frac{1}{2}\biggl(\frac{n_{C}}{(n_{C}-1)^2}+\frac{n_{E}}{(n_{E}-1)^2}\biggl)
(\#eq:lnVRvar-eq)
\end{equation}

However, due to both \(SD_{ir}\) and \(\textrm{ln}VR\) being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparison group (e.g., when \(\overline{x}_E\) is larger than \(\overline{x}_C\)). In this case, it is likely that the standard deviation will be larger in the group with the larger mean (e.g., \(s_{E}\) is larger than \(s_{C}\)). This mean-variance relationship is common for many variables and datasets\footnote{For one clear example, see figure 1A in Vigostky et al. (2020) who show that the mean and standard deviation for baseline strength values typically scale with one another across most studies.} and to highlight this we provide examples below. They also assume constant measurement error over the range of values for the mean, which can impact their utility for examining response variation (Tenan et al., 2020).

\hypertarget{log-ratio-of-coefficient-of-variation-textrmlncvr}{%
\subsection{\texorpdfstring{Log ratio of coefficient of variation (\(\textrm{ln}CVR\))}{Log ratio of coefficient of variation (\textbackslash textrm\{ln\}CVR)}}\label{log-ratio-of-coefficient-of-variation-textrmlncvr}}

The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present. \textcolor{red}{In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.} The natural logarithm of the ratio between the coefficients of variation from two groups, the \(\textrm{ln}CVR\) is thus a more generally applicable effect size statistic for examining variability between groups. Considering equations @ref(eq:lnRR-eq) and @ref(eq:lnVR-eq), the \(\textrm{ln}CVR\) is given by:

\begin{equation}
\textrm{ln}CVR=\textrm{ln}\biggl(\frac{CV_{E}}{CV_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
(\#eq:lnCVR-eq)
\end{equation}

where \(CV_{E}\) and \(CV_{C}\) are \({s_{E}}/{\overline{x}_{E}}\) and \({s_{C}}/{\overline{x}_{C}}\) respectively. Senior et al. (2020) derived the sampling variance, \(s^2_{\textrm{ln}CVR}\), as:

\begin{equation}
\begin{split}
s^2_{\textrm{ln}CVR}=\frac{s_{C}^2}{n_{C}\overline{x}_{C}^2}+\frac{s_{C}^4}{2n_{C}^2\overline{x}_{C}^4}+\frac{n_{C}}{(n_{C}-1)^2} \\
+\frac{s_{E}^2}{n_{E}\overline{x}_{E}^2}+\frac{s_{E}^4}{2n_{E}^2\overline{x}_{E}^4}+\frac{n_{E}}{(n_{E}-1)^2}
\end{split}
(\#eq:lnCVRvar-eq)
\end{equation}

\hypertarget{examples-using-resistance-training-studies}{%
\section{Examples using resistance training studies}\label{examples-using-resistance-training-studies}}

As noted, to facilitate understanding for those new to examination of variation, we provide primary examples of the approaches presented using data from RT studies included in a recent meta-analysis published in the \emph{Journal of Sport Sciences} (Polito et al., 2021). Here we have used their list of included studies and re-extracted data from 111 of these\footnote{The authors of the meta-analysis did not make their extracted data openly available, nor did they respond to our request for the extracted data. Further, their original analysis included 119 studies however we were unable to extract data for our analyses from 8 of these for a variety of reasons (e.g., only percentage change data was reported, no standard deviations for control groups reported).}. All analysis examples were performed in R (version 4.2.1, ``Funny-Looking Kid'', The R Foundation for Statistical Computing, 2022) using the \textbf{metafor} package (Viechtbauer, 2010). The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (\url{https://osf.io/2h9ma/}) \textcolor{red}{or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git)}.

\begin{table}

\caption{(\#tab:sample-size-tab)Sample sizes for resistance training and non training control groups for dataset.}
\centering
\begin{tabular}[t]{l|r}
\hline
\textbf{Arm} & \textbf{Sample Size}\\
\hline
\multicolumn{2}{l}{\textbf{RT}}\\
\hline
\hspace{1em}All & 2683\\
\hline
\hspace{1em}Minumum RT & 5\\
\hline
\hspace{1em}Median RT & 12\\
\hline
\hspace{1em}Maximum RT & 59\\
\hline
\multicolumn{2}{l}{\textbf{CON}}\\
\hline
\hspace{1em}All CON & 2349\\
\hline
\hspace{1em}Minumum CON & 4\\
\hline
\hspace{1em}Median CON & 10\\
\hline
\hspace{1em}Maximum CON & 44\\
\hline
\multicolumn{2}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{2}{l}{\rule{0pt}{1em}RT = resistance training}\\
\multicolumn{2}{l}{\rule{0pt}{1em}CON = non-training control}\\
\end{tabular}
\end{table}

Polito et al. (2021) conducted a systematic review and meta-analysis of randomised trials that included a RT intervention group(s) and a non-training control comparison group. Their analysis focused upon the SMD between the RT intervention group(s) and the control group from the studies included, with both overall effect estimate and moderator analyses (i.e., meta-regressions\footnote{\textcolor{red}{Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s). This can be extended to meta-analytic synthesis where the independent variables reflect characteristics associated with the effects included. For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.}}) performed. Given that Polito et al. (2021) included only studies with a non-training control group, their study selection offers a unique context to examine variation of interindividual responses specifically by comparing the variances in change scores between the RT intervention groups(s) and control group. Table @ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies. Indeed, this highlights the typically small samples used in sport and exercise science, and thus low power to detect difference in both means and variances in in individual studies (Yang et al., 2022), emphasising the value of meta-analysis to explore variation. Table @ref(tab:summary-characteristics-tab) shows the study and participant characteristics.

\begin{table}

\caption{(\#tab:summary-characteristics-tab)Summary of study and participant characteristics.}
\centering
\begin{tabular}[t]{l|c}
\hline
\textbf{Characteristic} & \textbf{Summary}\\
\hline
TESTEX & 7 (6, 8)\\
\hline
Age & 33 (23, 66)\\
\hline
Proportion Male & 100 (0, 100)\\
\hline
Weight & 74 (68, 78)\\
\hline
BMI & 26.62 (24.27, 27.34)\\
\hline
\multicolumn{2}{l}{Training Status}\\
\hline
\hspace{1em}Trained & 9 (4.6\%)\\
\hline
\hspace{1em}Untrained & 187 (95\%)\\
\hline
\multicolumn{2}{l}{Sample Type}\\
\hline
\hspace{1em}Clinical & 5 (2.6\%)\\
\hline
\hspace{1em}Healthy & 191 (97\%)\\
\hline
\multicolumn{2}{l}{RT Intervention Only?}\\
\hline
\hspace{1em}N & 9 (4.6\%)\\
\hline
\hspace{1em}Y & 187 (95\%)\\
\hline
Duration (weeks) & 12 (8, 16)\\
\hline
Weekly Frequency & 3.00 (2.00, 3.00)\\
\hline
Number of Exercises & 6 (2, 8)\\
\hline
Sets per Exercise & 3.00 (2.50, 3.00)\\
\hline
Number of Repetitions & 10.0 (8.0, 11.2)\\
\hline
Load (\%1RM) & 74 (65, 80)\\
\hline
\multicolumn{2}{l}{Task Failure?}\\
\hline
\hspace{1em}N & 29 (23\%)\\
\hline
\hspace{1em}Y & 95 (77\%)\\
\hline
\multicolumn{2}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{2}{l}{\rule{0pt}{1em}RT = resistance training; }\\
\multicolumn{2}{l}{\rule{0pt}{1em}Continuous variables are median (IQR); }\\
\multicolumn{2}{l}{\rule{0pt}{1em}Categorical variables are count (\%); }\\
\multicolumn{2}{l}{\rule{0pt}{1em}Not all studies reported full descriptive data (see dataset; https://osf.io/kg2z4)}\\
\end{tabular}
\end{table}

\hypertarget{detecting-the-presence-of-interindividual-response-variation-to-resistance-training-intervention}{%
\subsection{\texorpdfstring{Detecting the presence of interindividual response variation to resistance training intervention \textcolor{red}{with absolute variance statistics}}{Detecting the presence of interindividual response variation to resistance training intervention }}\label{detecting-the-presence-of-interindividual-response-variation-to-resistance-training-intervention}}

First we conducted a traditional SMD and \(\textrm{ln}RR\) based effect size\footnote{It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental intervention, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure). We presented the SMD and \(\textrm{ln}RR\) effect sizes in equations @ref(eq:SMD-eq) and @ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., (2000); Morris et al., (2007); Morris (2008); Lajeunesse (2011, 2015)) and these are the effect sizes used in the meta-analyses referred to here.} meta-analysis to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes\footnote{We also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases. There did not appear to be any obvious small study bias in the dataset (see \url{https://osf.io/stqr3}).}. Polito et al. (2021) originally used a normal random-effects meta-analysis as described in the introduction. However, the data we extracted were hierarchical in nature. \textcolor{red}{Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model (Van den Noortgate et al., 2013) with cluster-robust variance estimation (Hedges et al., 2010) was used with random intercepts for study, arm\footnote{We use the term \emph{arm} to refer to an intervention group-control group contrast to accommodate studies including multiple intervention groups. This is so as to not confuse the reader with the use of \emph{group} to designate either the RT intervention group(s) or control group separately. Thus, in the instances of models using effect sizes relating to comparisons between an intervention group and control group (i.e., SMD, \(\textrm{ln}RR\), \(SD_{ir}\), \(\textrm{ln}VR\), and \(\textrm{ln}CVR\)) we calculate comparisons \emph{between} each intervention group (i.e., arm) and the control group. Thus, where a study had for example two RT interventions and a control, two separate arms would be coded (RT intervention 1 compared to control, and RT intervention 2 compared to control). Data was coded such that study and arm had explicit nesting.}, and effect. \textcolor{red}{This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies. We can therefore describe the overall model as:}

\begin{equation}
\hat\theta_{ijk}=(\theta+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+m_{ijk}
(\#eq:ml-model-eq)
\end{equation}

\textcolor{red}{where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$. This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects}\footnote{\textcolor{red}{Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.}}. \textcolor{red}{The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).}

We then fit the same model for the \(SD_{ir}\) and \(\textrm{ln}VR\) effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls. A positive SMD or \(\textrm{ln}RR\) would indicate that RT interventions produced greater improvements in outcomes compared to controls, whilst a positive \(SD_{ir}\) and \(\textrm{ln}VR\) would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/forest-SMD-plot-1}

}

\caption{Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes.}(\#fig:forest-SMD-plot)
\end{figure}

The pattern of results from our models examining SMDs (figure @ref(fig:forest-SMD-plot)) were similar to those reported by Polito et al. (2021), albeit with slightly lower estimates for both outcome types; possibly due to our use of a multilevel mixed-effects meta-analysis model that allowed for each individual effect size to be more appropriately weighted (the relative amount of heterogeneity between effects for each level is presented as the \(I^2\) statistic). As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = 0.87 {[}95\%CI: 0.77 to 0.97{]}; \(I^2_{study}\) = 57.32\%, \(I^2_{arm}\) = 3\%, \(I^2_{effect}\) = 11.95\%) and hypertrophy outcomes (SMD = 0.34 {[}95\%CI: 0.29 to 0.39{]}; \(I^2_{study}\) = 54.62\%, \(I^2_{arm}\) = 0.62\%, \(I^2_{effect}\) = 2.79\%). Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure }@ref(fig:forest-SMD-plot)) \textcolor{red}{were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/forest-RR-plot-1}

}

\caption{Caterpillar plots of exponentiated RR effect sizes for strength (A) and hypertrophy (B) outcomes.}(\#fig:forest-RR-plot)
\end{figure}

For the \(\textrm{ln}RR\) results we exponentiated them and converted to percentages to be more interpretable. These were similar, with greater proportional increases in strength compared with hypertrophy (figure @ref(fig:forest-RR-plot)). Increases were seen for both strength (\(\textrm{exp}RR\) = 21.97 {[}95\%CI: 19.43 to 24.57{]}; \(I^2_{study}\) = 33.46\%, \(I^2_{arm}\) = 0\%, \(I^2_{effect}\) = 0\%) and hypertrophy (\(\textrm{exp}RR\) = 5.39 {[}95\%CI: 4.44 to 6.35{]}; \(I^2_{study}\) = 12.97\%, \(I^2_{arm}\) = 0\%, \(I^2_{effect}\) = 0\%). Confidence intervals were again precise for both outcomes, and whilst relative heterogeneity was lower compared to SMD models prediction intervals were still quite wide.

In addition to the SMD and \(\textrm{ln}RR\) results, both the \(SD_{ir}\) (figure @ref(fig:forest-SDir-plot)) and \(\textrm{ln}VR\) (figure @ref(fig:forest-logVR-plot)) were also positive for both strength (\(SD_{ir}\) = 0.91 {[}95\%CI: 0.36 to 1.47{]}; \(I^2_{study}\) = 53.85\%, \(I^2_{arm}\) = 0.04\%, \(I^2_{effect}\) = 0\%; \(\textrm{ln}VR\) = 0.9 {[}95\%CI: 0.77 to 1.02{]}; \(I^2_{study}\) = 56.36\%, \(I^2_{arm}\) = 0.77\%, \(I^2_{effect}\) = 27.36\%) and hypertrophy outcomes (\(SD_{ir}\) = 0.42 {[}95\%CI: 0.33 to 0.5{]}; \(I^2_{study}\) = 0.01\%, \(I^2_{arm}\) = 40.15\%, \(I^2_{effect}\) = 39.73\%; \(\textrm{ln}VR\) = 0.5 {[}95\%CI: 0.4 to 0.6{]}; \(I^2_{study}\) = 41.21\%, \(I^2_{arm}\) = 3.31\%, \(I^2_{effect}\) = 33.59\%) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation. Although, heterogeneity across the models and levels (study, arm, effect) were again relatively large and quite varied.

This additional variance might support previous perspectives (Carpinelli, 2017) that the considerable variation in responses to RT interventions typically observed are due to `true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is \emph{detectable} independent of the random error). However, as noted, both the \(SD_{ir}\) and \(\textrm{ln}VR\) assume constant variance over values of the mean \textcolor{red}{(i.e., that the variance is similar whether mean values are low or high)}. As we have seen from the SMD and \(\textrm{ln}RR\) models, RT interventions increase mean scores. Thus, if there is a mean-variance relationship in the data, an increase in the mean alone may be fully responsible for any apparent increase in variation. As such, we cannot rely solely on absolute comparisons of variance such as the \(SD_{ir}\) and \(\textrm{ln}VR\) to determine whether interindividual response variation is actually present. The \(\textrm{ln}CVR\) can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic. First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures \textcolor{red}{to emphasise the need to consider this assumption,} and introduce a modelling approach that can also be used to overcome some possible limitations with the \(\textrm{ln}CVR\) \textcolor{red}{and provide flexibility to accomodate wider applications}.

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/forest-SDir-plot-1}

}

\caption{Caterpillar plots of SDir effect sizes for strength (A) and hypertrophy (B) outcomes.}(\#fig:forest-SDir-plot)
\end{figure}
\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/forest-logVR-plot-1}

}

\caption{Caterpillar plots of ln$VR$ effect sizes for strength (A) and hypertrophy (B) outcomes.}(\#fig:forest-logVR-plot)
\end{figure}

\hypertarget{mean-variance-relationships-in-muscular-strength-and-hypertrophy}{%
\subsection{Mean-variance relationships in muscular strength and hypertrophy}\label{mean-variance-relationships-in-muscular-strength-and-hypertrophy}}

With meta-analytic models of variation we are not limited to solely exploring variation in responses to interventions (e.g., Halperin et al. (2022); Nuzzo et al. (2023b)). We can explore the relationships between variance in a number of outcomes and the impact of certain predictors on this in the form of meta-regression. For example, as noted, one possible predictor of variance is the mean itself. As such, we can model variance of each effect as the response itself with the mean of the effect as the predictor. The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models \textcolor{red}{(i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).} Therefore, we instead can use \(\textrm{ln}\hat\sigma\), which is unbounded. In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al. (2021).

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/mean-variance-pre-plot-1}

}

\caption{Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes.}(\#fig:mean-variance-pre-plot)
\end{figure}

As can be seen in figure @ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean (\(\overline{x}\)) and standard deviation (\(s\)). \textcolor{red}{The variance in standard deviations increases with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems (Taylor, 1961).

\begin{equation}
s^2=a\overline{x}^b
(\#eq:taylors-eq)
\end{equation}

where \(a\) and \(b\) are some constants. When this relationship holds, under most circumstances the standard deviation is not proportional to the mean. However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear \textcolor{red}{based upon the product and power logarithmic rules:}

\begin{equation}
2\textrm{ln}s=\textrm{ln}a+b\textrm{ln}\overline{x}
(\#eq:taylorsln-eq)
\end{equation}

Figure @ref(fig:mean-variance-pre-plot)(B) and (D) shows that the relationship between the mean and variance on the log scale better meets the assumption of normality. Given these the observations we have for \(\textrm{ln}\hat\sigma\) and \(\textrm{ln}\overline{x}\) come from multiple outcomes within multiple arms within studies we can also estimate this relationship using a multilevel mixed-effects meta-regression model similar to that applied above. In this case though we are including an additional \emph{predictor} variable, the \(\textrm{ln}\overline{x}\). For example, the following model specifies \(\textrm{ln}\overline{x}\) as a fixed effect with random intercepts for study, arm, and effect:

\begin{equation}
\textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
(\#eq:lnm-lns-model-eq)
\end{equation}

where \(\textrm{ln}\hat\sigma_{ijk}\) is the \(k\textrm{th}\) effect size, as in equation @ref(eq:lnsigma-eq), from the \(j\textrm{th}\) arm (\(j = 1,2,\cdots,N_{j}\); where \(N_{j}\) is the number of arms\footnote{In contrast to the models presented examining effect sizes relating to comparisons between and intervention group and control group, in the models examining \(\textrm{ln}\sigma_{ijk}\) and \(\textrm{ln}\overline{x}_{ijk}\) as a predictor the term \emph{arm} refers to both the intervention groups(s) and control group. Thus, where a study had for example two RT interventions and a control, three separate arms would be coded (RT intervention 1, RT intervention 2, and control). Data were again coded such that study and arm had explicit nesting.}) in the \(i\textrm{th}\) study (\(i = 1,2,\cdots,N_{i}\); where \(N_{i}\) is the number of studies), \(\textrm{ln}\overline{x}_{ijk}\) is the mean estimate for each effect size, \(\beta_{0}\) is the intercept or overall mean of the effects, \(\beta_{1}\) is the slope or regression coefficient for \(\textrm{ln}\overline{x}\), \(\tau_{i}\) is the deviation from \(\beta_{0}\) for the \(i\textrm{th}\) study, \(\tau_{j}\) is the deviation for the \(j\textrm{th}\) arm, \(\tau_{k}\) is the deviation for the \(k\textrm{th}\) effect, \(\epsilon_{ijk}\) is the residual for each effect size which is normally distributed with \(\sigma^2_{\epsilon}\), and \(m_{ijk}\) is the sampling error for each effect size normally distributed with \(\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}\).

\textcolor{red}{These kinds of models are highly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include \(\beta_{2}Outcome\) in the model with \(Outcome\) as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where \(\beta_{2}\) is the slope or regression coefficient for \(Outcome\) (most intuitively thought of as the difference between the two outcome types)\footnote{We do not have to limit ourselves to only fixed effect predictor terms as we have here. Indeed, for mixed effect models generally some argue that models should use a \emph{maximal} random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences} (Barr et al., 2013). We could model a categorical variable for the outcome type and using random effects include \((\beta_{2} + \varphi_{i})Outcome\) or \((\beta_{2} + \varphi_{i} + \varphi_{j})Outcome\) in the model with \(Outcome\) as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where \(\beta_{2}\) is the overall average slope or regression coefficient for \(Outcome\), and \(\varphi_{i}\) is the deviation (random slope) from \(\beta_{2}\) for the \(i\textrm{th}\) study and \(\varphi_{j}\) is the deviation for the \(j\textrm{th}\) arm. \textcolor{red}{These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms. We could also do the same and include random slopes for $(\beta_{1}$ on $\textrm{ln}\overline{x}$ thus allowing for the strength of the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ to also vary between studies and arms. Indeed, we fit a range of models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Outcome$ by study, (5) the inclusion of correlated random slopes for $Outcome$ by study and arm, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study, (7) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and arm.  The comparison of these models using $2\times\textrm{log}BF$} (Kass \& Raftery, 1995) \textcolor{red}{from approximate Bayesian information criterion} (Wagenmakers, 2007) \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (https://osf.io/3tv6x). There was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.}}.

Figure @ref(fig:model-mean-variance-pre-plot) shows this model fit visually where the size of the points reflects their weight in the model. Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight but systematically greater degree of variance in strength measures compared to hypertrophy for a given mean score.

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/model-mean-variance-pre-plot-1}

}

\caption{Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores.}(\#fig:model-mean-variance-pre-plot)
\end{figure}

The presence of Taylor's law type relationships should be examined in datasets prior to deciding on which variance effect size statistic should be employed. Returning to the context of interindividual response variation to interventions, the presence of a mean-variance relationship in the data would imply that we cannot rely on absolute comparisons of variance (i.e., \(SD_{ir}\) or \(\textrm{ln}VR\)) to determine whether interindividual response variation is actually present. So we should also explore this assumption for the change-scores in the RT and control groups and determining the appropriate effect sizes to explore.

\hypertarget{reanalysis-of-interindividual-response-variation-using-textrmlncvr}{%
\subsection{\texorpdfstring{Reanalysis of interindividual response variation using \(\textrm{ln}CVR\)}{Reanalysis of interindividual response variation using \textbackslash textrm\{ln\}CVR}}\label{reanalysis-of-interindividual-response-variation-using-textrmlncvr}}

As can be seen in figures @ref(fig:mean-variance-delta-plot)(A) and (C) there is also a mean-variance relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation. Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures @ref(fig:mean-variance-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seen with the pre-intervention scores above. As such, in this case, we cannot rely solely on absolute comparisons of variance such as the \(SD_{ir}\) and \(\textrm{ln}VR\) to determine whether interindividual response variation is actually present.

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/mean-variance-delta-plot-1}

}

\caption{Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes.}(\#fig:mean-variance-delta-plot)
\end{figure}

The \(\textrm{ln}CVR\) can be used to overcome this issue though. Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before (see equation @ref(eq:ml-model-eq)) using the \(\textrm{ln}CVR\) as the effect size statistic leads to different conclusions compared to absolute variance comparisons using \(SD_{ir}\) or \(\textrm{ln}VR\). The introduction of an RT intervention actually \emph{reduces} the relative variation seen in change scores for strength (\(\textrm{ln}CVR\) = -0.61 {[}95\%CI: -0.76 to -0.47{]}; \(I^2_{study}\) = 23.18\%, \(I^2_{arm}\) = 0\%, \(I^2_{effect}\) = 0\%) and hypertrophy (\(\textrm{ln}CVR\) = -0.45 {[}95\%CI: -0.61 to -0.29{]}; \(I^2_{study}\) = 10.03\%, \(I^2_{arm}\) = 0\%, \(I^2_{effect}\) = 0\%) and further there is lower relative heterogeneity between studies in the effect estimates (figure @ref(fig:forest-logCVR-plot)).

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/forest-logCVR-plot-1}

}

\caption{Caterpillar plots of ln$CVR$ effect sizes for strength (A) and hypertrophy (B) outcomes.}(\#fig:forest-logCVR-plot)
\end{figure}

There is, however, a potential limitation for the \(\textrm{ln}CVR\) also that may need to be considered. Firstly, it is limited to the use of ratio scale data (which is not the case for the \(\textrm{ln}\hat\sigma\) or \(\textrm{ln}VR\)); hence the need to transform the change scores to be positively signed in this specific case. Secondly, whilst the \(\textrm{ln}CVR\) is useful in situations where there is a mean-variance relationship, the use of the \(CV\) in the effect size statistic assumes proportionality between standard deviation and mean. Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure @ref(fig:mean-variance-delta-plot)) an alternative yet \textcolor{red}{comparable} approach might be desirable. \textcolor{red}{Lastly, this statistic is limited to examination of pairwise comparisons of variance. In this example this approach poses no issue as we are comparing RT intervention group(s) to a control group. But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a \textcolor{red}{comparable} effect of interest to the $\textrm{ln}CVR$ one just applied.}

\hypertarget{meta-regression-of-textrmlnhatsigma-with-textrmlnoverlinex-and-group}{%
\subsection{\texorpdfstring{Meta-regression of \(\textrm{ln}\hat\sigma\) with \(\textrm{ln}\overline{x}\) and \(Group\)}{Meta-regression of \textbackslash textrm\{ln\}\textbackslash hat\textbackslash sigma with \textbackslash textrm\{ln\}\textbackslash overline\{x\} and Group}}\label{meta-regression-of-textrmlnhatsigma-with-textrmlnoverlinex-and-group}}

Instead, we can use a version of the meta-regression model described above (see equation @ref(eq:lnm-lns-model-eq) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using \(\textrm{ln}\hat\sigma\) and \(\textrm{ln}\overline{x}\). In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes \(\beta_{2}Group\) with \(Group\) as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where \(\beta_{2}\) is the slope or regression coefficient for \(Group\). \textcolor{red}{This kind of meta-regression model with  $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups. This reflects the pairwise nature of the $\textrm{ln}CVR$.}

\begin{figure}

{\centering \includegraphics{Revised-Paper---Meta-Analysis-of-Variation-in-Sport-and-Exercise-Science---anonymous_files/figure-latex/model-mean-variance-delta-plot-1}

}

\caption{Meta-analytic scatter plot of the log mean and log standard deviation of change scores.}(\#fig:model-mean-variance-delta-plot)
\end{figure}

Given the heteroskedasticity in the change scores means and standard deviations (see figure @ref(fig:mean-variance-delta-plot)), we fit this model and other model specifications\footnote{Note, as with the models examining \(Outcome\) upon baseline scores, we similarly explored \(\textrm{ln}\hat\sigma\) with \(\textrm{ln}\overline{x}\) and \(Group\) as a predictor with \textcolor{red}{(1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).  The comparison of these models using $2\times\textrm{log}BF$} (Kass \& Raftery, 1995) \textcolor{red}{from approximate Bayesian information criterion} (Wagenmakers, 2007) \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (see https://osf.io/b5deh for strength and https://osf.io/5ektd for hypertrophy). Similar to the models including $Outcome$ there was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here. All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (https://osf.io/5g7ce) all of which revealed similar conclusions.}}\textsuperscript{,}\footnote{\textcolor{red}{It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variance relationship. For example, the $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; in the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one. In both however this is a *strong* assumption. The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in model where random slopes are included this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others). Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression models allow). It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case. Given that for most effects that might be included in such models we can determine the sampling variance for $\overline{x}$ one approach to address this might be to employ models that incorporate the variance on this predictor (i.e., measurement error or errors in variables models). This is beyond the scope of this paper to discuss. It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.}} to the dataset (see figure @ref(fig:model-mean-variance-delta-plot)). The results were largely similar to those found using the \(\textrm{ln}CVR\) model for strength (\(\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}\) = -0.34 {[}95\%CI: -0.54 to -0.13{]}) and hypertrophy (\(\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}\) = -0.47 {[}95\%CI: -0.71 to -0.22{]}). Both models suggest that the introduction of the RT interventions \emph{reduced} variance in change scores.

\hypertarget{moderators-of-variance-effects}{%
\subsection{Moderators of variance effects}\label{moderators-of-variance-effects}}

\textcolor{red}{Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework. Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)}\footnote{See supplementary materials (\url{https://osf.io/e6vpr}) for examples from model estimates for both SMD and \(\textrm{ln}CVR\), (used for simplicity of presenting moderator analysis results) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes. There were no obvious moderators of \(\textrm{ln}CVR\) in particular.}. \textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size. The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis. Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means\footnote{Indeed, it can be seen from figures @ref(fig:forest-SDir-plot), @ref(fig:forest-logVR-plot), and @ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.} (Yang et al., 2022).

It is of particular interest to note the different conclusions drawn here are dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is \emph{detectable} inter-individual response variation present. \textcolor{red}{We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications. Given the different conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of simpler approaches are met, or indeed seem reasonable, before their application in specific situations. Where for example there is not an obvious mean-variance relationship it may be appropriate to utilise the $SD_{ir}$ or $\textrm{ln}VR$. However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate, though assumptions regarding whether this relationship is proportional and fixed should also be considered. The structure of the data also impacts the specific modelling approach to be employed and whether or not assumptions that the mean-variance relationship varies within study or arm levels can be incorporated. Further, depending on the exact research questions it is worth considering balancing model complexity with its ability to provide an answer. If a simple pairwise comparison across a categorical variable is of interest then this can be explored with comparable models using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$. But where more complicated predictors, including categories extending beyond just two, are of interest then the flexibility of the latter is desirable to explore. Lastly, it is possible that substantive conclusions might be sensitive to the exact model specifications used. As such, it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis. With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions} (Olsson-Collentine et al., 2023; Steegen et al., 2016).

The examples presented herein used data from RT studies included in a recent meta-analysis published in the \emph{Journal of Sport Sciences} (Polito et al., 2021), which hopefully makes them more relatable for researchers in sport and exercise sciences. Examining this data using absolute comparisons of variance such as \(SD_{ir}\) and \(\textrm{ln}VR\) gave the impression that the introduction of the RT intervention likely \emph{increased} variance above random error, suggesting the presence of inter-individual response variation. In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive (as seen in the SMD and \(\textrm{ln}RR\) models), yet combined with the results from \(SD_{ir}\) and \(\textrm{ln}VR\) we might conclude that while all likely benefit, some benefit more so than others. Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some (Usui et al., 2021). Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.

However, similar to the cross-sectional pre-intervention score models reported here (and indeed most physical and biological variables), change scores also demonstrated a mean-variance relationship in addition to heteroskedasticity. The likely more appropriate analyses in this case using the \(\textrm{ln}CVR\) or meta-regression of \(\textrm{ln}\hat\sigma\) upon \(\textrm{ln}\overline{x}\) revealed conclusions in the opposite substantive direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly \emph{decreased} change score variance, implying that there is likely little to no interindividual response variation to explain. Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve (Usui et al., 2021); that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect (Cortés Martínez, 2021). Interventions such as these are valuable for the simplification of guidelines and recommendations. For example, muscle strengthening interventions such as RT are recommended for \emph{everyone} in current physical activity guidelines and in such applications there is likely value in a simple approach to such recommendations (Steele et al., 2017, 2022).

The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis. Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; Halliday et al. (2017)), or a ceiling effect on change (i.e., plateau in response; Steele et al. (2022)) has a constraining effect (Cortés Martínez, 2021). However, this potentially represents another interesting area of future study regarding variation; specifically, how to produce interventions that actually reduce variance in an outcome. In other contexts such as sporting performance, interventions to not only positively affect mean performance but also those that reduce variation in performance would be highly desirable.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Embracing variability and focusing on more than merely the mean differences between groups or conditions, such as intervention and control comparisons, has the potential to inform experimental design and lead to changes in both the approach and direction of follow-up studies. Whether there is evidence of meaningful average intervention effects or not, where considerable variance effects are present it suggests that a meaningful line of research would be to aim at identifying subgroup- or participant-by-intervention interactions using appropriate study designs (Hecksteden et al., 2015). Where variance effects are limited this instead suggests that translational work towards generalisable implementation might be the most meaningful line of future research. Finally, there may be cases where it is in fact desirable to identify interventions that actually reduce variance; for example, improvements in methodological approaches to enhance research (Usui et al., 2021), or interventions to reduce variation in sport performances. Thus, researchers in sport and exercise science should consider exploring variance more systematically, and indeed utilise the meta-analytic framework to support this. This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches (Atkinson et al., 2019; Atkinson \& Batterham, 2015; Hopkins, 2015; Mills et al., 2021; Nakagawa et al., 2015; Usui et al., 2021) presented here and echoing the efforts of other recent work (Bonafiglia et al., 2022; Esteves et al., 2021; Fisher et al., 2022; Kelley et al., 2020, 2022; Steele et al., 2021). \textcolor{red}{Researchers should be aware though that the meta-analytic modelling of variance, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make. Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.}

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-atkinsonTrueFalseInterindividual2015}{}}%
Atkinson, G., \& Batterham, A. M. (2015). True and false interindividual differences in the physiological response to an intervention. \emph{Experimental Physiology}, \emph{100}(6), 577--588. \url{https://doi.org/10.1113/EP085070}

\leavevmode\vadjust pre{\hypertarget{ref-atkinsonIssuesDeterminationResponders2019}{}}%
Atkinson, G., Williamson, P., \& Batterham, A. M. (2019). Issues in the determination of 'responders' and 'non-responders' in physiological research. \emph{Experimental Physiology}, \emph{104}(8), 1215--1225. \url{https://doi.org/10.1113/EP087712}

\leavevmode\vadjust pre{\hypertarget{ref-barrRandomEffectsStructure2013}{}}%
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: {Keep} it maximal. \emph{Journal of Memory and Language}, \emph{68}(3), 10.1016/j.jml.2012.11.001. \url{https://doi.org/10.1016/j.jml.2012.11.001}

\leavevmode\vadjust pre{\hypertarget{ref-bernardez-vazquezResistanceTrainingVariables2022}{}}%
Bernárdez-Vázquez, R., Raya-González, J., Castillo, D., \& Beato, M. (2022). Resistance {Training} {Variables} for {Optimization} of {Muscle} {Hypertrophy}: {An} {Umbrella} {Review}. \emph{Frontiers in Sports and Active Living}, \emph{4}. \url{https://www.frontiersin.org/articles/10.3389/fspor.2022.949021}

\leavevmode\vadjust pre{\hypertarget{ref-bonafigliaInterindividualDifferencesTrainability2022}{}}%
Bonafiglia, J. T., Swinton, P. A., Ross, R., Johannsen, N. M., Martin, C. K., Church, T. S., Slentz, C. A., Ross, L. M., Kraus, W. E., Walsh, J. J., Kenny, G. P., Goldfield, G. S., Prud'homme, D., Sigal, R. J., Earnest, C. P., \& Gurd, B. J. (2022). Interindividual {Differences} in {Trainability} and {Moderators} of {Cardiorespiratory} {Fitness}, {Waist} {Circumference}, and {Body} {Mass} {Responses}: {A} {Large}-{Scale} {Individual} {Participant} {Data} {Meta}-analysis. \emph{Sports Medicine (Auckland, N.Z.)}. \url{https://doi.org/10.1007/s40279-022-01725-9}

\leavevmode\vadjust pre{\hypertarget{ref-borensteinBasicIntroductionFixedeffect2010}{}}%
Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2010). A basic introduction to fixed-effect and random-effects models for meta-analysis. \emph{Research Synthesis Methods}, \emph{1}(2), 97--111. \url{https://doi.org/10.1002/jrsm.12}

\leavevmode\vadjust pre{\hypertarget{ref-borensteinIntroductionMetaAnalysis2021}{}}%
Borenstein, M., Hedges, L. V., Higgins, J. P. T., \& Rothstein, H. R. (2021). \emph{Introduction to {Meta}-{Analysis}}. John Wiley \& Sons.

\leavevmode\vadjust pre{\hypertarget{ref-caldwellCaseDefaultEffect2020}{}}%
Caldwell, A., \& Vigotsky, A. D. (2020). A case against default effect sizes in sport and exercise science. \emph{PeerJ}, \emph{8}, e10314. \url{https://doi.org/10.7717/peerj.10314}

\leavevmode\vadjust pre{\hypertarget{ref-carpinelliInterindividualHeterogeneityAdaptations2017}{}}%
Carpinelli, R. N. (2017). Interindividual heterogeneity of adaptations to resistance training. \emph{Medicina Sportiva Practica}, \emph{18}(4), 79--94.

\leavevmode\vadjust pre{\hypertarget{ref-cohenStatisticalPowerAnalysis1988}{}}%
Cohen, J. (1988). \emph{Statistical {Power} {Analysis} for the {Behavioral} {Sciences}} (2nd ed.). Routledge. \url{https://doi.org/10.4324/9780203771587}

\leavevmode\vadjust pre{\hypertarget{ref-cortesmartinezConstantEffectRandomized2021}{}}%
Cortés Martínez, J. (2021). Constant effect in randomized clinical trials with quantitative outcome : A methodological review. \emph{TDX (Tesis Doctorals En Xarxa)}. \url{https://upcommons.upc.edu/handle/2117/349575}

\leavevmode\vadjust pre{\hypertarget{ref-curtisMetaanalysisElevatedCO21998}{}}%
Curtis, P. S., \& Wang, X. (1998). A meta-analysis of elevated {CO2} effects on woody plant mass, form, and physiology. \emph{Oecologia}, \emph{113}(3), 299--313. \url{https://doi.org/10.1007/s004420050381}

\leavevmode\vadjust pre{\hypertarget{ref-debQuantifyingEffectsAcute2018}{}}%
Deb, S. K., Brown, D. R., Gough, L. A., Mclellan, C. P., Swinton, P. A., Andy Sparks, S., \& Mcnaughton, L. R. (2018). Quantifying the effects of acute hypoxic exposure on exercise performance and capacity: {A} systematic review and meta-regression. \emph{European Journal of Sport Science}, \emph{18}(2), 243--256. \url{https://doi.org/10.1080/17461391.2017.1410233}

\leavevmode\vadjust pre{\hypertarget{ref-estevesIndividualParticipantData2021}{}}%
Esteves, G. P., Swinton, P., Sale, C., James, R. M., Artioli, G. G., Roschel, H., Gualano, B., Saunders, B., \& Dolan, E. (2021). Individual {Participant} {Data} {Meta}-{Analysis} {Provides} {No} {Evidence} of {Intervention} {Response} {Variation} in {Individuals} {Supplementing} {With} {Beta}-{Alanine}. \emph{International Journal of Sport Nutrition and Exercise Metabolism}, \emph{31}(4), 305--313. \url{https://doi.org/10.1123/ijsnem.2021-0038}

\leavevmode\vadjust pre{\hypertarget{ref-exnerDoesPerformingResistance2022}{}}%
Exner, R. J., Patel, M. H., Whitener, D. V., Buckner, S. L., Jessee, M. B., \& Dankel, S. J. (2022). Does performing resistance exercise to failure homogenize the training stimulus by accounting for differences in local muscular endurance? \emph{European Journal of Sport Science}, 1--10. \url{https://doi.org/10.1080/17461391.2021.2023657}

\leavevmode\vadjust pre{\hypertarget{ref-fisherRoleSupervisionResistance2022}{}}%
Fisher, J., Steele, J., Wolf, M., Korakakis, P. A., Smith, D., \& Giessing, J. (2022). The {Role} of {Supervision} in {Resistance} {Training}; an {Exploratory} {Systematic} {Review} and {Meta}-{Analysis}: \emph{International Journal of Strength and Conditioning}, \emph{2}(1). \url{https://doi.org/10.47206/ijsc.v2i1.101}

\leavevmode\vadjust pre{\hypertarget{ref-glassPrimarySecondaryMetaAnalysis1976}{}}%
Glass, G. V. (1976). Primary, {Secondary}, and {Meta}-{Analysis} of {Research}. \emph{Educational Researcher}, \emph{5}(10), 3--8. \url{https://doi.org/10.2307/1174772}

\leavevmode\vadjust pre{\hypertarget{ref-gurevitchInteractionCompetitionPredation2000}{}}%
Gurevitch, J., Morrison, J. A., \& Hedges, L. V. (2000). The {Interaction} between {Competition} and {Predation}: {A} {Meta}-analysis of {Field} {Experiments}. \emph{The American Naturalist}, \emph{155}(4), 435--453. \url{https://doi.org/10.1086/303337}

\leavevmode\vadjust pre{\hypertarget{ref-haggerMetaanalysisSportExercise2006}{}}%
Hagger, M. (2006). Meta-analysis in sport and exercise research: {Review}, recent developments, and recommendations. \emph{European Journal of Sport Science}, \emph{6}(2), 103--115. \url{https://doi.org/10.1080/17461390500528527}

\leavevmode\vadjust pre{\hypertarget{ref-haggerMetaanalysis2022}{}}%
Hagger, M. (2022). Meta-analysis. \emph{International Review of Sport and Exercise Psychology}, \emph{15}(1), 120--151. \url{https://doi.org/10.1080/1750984X.2021.1966824}

\leavevmode\vadjust pre{\hypertarget{ref-hallidayResistanceTrainingAssociated2017}{}}%
Halliday, T. M., Savla, J., Marinik, E. L., Hedrick, V. E., Winett, R. A., \& Davy, B. M. (2017). Resistance training is associated with spontaneous changes in aerobic physical activity but not overall diet quality in adults with prediabetes. \emph{Physiology \& Behavior}, \emph{177}, 49--56. \url{https://doi.org/10.1016/j.physbeh.2017.04.013}

\leavevmode\vadjust pre{\hypertarget{ref-halperinAccuracyPredictingRepetitions2022}{}}%
Halperin, I., Malleron, T., Har-Nir, I., Androulakis-Korakakis, P., Wolf, M., Fisher, J., \& Steele, J. (2022). Accuracy in {Predicting} {Repetitions} to {Task} {Failure} in {Resistance} {Exercise}: {A} {Scoping} {Review} and {Exploratory} {Meta}-analysis. \emph{Sports Medicine}, \emph{52}(2), 377--390. \url{https://doi.org/10.1007/s40279-021-01559-x}

\leavevmode\vadjust pre{\hypertarget{ref-heckstedenIndividualResponseExercise2015}{}}%
Hecksteden, A., Kraushaar, J., Scharhag-Rosenberger, F., Theisen, D., Senn, S., \& Meyer, T. (2015). Individual response to exercise training - a statistical perspective. \emph{Journal of Applied Physiology (Bethesda, Md.: 1985)}, \emph{118}(12), 1450--1459. \url{https://doi.org/10.1152/japplphysiol.00714.2014}

\leavevmode\vadjust pre{\hypertarget{ref-hedgesMetaAnalysisResponseRatios1999}{}}%
Hedges, L. V., Gurevitch, J., \& Curtis, P. S. (1999). The {Meta}-{Analysis} of {Response} {Ratios} in {Experimental} {Ecology}. \emph{Ecology}, \emph{80}(4), 1150--1156. \url{https://doi.org/10.1890/0012-9658(1999)080\%5B1150:TMAORR\%5D2.0.CO;2}

\leavevmode\vadjust pre{\hypertarget{ref-hedgesSexDifferencesMental1995}{}}%
Hedges, L. V., \& Nowell, A. (1995). Sex differences in mental test scores, variability, and numbers of high-scoring individuals. \emph{Science (New York, N.Y.)}, \emph{269}(5220), 41--45. \url{https://doi.org/10.1126/science.7604277}

\leavevmode\vadjust pre{\hypertarget{ref-hedgesStatisticalMethodsMetaAnalysis2014}{}}%
Hedges, L. V., \& Olkin, I. (2014). \emph{Statistical {Methods} for {Meta}-{Analysis}}. Academic Press.

\leavevmode\vadjust pre{\hypertarget{ref-hedgesRobustVarianceEstimation2010}{}}%
Hedges, L. V., Tipton, E., \& Johnson, M. C. (2010). Robust variance estimation in meta-regression with dependent effect size estimates. \emph{Research Synthesis Methods}, \emph{1}(1), 39--65. \url{https://doi.org/10.1002/jrsm.5}

\leavevmode\vadjust pre{\hypertarget{ref-heidelMachinesFreeWeight2022}{}}%
Heidel, K. A., Novak, Z. J., \& Dankel, S. J. (2022). Machines and free weight exercises: A systematic review and meta-analysis comparing changes in muscle size, strength, and power. \emph{The Journal of Sports Medicine and Physical Fitness}, \emph{62}(8), 1061--1070. \url{https://doi.org/10.23736/S0022-4707.21.12929-9}

\leavevmode\vadjust pre{\hypertarget{ref-hopkinsIndividualResponsesMade2015}{}}%
Hopkins, W. G. (2015). Individual responses made easy. \emph{Journal of Applied Physiology (Bethesda, Md.: 1985)}, \emph{118}(12), 1444--1446. \url{https://doi.org/10.1152/japplphysiol.00098.2015}

\leavevmode\vadjust pre{\hypertarget{ref-hrubeniukDirectionsExerciseTreatment2022}{}}%
Hrubeniuk, T. J., Bonafiglia, J. T., Bouchard, D. R., Gurd, B. J., \& Sénéchal, M. (2022). Directions for {Exercise} {Treatment} {Response} {Heterogeneity} and {Individual} {Response} {Research}. \emph{International Journal of Sports Medicine}, \emph{43}(1), 11--22. \url{https://doi.org/10.1055/a-1548-7026}

\leavevmode\vadjust pre{\hypertarget{ref-kassBayesFactors1995}{}}%
Kass, R. E., \& Raftery, A. E. (1995). Bayes {Factors}. \emph{Journal of the American Statistical Association}, \emph{90}(430), 773--795. \url{https://doi.org/10.1080/01621459.1995.10476572}

\leavevmode\vadjust pre{\hypertarget{ref-kelleyPrecisionExerciseMedicine2022}{}}%
Kelley, G. A. (2022). Precision exercise medicine in rheumatology: {Don}'t put the cart before the horse. \emph{Clinical Rheumatology}, \emph{41}(8), 2277--2279. \url{https://doi.org/10.1007/s10067-022-06260-6}

\leavevmode\vadjust pre{\hypertarget{ref-kelleyAreThereInterindividual2022}{}}%
Kelley, G. A., Kelley, K. S., \& Callahan, L. F. (2022). Are {There} {Interindividual} {Differences} in {Anxiety} as a {Result} of {Aerobic} {Exercise} {Training} in {Adults} with {Fibromyalgia}? {An} {Ancillary} {Meta}-analysis of {Randomized} {Controlled} {Trials}. \emph{Archives of Physical Medicine and Rehabilitation}, S0003-9993(22)00007-7. \url{https://doi.org/10.1016/j.apmr.2021.12.019}

\leavevmode\vadjust pre{\hypertarget{ref-kelleyAreThereInterIndividual2020}{}}%
Kelley, G. A., Kelley, K. S., \& Pate, R. R. (2020). Are {There} {Inter}-{Individual} {Differences} in {Fat} {Mass} and {Percent} {Body} {Fat} as a {Result} of {Aerobic} {Exercise} {Training} in {Overweight} and {Obese} {Children} and {Adolescents}? {A} {Meta}-{Analytic} {Perspective}. \emph{Childhood Obesity (Print)}, \emph{16}(5), 301--306. \url{https://doi.org/10.1089/chi.2020.0056}

\leavevmode\vadjust pre{\hypertarget{ref-lajeunesseMetaanalysisResponseRatios2011}{}}%
Lajeunesse, M. J. (2011). On the meta-analysis of response ratios for studies with correlated and multi-group designs. \emph{Ecology}, \emph{92}(11), 2049--2055. \url{https://doi.org/10.1890/11-0423.1}

\leavevmode\vadjust pre{\hypertarget{ref-lajeunesseBiasCorrectionLog2015}{}}%
Lajeunesse, M. J. (2015). Bias and correction for the log response ratio in ecological meta-analysis. \emph{Ecology}, \emph{96}(8), 2056--2063. \url{https://doi.org/10.1890/14-2402.1}

\leavevmode\vadjust pre{\hypertarget{ref-millsDetectingHeterogeneityIntervention2021}{}}%
Mills, H. L., Higgins, J. P. T., Morris, R. W., Kessler, D., Heron, J., Wiles, N., Davey Smith, G., \& Tilling, K. (2021). Detecting {Heterogeneity} of {Intervention} {Effects} {Using} {Analysis} and {Meta}-analysis of {Differences} in {Variance} {Between} {Trial} {Arms}. \emph{Epidemiology (Cambridge, Mass.)}, \emph{32}(6), 846--854. \url{https://doi.org/10.1097/EDE.0000000000001401}

\leavevmode\vadjust pre{\hypertarget{ref-morrisEstimatingEffectSizes2008}{}}%
Morris, S. B. (2008). Estimating {Effect} {Sizes} {From} {Pretest}-{Posttest}-{Control} {Group} {Designs}. \emph{Organizational Research Methods}, \emph{11}(2), 364--386. \url{https://doi.org/10.1177/1094428106291059}

\leavevmode\vadjust pre{\hypertarget{ref-morrisDirectInteractiveEffects2007}{}}%
Morris, W. F., Hufbauer, R. A., Agrawal, A. A., Bever, J. D., Borowicz, V. A., Gilbert, G. S., Maron, J. L., Mitchell, C. E., Parker, I. M., Power, A. G., Torchin, M. E., \& Vázquez, D. P. (2007). Direct and interactive effects of enemies and mutualists on plant performance: A meta-analysis. \emph{Ecology}, \emph{88}(4), 1021--1029. \url{https://doi.org/10.1890/06-0442}

\leavevmode\vadjust pre{\hypertarget{ref-nakagawaEffectSizeConfidence2007}{}}%
Nakagawa, S., \& Cuthill, I. C. (2007). Effect size, confidence interval and statistical significance: A practical guide for biologists. \emph{Biological Reviews of the Cambridge Philosophical Society}, \emph{82}(4), 591--605. \url{https://doi.org/10.1111/j.1469-185X.2007.00027.x}

\leavevmode\vadjust pre{\hypertarget{ref-nakagawaMetaanalysisVariationEcological2015}{}}%
Nakagawa, S., Poulin, R., Mengersen, K., Reinhold, K., Engqvist, L., Lagisz, M., \& Senior, A. M. (2015). Meta-analysis of variation: Ecological and evolutionary applications and beyond. \emph{Methods in Ecology and Evolution}, \emph{6}(2), 143--152. \url{https://doi.org/10.1111/2041-210X.12309}

\leavevmode\vadjust pre{\hypertarget{ref-nakagawaMeanStrikesBack2012}{}}%
Nakagawa, S., \& Schielzeth, H. (2012). The mean strikes back: Mean-variance relationships and heteroscedasticity. \emph{Trends in Ecology \& Evolution}, \emph{27}(9), 474-475; author reply 475-476. \url{https://doi.org/10.1016/j.tree.2012.04.003}

\leavevmode\vadjust pre{\hypertarget{ref-nuzzoEccentricConcentricStrength2023}{}}%
Nuzzo, J., Pinto, M. D., Nosaka, K., \& Steele, J. (2023a). The {Eccentric}:{Concentric} {Strength} {Ratio} of {Human} {Skeletal} {Muscle} {In} {Vivo}: {Meta}-analysis of the {Influences} of {Sex}, {Age}, {Joint} {Action}, and {Velocity}. \emph{Sports Medicine}, \emph{53}(6), 1125--1136. \url{https://doi.org/10.1007/s40279-023-01851-y}

\leavevmode\vadjust pre{\hypertarget{ref-nuzzoMaximalNumberRepetitions2023}{}}%
Nuzzo, J., Pinto, M., Nosaka, K., \& Steele, J. (2023b). \emph{Maximal number of repetitions at percentages of the one repetition maximum: {A} meta-regression and moderator analysis of sex, age, training status, and exercise}. SportRxiv. \url{https://doi.org/10.51224/SRXIV.291}

\leavevmode\vadjust pre{\hypertarget{ref-olsson-collentineMetaanalyzingMultiversePeek2023}{}}%
Olsson-Collentine, A., Van Aert, R. C. M., Bakker, M., \& Wicherts, J. (2023). Meta-analyzing the multiverse: {A} peek under the hood of selective reporting. \emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000559}

\leavevmode\vadjust pre{\hypertarget{ref-pickeringNonRespondersExerciseExistand2019}{}}%
Pickering, C., \& Kiely, J. (2019). Do {Non}-{Responders} to {Exercise} {Exist}-and {If} {So}, {What} {Should} {We} {Do} {About} {Them}? \emph{Sports Medicine (Auckland, N.Z.)}, \emph{49}(1), 1--7. \url{https://doi.org/10.1007/s40279-018-01041-1}

\leavevmode\vadjust pre{\hypertarget{ref-plackettStudiesHistoryProbability1958}{}}%
Plackett, R. L. (1958). Studies in the {History} of {Probability} and {Statistics}: {VII}. {The} {Principle} of the {Arithmetic} {Mean}. \emph{Biometrika}, \emph{45}(1/2), 130--135. \url{https://doi.org/10.2307/2333051}

\leavevmode\vadjust pre{\hypertarget{ref-politoModeratorsStrengthGains2021}{}}%
Polito, M. D., Papst, R. R., \& Farinatti, P. (2021). Moderators of strength gains and hypertrophy in resistance training: {A} systematic review and meta-analysis. \emph{Journal of Sports Sciences}, \emph{39}(19), 2189--2198. \url{https://doi.org/10.1080/02640414.2021.1924978}

\leavevmode\vadjust pre{\hypertarget{ref-rossPrecisionExerciseMedicine2019}{}}%
Ross, R., Goodpaster, B. H., Koch, L. G., Sarzynski, M. A., Kohrt, W. M., Johannsen, N. M., Skinner, J. S., Castro, A., Irving, B. A., Noland, R. C., Sparks, L. M., Spielmann, G., Day, A. G., Pitsch, W., Hopkins, W. G., \& Bouchard, C. (2019). Precision exercise medicine: Understanding exercise response variability. \emph{British Journal of Sports Medicine}, \emph{53}(18), 1141--1153. \url{https://doi.org/10.1136/bjsports-2018-100328}

\leavevmode\vadjust pre{\hypertarget{ref-seniorRevisitingExpandingMetaanalysis2020}{}}%
Senior, A. M., Viechtbauer, W., \& Nakagawa, S. (2020). Revisiting and expanding the meta-analysis of variation: {The} log coefficient of variation ratio. \emph{Research Synthesis Methods}, \emph{11}(4), 553--567. \url{https://doi.org/10.1002/jrsm.1423}

\leavevmode\vadjust pre{\hypertarget{ref-steegenIncreasingTransparencyMultiverse2016}{}}%
Steegen, S., Tuerlinckx, F., Gelman, A., \& Vanpaemel, W. (2016). Increasing {Transparency} {Through} a {Multiverse} {Analysis}. \emph{Perspectives on Psychological Science}, \emph{11}(5), 702--712. \url{https://doi.org/10.1177/1745691616658637}

\leavevmode\vadjust pre{\hypertarget{ref-steeleLongTermTimeCourseStrength2022}{}}%
Steele, J., Fisher, J. P., Giessing, J., Androulakis-Korakakis, P., Wolf, M., Kroeske, B., \& Reuters, R. (2022). Long-{Term} {Time}-{Course} of {Strength} {Adaptation} to {Minimal} {Dose} {Resistance} {Training} {Through} {Retrospective} {Longitudinal} {Growth} {Modeling}. \emph{Research Quarterly for Exercise and Sport}, \emph{0}(0), 1--18. \url{https://doi.org/10.1080/02701367.2022.2070592}

\leavevmode\vadjust pre{\hypertarget{ref-steeleHigherEffortbasedParadigm2017}{}}%
Steele, J., Fisher, J., Skivington, M., Dunn, C., Arnold, J., Tew, G., Batterham, A. M., Nunan, D., O'Driscoll, J. M., Mann, S., Beedie, C., Jobson, S., Smith, D., Vigotsky, A., Phillips, S., Estabrooks, P., \& Winett, R. (2017). A higher effort-based paradigm in physical activity and exercise for public health: Making the case for a greater emphasis on resistance training. \emph{BMC Public Health}, \emph{17}(1), 300. \url{https://doi.org/10.1186/s12889-017-4209-8}

\leavevmode\vadjust pre{\hypertarget{ref-steeleSlowSteadyHard2021}{}}%
Steele, J., Plotkin, D., Van Every, D., Rosa, A., Zambrano, H., Mendelovits, B., Carrasquillo-Mercado, M., Grgic, J., \& Schoenfeld, B. J. (2021). Slow and {Steady}, or {Hard} and {Fast}? {A} {Systematic} {Review} and {Meta}-{Analysis} of {Studies} {Comparing} {Body} {Composition} {Changes} between {Interval} {Training} and {Moderate} {Intensity} {Continuous} {Training}. \emph{Sports (Basel, Switzerland)}, \emph{9}(11), 155. \url{https://doi.org/10.3390/sports9110155}

\leavevmode\vadjust pre{\hypertarget{ref-swintonStatisticalFrameworkInterpret2018}{}}%
Swinton, P. A., Hemingway, B. S., Saunders, B., Gualano, B., \& Dolan, E. (2018). A {Statistical} {Framework} to {Interpret} {Individual} {Response} to {Intervention}: {Paving} the {Way} for {Personalized} {Nutrition} and {Exercise} {Prescription}. \emph{Frontiers in Nutrition}, \emph{5}, 41. \url{https://doi.org/10.3389/fnut.2018.00041}

\leavevmode\vadjust pre{\hypertarget{ref-swintonInterpretingMagnitudeChange2022}{}}%
Swinton, P. A., Katherine, B., Andy, H., Leon, G., John, P., Rodrigo, R. A., Patrick, M., \& Andrew, M. (2022). Interpreting magnitude of change in strength and conditioning: {Effect} size selection, threshold values and {Bayesian} updating {Journal} of {Sports} {Sciences}. \emph{Journal of Sports Sciences}.

\leavevmode\vadjust pre{\hypertarget{ref-taylorAggregationVarianceMean1961}{}}%
Taylor, L. R. (1961). Aggregation, {Variance} and the {Mean}. \emph{Nature}, \emph{189}(4766), 732--735. \url{https://doi.org/10.1038/189732a0}

\leavevmode\vadjust pre{\hypertarget{ref-tenanCommentMethodStop2020}{}}%
Tenan, M. S., Vigotsky, A. D., \& Caldwell, A. R. (2020). Comment on: {``{A} {Method} to {Stop} {Analyzing} {Random} {Error} and {Start} {Analyzing} {Differential} {Responders} to {Exercise}.''} \emph{Sports Medicine}, \emph{50}(2), 431--434. \url{https://doi.org/10.1007/s40279-019-01249-9}

\leavevmode\vadjust pre{\hypertarget{ref-usuiMetaanalysisVariationSuggests2021}{}}%
Usui, T., Macleod, M. R., McCann, S. K., Senior, A. M., \& Nakagawa, S. (2021). Meta-analysis of variation suggests that embracing variability improves both replicability and generalizability in preclinical research. \emph{PLOS Biology}, \emph{19}(5), e3001009. \url{https://doi.org/10.1371/journal.pbio.3001009}

\leavevmode\vadjust pre{\hypertarget{ref-vandennoortgateThreelevelMetaanalysisDependent2013}{}}%
Van den Noortgate, W., López-López, J. A., Marín-Martínez, F., \& Sánchez-Meca, J. (2013). Three-level meta-analysis of dependent effect sizes. \emph{Behavior Research Methods}, \emph{45}(2), 576--594. \url{https://doi.org/10.3758/s13428-012-0261-6}

\leavevmode\vadjust pre{\hypertarget{ref-viechtbauerConductingMetaAnalysesMetafor2010}{}}%
Viechtbauer, W. (2010). Conducting {Meta}-{Analyses} in {R} with the metafor {Package}. \emph{Journal of Statistical Software}, \emph{36}, 1--48. \url{https://doi.org/10.18637/jss.v036.i03}

\leavevmode\vadjust pre{\hypertarget{ref-vigotskyImprobableDataPatterns2020}{}}%
Vigotsky, A., Nuckols, G. L., Fisher, J., Heathers, J., Krieger, J., Schoenfeld, B. J., Giessing, J., \& Steele, J. (2020). \emph{Improbable data patterns in the work of {Barbalho} et al.} SportRxiv. \url{https://doi.org/10.31236/osf.io/sg3wm}

\leavevmode\vadjust pre{\hypertarget{ref-wagenmakersPracticalSolutionPervasive2007}{}}%
Wagenmakers, E.-J. (2007). A practical solution to the pervasive problems ofp values. \emph{Psychonomic Bulletin \& Review}, \emph{14}(5), 779--804. \url{https://doi.org/10.3758/BF03194105}

\leavevmode\vadjust pre{\hypertarget{ref-wolfPartialVsFull2023}{}}%
Wolf, M., Androulakis-Korakakis, P., Fisher, J., Schoenfeld, B., \& Steele, J. (2023). Partial {Vs} {Full} {Range} of {Motion} {Resistance} {Training}: {A} {Systematic} {Review} and {Meta}-{Analysis}. \emph{International Journal of Strength and Conditioning}, \emph{3}(1). \url{https://doi.org/10.47206/ijsc.v3i1.182}

\leavevmode\vadjust pre{\hypertarget{ref-yangLowStatisticalPower2022}{}}%
Yang, Y., Hillebrand, H., Lagisz, M., Cleasby, I., \& Nakagawa, S. (2022). Low statistical power and overestimated anthropogenic impacts, exacerbated by publication bias, dominate field studies in global change biology. \emph{Global Change Biology}, \emph{28}(3), 969--989. \url{https://doi.org/10.1111/gcb.15972}

\end{CSLReferences}

\end{document}

### 2023-09-14 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -971,9 +971,6 @@ MultiLevelModel_ri_group_slope_study_log_mean_variance <- rma.mv(SD_log, V=SD_lo
 
 save(MultiLevelModel_ri_group_slope_study_log_mean_variance, file = "models/MultiLevelModel_ri_group_slope_study_log_mean_variance")
 
-### Calculate I^2 
-I2_rs_study__log_mean_variance <- i2_ml(MultiLevelModel_ri_group_slope_study_log_mean_variance)
-
 ### Calculate robust estimate from multi-level model
 RobuEstMultiLevelModel_ri_group_slope_study_log_mean_variance <- robust(MultiLevelModel_ri_group_slope_study_log_mean_variance, Data_long_pre$study)
 
@@ -989,9 +986,6 @@ MultiLevelModel_ri_group_slope_study_arm_log_mean_variance <- rma.mv(SD_log, V=S
                                             )
 
 save(MultiLevelModel_ri_group_slope_study_arm_log_mean_variance, file = "models/MultiLevelModel_ri_group_slope_study_arm_log_mean_variance")
-
-### Calculate I^2 
-I2_rs_both__log_mean_variance <- i2_ml(MultiLevelModel_ri_group_slope_study_arm_log_mean_variance)
 
 ### Calculate robust estimate from multi-level model
 RobuEstMultiLevelModel_ri_group_slope_study_arm_log_mean_variance <- robust(MultiLevelModel_ri_group_slope_study_arm_log_mean_variance, Data_long_pre$study)
@@ -1009,9 +1003,6 @@ MultiLevelModel_ri_mean_slope_study_log_mean_variance <- rma.mv(SD_log, V=SD_log
 
 save(MultiLevelModel_ri_mean_slope_study_log_mean_variance, file = "models/MultiLevelModel_ri_mean_slope_study_log_mean_variance")
 
-### Calculate I^2 
-I2_rs_both__log_mean_variance <- i2_ml(MultiLevelModel_ri_mean_slope_study_log_mean_variance)
-
 ### Calculate robust estimate from multi-level model
 RobuEstMultiLevelModel_ri_mean_slope_study_log_mean_variance <- robust(MultiLevelModel_ri_mean_slope_study_log_mean_variance, Data_long_pre$study)
 
@@ -1027,9 +1018,6 @@ MultiLevelModel_ri_mean_slope_study_arm_log_mean_variance <- rma.mv(SD_log, V=SD
 )
 
[…]

### 2023-09-15 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Render revised paper for submission with author names

--- 2023-09-14 paper/Revised Pre-print - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
@@
 thanks: |
-  Preprint, please cite as: Steele, J., Fisher, J., Smith, D., Schoenfeld, B., Yang, R., & Nakagawa, S.
-(2022).
-Meta-Analysis of Variation in Sport and Exercise Science: Examples of Application Within Resistance Training Research.
-$\textrm{sportr}\chi\textrm{iv}$ DOI: https://doi.org/10.51224/SRXIV.214.
   Address for correspondence: james.steele@solent.ac.uk

### 2023-09-15 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Add track changes to revised named paper

@@
    - \captionsetup[figure]{font=small}
----
-
+   - \usepackage{lineno}
+   - \linenumbers
 ---
@@
 Accordingly, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
-However, to begin we provide a conceptual overview of meta-analyses and effect sizes as typically used within sport and exercise science.
+\textcolor{red}{However, to begin we provide a conceptual overview of meta-analyses and effect sizes as typically used within sport and exercise science.}
 
@@
 
-There are two popular models[^1] used for meta-analysis: the fixed-effect model and the random-effects model [@borensteinBasicIntroductionFixedeffect2010].
-The fixed effect model assumes that there is one true effect size[^2] that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error.
+\textcolor{red}{There are two popular models}[^1] \textcolor{red}{used for meta-analysis: the fixed-effect model and the random-effects model} [@borensteinBasicIntroductionFixedeffect2010].
+\textcolor{red}{The fixed effect model assumes that there is one true effect size}[^2] \textcolor{red}{that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error.
 This essentially means that there is a single common effect which is fixed across studies and each study takes samples of individuals from the population to estimate this effect.
-We can express this model in the following formula:
+We can express this model in the following formula:}
 
-[^1]: To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics.
+[^1]: \textcolor{red}{To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics.
 A statistical model essentially is specification of what we think the data generating process might be for a given situation.
 In the context of meta-analyses the data are usually the individual effects that we have extracted from studies i.e., the results of each study.
-The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.
+The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.}
 
-[^2]: *Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
-This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between [@caldwellCaseDefaultEffect2020] including, as we shall see, statistics describing variation.
+[^2]: \textcolor{red}{*Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
+This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between} [@caldwellCaseDefaultEffect2020] \textcolor{red}{including, as we shall see, statistics describing variation.}
 
@@
 
-where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
+\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
 Studies with smaller standard errors of their effect estimates have smaller sampling errors and so these higher precision estimates are given greater weight in the model.
-The weighting given to each study is calculated as:
+The weighting given to each study is calculated as:}
 
@@
 
-Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
+\textcolor{red}{Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
 It is referred to as inverse-variance weighting.
-Then, the overall weighted mean effect estimate from the model is then calculated as:
+Then, the overall weighted mean effect estimate from the model is then calculated as:}
 
@@
 
-Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
+\textcolor{red}{Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
 Instead it allows for the true effects that each study estimates to differ.
@@
 Hence, in the fixed-effect model there is one effect and it is assumed to be *fixed* across studies, whereas in the random-effects model there are many and we examine an assumed *random* sample of them.
-We can express this model in the following formula:
+We can express this model in the following formula:}
 
@@
 
-where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
+\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
 Essentially this model assumes that each individual study estimates an effect and there will be some sampling error in estimating it.
 But, the effects each study estimates comes from an overall distribution of true effects with a mean value (i.e., $\theta_{i}$).
-The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated[^3].
-This essentially describes the heterogeneity between the effects included.
-Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:
+The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated}[^3].
+\textcolor{red}{This essentially describes the heterogeneity between the effects included.
+Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:}
 
-[^3]: This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
+[^3]: \textcolor{red}{This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
 This is beyond the scope of this paper to discuss.
-We note however that the models we present all utilise Restricted Maximum Likelihood estimation.
+We note however that the models we present all utilise Restricted Maximum Likelihood estimation.}
 
@@
 
-Then the average of the distribution of effects can be calculated as per equation \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.
+\textcolor{red}{Then the average of the distribution of effects can be calculated as per equation} \@ref(eq:fixed-average-eq) \textcolor{red}{substituting $w_{i}$ for $w^*_{i}$.}
 
-Although historically fixed effect models were commonplace in the field of sport and exercise science [@haggerMetaanalysisSportExercise2006] nowadays the random effects model is more often employed [@haggerMetaanalysis2022].
-This is likely due to the fact that direct (i.e., exact) replication of studies is rare[^4] and instead studies often explore similar effects across varying moderating factors such as those noted above.
+\textcolor{red}{Although historically fixed effect models were commonplace in the field of sport and exercise science} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays the random effects model is more often employed} [@haggerMetaanalysis2022].
+\textcolor{red}{This is likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above.}
 
-[^4]: Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).
+[^4]: \textcolor{red}{Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).}
 
-As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable.
-The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
+As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] \textcolor{red}{likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable.
+The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping}; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
 Indeed, a recent umbrella review [@bernardez-vazquezResistanceTrainingVariables2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes, all of which focused on the comparison of mean changes between different intervention groups.
 
-Often, due to the varying operationalisations used for broad outcome concepts[^5], an effect size is used which is "standardised" across studies to allow for their synthesis.
-Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
+\textcolor{red}{Often, due to the varying operationalisations used for broad outcome concepts}[^5]\textcolor{red}{, an effect size is used which is "standardised" across studies to allow for their synthesis.} Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
 This statistic is usually a version of Cohen's *d* [@cohenStatisticalPowerAnalysis1988], or its bias-corrected[^7] metric referred to as Hedges' *g* [@hedgesStatisticalMethodsMetaAnalysis2014; @borensteinIntroductionMetaAnalysis2021; @nakagawaEffectSizeConfidence2007][^8].
@@
 
-[^5]: For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions.
-Or the same operationalisations may be employed but different exercises such as the squat or bench press.
+[^5]: \textcolor{red}{For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions.
+Or the same operationalisations may be employed but different exercises such as the squat or bench press.}
 
@@
 
-[^7]: For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct.
-Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.
+[^7]: \textcolor{red}{For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct.
+Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.}
 
 [^8]: We will refer to both merely as the SMD throughout the manuscript for simplicity and note that throughout when reporting a 'SMD' we are reporting the bias-corrected version.
-We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance; @caldwellCaseDefaultEffect2020).
+\textcolor{red}{We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance;} @caldwellCaseDefaultEffect2020).
 
@@
 For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawaMeanStrikesBack2012].
-Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] and in the number of repetitions that can be performed at different percentages of one repetition maximum [@nuzzoMaximalNumberRepetitions2023].
+Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] \textcolor{red}{and in the number of repetitions that can be performed at different percentages of one repetition maximum} [@nuzzoMaximalNumberRepetitions2023].
 
@@
 
-Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
+\textcolor{red}{Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
 We hope this will make clear why we offer such recommendations in our discussion and conclusion.
-As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.
+As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.}
 
@@
 The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present.
-In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.
-The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
+\textcolor{red}{In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.} The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
 Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), the $\textrm{ln}CVR$ is given by:
@@
 All analysis examples were performed in R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauerConductingMetaAnalysesMetafor2010].
-The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/) or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git).
+The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/) \textcolor{red}{or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git)}.
 
@@
 
-[^13]: Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
+[^13]: \textcolor{red}{Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
 This can be extended to meta-analytic synthesis where the independent variables reflect characteristics associated with the effects included.
-For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.
+For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.}
 
@@
 
-## Detecting the presence of interindividual response variation to resistance training intervention with absolute variance statistics
+## Detecting the presence of interindividual response variation to resistance training intervention \textcolor{red}{with absolute variance statistics}
 First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^14] meta-analysis to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^15].
@@
 However, the data we extracted were hierarchical in nature.
-Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.
-Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
-This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
-We can therefore describe the overall model as:
+\textcolor{red}{Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
+\textcolor{red}{This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
+We can therefore describe the overall model as:}
 
@@
 
-where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
-This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects[^17].
-The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).
+\textcolor{red}{where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
+This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects}[^17].
+\textcolor{red}{The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).}
 
-[^17]: Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.
+[^17]: \textcolor{red}{Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.}
 
@@
 As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%).
-Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).
+Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure }\@ref(fig:forest-SMD-plot)) \textcolor{red}{were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}
 
@@
 This additional variance might support previous perspectives [@carpinelliInterindividualHeterogeneityAdaptations2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* independent of the random error).
-However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean (i.e., that the variance is similar whether mean values are low or high).
+However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean \textcolor{red}{(i.e., that the variance is similar whether mean values are low or high)}.
 As we have seen from the SMD and $\textrm{ln}RR$ models, RT interventions increase mean scores.
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic.
-First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures to emphasise the need to consider this assumption, and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ and provide flexibility to accomodate wider applications.
+First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures \textcolor{red}{to emphasise the need to consider this assumption,} and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ \textcolor{red}{and provide flexibility to accomodate wider applications}.
 
@@
 As such, we can model variance of each effect as the response itself with the mean of the effect as the predictor.
-The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models (i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).
-Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
+The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models \textcolor{red}{(i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).} Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
 In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al.
@@
 As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
-The variance in standard deviations increases with higher mean values.
-This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
+\textcolor{red}{The variance in standard deviations increases with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
 
@@
 When this relationship holds, under most circumstances the standard deviation is not proportional to the mean.
-However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear based upon the product and power logarithmic rules:
+However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear \textcolor{red}{based upon the product and power logarithmic rules:}
 
@@
 
-These kinds of models are highly flexible.
-Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
+\textcolor{red}{These kinds of models are highly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
 
@@
 [^19]: We do not have to limit ourselves to only fixed effect predictor terms as we have here.
-Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes (i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences [@barrRandomEffectsStructure2013].
+Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences} [@barrRandomEffectsStructure2013].
 We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the overall average slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
-These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms.
+\textcolor{red}{These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms.
 We could also do the same and include random slopes for $(\beta_{1}$ on $\textrm{ln}\overline{x}$ thus allowing for the strength of the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ to also vary between studies and arms.
 Indeed, we fit a range of models  using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Outcome$ by study, (5) the inclusion of correlated random slopes for $Outcome$ by study and arm, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study, (7) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and arm.
-The comparison of these models using $2\times\textrm{log}BF$ [@kassBayesFactors1995] from approximate Bayesian information criterion [@wagenmakersPracticalSolutionPervasive2007] to determine under which is the observed data most likely is included in the supplementary materials (https://osf.io/3tv6x).
-There was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.
+The comparison of these models using $2\times\textrm{log}BF$} [@kassBayesFactors1995] \textcolor{red}{from approximate Bayesian information criterion} [@wagenmakersPracticalSolutionPervasive2007] \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (https://osf.io/3tv6x).
+There was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.}
 
@@
 Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
-Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet comparable approach might be desirable.
-Lastly, this statistic is limited to examination of pairwise comparisons of variance.
+Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet \textcolor{red}{comparable} approach might be desirable.
+\textcolor{red}{Lastly, this statistic is limited to examination of pairwise comparisons of variance.
 In this example this approach poses no issue as we are comparing RT intervention group(s) to a control group.
-But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a comparable effect of interest to the $\textrm{ln}CVR$ one just applied.
+But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a \textcolor{red}{comparable} effect of interest to the $\textrm{ln}CVR$ one just applied.}
 
@@
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-This kind of meta-regression model with  $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
-This reflects the pairwise nature of the $\textrm{ln}CVR$.
+\textcolor{red}{This kind of meta-regression model with  $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
+This reflects the pairwise nature of the $\textrm{ln}CVR$.}
 
@@
 
-[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).
-The comparison of these models using $2\times\textrm{log}BF$ [@kassBayesFactors1995] from approximate Bayesian information criterion [@wagenmakersPracticalSolutionPervasive2007] to determine under which is the observed data most likely is included in the supplementary materials (see https://osf.io/b5deh for strength and https://osf.io/5ektd for hypertrophy).
+[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with \textcolor{red}{(1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).
+The comparison of these models using $2\times\textrm{log}BF$} [@kassBayesFactors1995] \textcolor{red}{from approximate Bayesian information criterion} [@wagenmakersPracticalSolutionPervasive2007] \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (see https://osf.io/b5deh for strength and https://osf.io/5ektd for hypertrophy).
 Similar to the models including $Outcome$ there was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.
-All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (https://osf.io/5g7ce) all of which revealed similar conclusions.
+All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (https://osf.io/5g7ce) all of which revealed similar conclusions.}
 
-[^21]: It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variance relationship.
+[^21]: \textcolor{red}{It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variance relationship.
 For example, the $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; in the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
@@
 This is beyond the scope of this paper to discuss.
-It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.
+It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.}
 
 ## Moderators of variance effects
-Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework.
-Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)[^22].
-But these are limited to the pairwise comparisons of a categorical variable as the effect size.
-The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.
+\textcolor{red}{Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework.
+Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)}[^22].
+\textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size.
+The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}
 
@@
 It is of particular interest to note the different conclusions drawn here are dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
-We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
+\textcolor{red}{We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
 Given the different conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of simpler approaches are met, or indeed seem reasonable, before their application in specific situations.
@@
 As such, it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
-With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
+With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions} [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
 
@@
 This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawaMetaanalysisVariationEcological2015; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] presented here and echoing the efforts of other recent work [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
-Researchers should be aware though that the meta-analytic modelling of variance, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make.
-Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.
+\textcolor{red}{Researchers should be aware though that the meta-analytic modelling of variance, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make.
+Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.}
 

### 2023-11-13 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
Second round of revisions

@@
 subtitle: "Examples of Application Within Resistance Training Research"
-date: "14th September 2023"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 abstract: "Meta-analysis has become commonplace within sport and exercise science for synthesising and summarising empirical studies.
 However, most research in the field focuses upon mean effects; particularly the effects of interventions to improve outcomes such as fitness or performance.
-It is thought that \\textcolor{red}{individual responses to interventions vary considerably}.
+It is thought that individual responses to interventions vary considerably.
 Hence, interest has increased in exploring *precision* or *personalised* exercise approaches.
-Not only is the mean often affected by interventions, but variances may also be impacted.
-Exploration of variances in studies such as randomised controlled trials (RCTs) can yield insight into interindividual heterogeneity in response to interventions and help determine generalisability of effects.
-Yet, larger samples sizes than those used for typical mean effects are required when probing variances.
-Thus, \\textcolor{red}{in} a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
-\\textcolor{red}{Despite} the value of embracing and exploring variation alongside mean effects in sport and exercise science \\textcolor{red}{it is rarely applied to research synthesis through meta-analysis}.\\textcolor{red}{We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs}."
+Not only is the mean often affected by interventions, but variation may also be impacted.
+Exploration of variation in studies such as randomised controlled trials (RCTs) can yield insight into interindividual heterogeneity in response to interventions and help determine generalisability of effects.
+Yet, larger samples sizes than those used for typical mean effects are required when probing variation.
+Thus, in a field with small samples such as sport and exercise science, exploration of variation through a meta-analytic framework is appealing.
+Despite the value of embracing and exploring variation alongside mean effects in sport and exercise science it is rarely applied to research synthesis through meta-analysis.We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs."
 output: 
@@
 ```
+
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
@@
 ```
+
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
@@
 ```
+
 # Introduction
+
 Although the quantitative synthesis of results across studies has existed since the 17th century [@plackettStudiesHistoryProbability1958], the modern-day term "meta-analysis" was coined by Gene Glass [-@glassPrimarySecondaryMetaAnalysis1976].
@@
 
-\textcolor{red}{There are two popular models}[^1] \textcolor{red}{used for meta-analysis: the fixed-effect model and the random-effects model} [@borensteinBasicIntroductionFixedeffect2010].
-\textcolor{red}{The fixed effect model assumes that there is one true effect size}[^2] \textcolor{red}{that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error.
+There are two popular models[^1] used for meta-analysis: the fixed-effect model and the random-effects model [@borensteinBasicIntroductionFixedeffect2010].
+The fixed effect model assumes that there is one true effect size[^2] that each study included in the meta-analysis has estimated and that any differences in the estimates between individual studies are due to only sampling error.
 This essentially means that there is a single common effect which is fixed across studies and each study takes samples of individuals from the population to estimate this effect.
-We can express this model in the following formula:}
+We can express this model in the following formula:
 
-[^1]: \textcolor{red}{To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics.
+[^1]: To clarify language here for those unfamiliar, the term and concept *model* is used commonly in statistics.
 A statistical model essentially is specification of what we think the data generating process might be for a given situation.
 In the context of meta-analyses the data are usually the individual effects that we have extracted from studies i.e., the results of each study.
-The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.}
+The model, in mathematical formulae, is intended to approximate the processes that we assume led to the generation of the data.
 
-[^2]: \textcolor{red}{*Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
-This includes descriptive statistics ranging from mean raw values to correlation coeffients and everything in between} [@caldwellCaseDefaultEffect2020] \textcolor{red}{including, as we shall see, statistics describing variation.}
+[^2]: *Effect size* is an agnostic term used for a family of statistics which communicate the strength of a given 'effect' resulting from research.
+This includes descriptive statistics ranging from mean raw values to correlation coefficients and everything in between [@caldwellCaseDefaultEffect2020] including, as we shall see, statistics describing variation.
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
+```
+where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
+Studies with smaller standard errors of their effect estimates have smaller sampling errors and so these higher precision estimates are given greater weight in the model.
+The weighting given to each study is calculated as:
 
-\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
-Studies with smaller standard errors of their effect estimates have smaller sampling errors and so these higher precision estimates are given greater weight in the model.
-The weighting given to each study is calculated as:}
-
+```{=tex}
 \begin{equation}
@@
 \end{equation}
+```
+Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
+It is referred to as inverse-variance weighting.
+Then, the overall weighted mean effect estimate from the model is then calculated as:
 
-\textcolor{red}{Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
-It is referred to as inverse-variance weighting.
-Then, the overall weighted mean effect estimate from the model is then calculated as:}
-
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-\textcolor{red}{Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
+```
+Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
 Instead it allows for the true effects that each study estimates to differ.
@@
 Hence, in the fixed-effect model there is one effect and it is assumed to be *fixed* across studies, whereas in the random-effects model there are many and we examine an assumed *random* sample of them.
-We can express this model in the following formula:}
+We can express this model in the following formula:
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-\textcolor{red}{where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
+```
+where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
 Essentially this model assumes that each individual study estimates an effect and there will be some sampling error in estimating it.
 But, the effects each study estimates comes from an overall distribution of true effects with a mean value (i.e., $\theta_{i}$).
-The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated}[^3].
-\textcolor{red}{This essentially describes the heterogeneity between the effects included.
-Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:}
+The weighting in any model employing random effects such as this requires a different approach as the variance of the distribution of the effect sizes the model assumes, known as $\tau^2$, must also be estimated[^3].
+This essentially describes the heterogeneity between the effects included.
+Once this has been estimated we can calculate a weight that is adjusted for the random effects for each effect as:
 
-[^3]: \textcolor{red}{This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
+[^3]: This estimation can be done using a variety of methods and is an area of ongoing investigation as to how different methods perform.
 This is beyond the scope of this paper to discuss.
-We note however that the models we present all utilise Restricted Maximum Likelihood estimation.}
+We note however that the models we present all utilise Restricted Maximum Likelihood estimation.
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
+```
+Then the average of the distribution of effects can be calculated as per equation \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.
 
-\textcolor{red}{Then the average of the distribution of effects can be calculated as per equation} \@ref(eq:fixed-average-eq) \textcolor{red}{substituting $w_{i}$ for $w^*_{i}$.}
+Although historically fixed effect models were commonplace in the field of sport and exercise science [@haggerMetaanalysisSportExercise2006] nowadays the random effects model is more often employed [@haggerMetaanalysis2022].
+This is likely due to the fact that direct (i.e., exact) replication of studies is rare[^4] and instead studies often explore similar effects across varying moderating factors such as those noted above.
 
-\textcolor{red}{Although historically fixed effect models were commonplace in the field of sport and exercise science} [@haggerMetaanalysisSportExercise2006] \textcolor{red}{nowadays the random effects model is more often employed} [@haggerMetaanalysis2022].
-\textcolor{red}{This is likely due to the fact that direct (i.e., exact) replication of studies is rare}[^4] \textcolor{red}{and instead studies often explore similar effects across varying moderating factors such as those noted above.}
+[^4]: Hence current efforts to conduct direct replications (see <https://ssreplicationcentre.com/>).
 
-[^4]: \textcolor{red}{Hence current efforts to conduct direct replications (see https://ssreplicationcentre.com/).}
-
-As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] \textcolor{red}{likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the *effect* of an independent variable upon some dependent outcome variable.
-The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping}; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
+As with many other fields [@nakagawaMetaanalysisVariationEcological2015; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] likely the most common aim in meta-analysis in sport and exercise science, and indeed primary empirical research too, is to estimate the average *effect* of an independent variable upon some dependent outcome variable.
+The dependent variable is often the mean of a measurement and the independent variable is often a categorical grouping; for example, the comparison of an intervention group(s) and a control group, the comparison of intervention groups between one another, or comparison between non-manipulated categories such as biological sex.
 Indeed, a recent umbrella review [@bernardez-vazquezResistanceTrainingVariables2022] of meta-analyses in RT identified 14 studies examining the manipulation of RT intervention variables (i.e., the comparison of one intervention to another whereby a variable in the intervention was manipulated) on hypertrophy outcomes, all of which focused on the comparison of mean changes between different intervention groups.
 
-\textcolor{red}{Often, due to the varying operationalisations used for broad outcome concepts}[^5]\textcolor{red}{, an effect size is used which is "standardised" across studies to allow for their synthesis.} Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
-This statistic is usually a version of Cohen's *d* [@cohenStatisticalPowerAnalysis1988], or its bias-corrected[^7] metric referred to as Hedges' *g* [@hedgesStatisticalMethodsMetaAnalysis2014; @borensteinIntroductionMetaAnalysis2021; @nakagawaEffectSizeConfidence2007][^8].
+Often, due to the varying operationalisations used for broad outcome concepts[^5], an effect size is used which is "standardised" across studies to allow for their synthesis.
+Most commonly, a magnitude based[^6] effect size statistic [@caldwellCaseDefaultEffect2020], the standardised mean difference (SMD), is used to compare means between groups or conditions.
+This statistic is usually a version of Cohen's *d* [@cohenStatisticalPowerAnalysis1988], or its bias-corrected[^7] metric referred to as Hedges' *g*[^8] [@hedgesStatisticalMethodsMetaAnalysis2014; @borensteinIntroductionMetaAnalysis2021; @nakagawaEffectSizeConfidence2007].
 The SMD, and its sampling variance, $s^2_{SMD}$ are given by:
 
-[^5]: \textcolor{red}{For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions.
-Or the same operationalisations may be employed but different exercises such as the squat or bench press.}
+[^5]: For example, *strength* might be examined in different studies using different operationalisations including one repetition maximum testing or maximum voluntary contractions.
+Or the same operationalisations may be employed but different exercises such as the squat or bench press.
 
@@
 
-[^7]: \textcolor{red}{For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct.
-Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.}
+[^7]: For those unfamiliar with the terminology, an estimator for a statistic is unbiased if it produces parameter estimates that are on average correct.
+Thus a bias corrected statistic is one which would be biased without the correction applied, but otherwise has been shown to be unbiased.
 
 [^8]: We will refer to both merely as the SMD throughout the manuscript for simplicity and note that throughout when reporting a 'SMD' we are reporting the bias-corrected version.
-\textcolor{red}{We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance;} @caldwellCaseDefaultEffect2020).
+We also note that another magnitude based effect size, Glass' $\Delta$, is commonly recommended as it is the simplest form of SMD though makes assumptions about the impact of the intervention having no effect on the denominator (i.e., variance; @caldwellCaseDefaultEffect2020).
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 where $\overline{x}_C$ and $\overline{x}_E$ are the sample means of the control group (C) and experimental (E) or intervention group respectively, $s_C$ and $s_E$ are the standard deviations of the two groups, $n_C$ and $n_E$ are the sample sizes of the two groups, and $J$ is a bias correction for small sample sizes.
@@
 The natural logarithm of the ratio of two means ($\textrm{ln}RR$) is also another effect size statistic that can be used [@curtisMetaanalysisElevatedCO21998; @hedgesMetaAnalysisResponseRatios1999; @lajeunesseMetaanalysisResponseRatios2011; @lajeunesseBiasCorrectionLog2015].
-The lnRR, and its sampling variance, $s^2_{\textrm{ln}RR}$ are given by:
+The bias corrected lnRR, and its sampling variance, $s^2_{\textrm{ln}RR}$ are given by:
 
+```{=tex}
 \begin{equation}
-\textrm{ln}RR=\textrm{ln}\frac{\overline{x}_E}{\overline{x}_C}
+\textrm{ln}RR=\textrm{ln}\frac{\overline{x}_E}{\overline{x}_C} + \frac{1}{2}(\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}-\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}})
 (\#eq:lnRR-eq)
 \end{equation}
-
+```
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 Due to its calculation the SMD is affected not only by the difference in means of the two groups, but also by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$ in the denominator.
-In contrast, the $\textrm{ln}RR$ is uninfluenced by the standard deviations in either groups (see equation \@ref(eq:lnRR-eq)), which only affects the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
+In contrast, the $\textrm{ln}RR$ is uninfluenced in magnitude by the standard deviations in either groups except for the bias correction factor (see second term in equation \@ref(eq:lnRR-eq)), and the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
 Despite this, the use of effect sizes like the $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@debQuantifyingEffectsAcute2018; @nuzzoEccentricConcentricStrength2023] and to our knowledge only a couple of meta-analyses of RT interventions has used this kind of effect size [@swintonInterpretingMagnitudeChange2022; @wolfPartialVsFull2023].
@@
 The increased interest in *precision* or *personalised* approaches to exercise prescription has resulted in a number of opinion and methodological review articles discussing statistical approaches to understanding interindividual response heterogeneity to exercise interventions [@heckstedenIndividualResponseExercise2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @rossPrecisionExerciseMedicine2019; @swintonStatisticalFrameworkInterpret2018; @hopkinsIndividualResponsesMade2015; @kelleyPrecisionExerciseMedicine2022; @hrubeniukDirectionsExerciseTreatment2022; @pickeringNonRespondersExerciseExistand2019].
-However, despite the availability of approaches to compare variances between groups, in sport and exercise science this is rarely explored in primary research [@bonafigliaInterindividualDifferencesTrainability2022].
-Moreover, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variances or explicitly aim to investigate the latter [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
+However, despite the availability of approaches to compare variation between groups, in sport and exercise science this is rarely explored in primary research [@bonafigliaInterindividualDifferencesTrainability2022].
+Moreover, although there has been increased interest in recent years, few meta-analyses in sport and exercise include both comparisons of means and variation or explicitly aim to investigate the latter [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
 Examination of interindividual heterogeneity in response to interventions presents considerable value to researchers and practitioners in sport and exercise science; interventions with low interindividual variation are likely to be widely generalisable, whilst an intervention with high interindividual variation is likely to have effects that are either subgroup or individual specific.
@@
 
-Comparison of heterogeneity in responses, such as post-scores or change scores to interventions, are not the only possible use of statistical methods for comparing variances.
-For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawaMeanStrikesBack2012].
-Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] \textcolor{red}{and in the number of repetitions that can be performed at different percentages of one repetition maximum} [@nuzzoMaximalNumberRepetitions2023].
+Comparison of heterogeneity in responses, such as post-scores or change scores to interventions, are not the only possible use of statistical methods for comparing variation For example, in other fields such as ecology there have been calls to shift focus of analysis onto the exploration of dispersion of traits between groups in non-experimental or intervention designs [@nakagawaMeanStrikesBack2012].
+Some recent examples from sport and exercise science, and RT in particular, include primary research exploring between-participant acute response variation for the purposes of identifying methods[^9] to reduce RT stimulus heterogeneity, [@exnerDoesPerformingResistance2022] as well as a meta-analysis exploring between-participant heterogeneity of accuracy in predicting proximity to task failure during RT [@halperinAccuracyPredictingRepetitions2022] and in the number of repetitions that can be performed at different percentages of one repetition maximum [@nuzzoMaximalNumberRepetitions2023].
 
@@
 
-\textcolor{red}{Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
+Although ultimately we will recommend the thoughtful consideration of assumptions for certain approaches regarding what effect size statistics and models to employ in examining variation through use of meta-analysis, we provide examples throughout using all approaches described in order to aid the reader in understanding their strengths and weaknesses.
 We hope this will make clear why we offer such recommendations in our discussion and conclusion.
-As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.}
+As we have up to this point, we will also make an effort to both present the mathematical formulations of the models described, as well as to provide an explanation of them in plain English.
 
 # Effect size statistics for meta-analytic comparisons of variation
+
 Until recent years there has been a dearth of effect size statistics available for the examination of variation in a meta-analytic framework.
@@
 ## Standard deviation for individual responses ($SD_{ir}$)
+
 In the context of *precision* or *personalised* approaches to exercise prescription the $SD_{ir}$ has been proposed as an approach to determine the extent to which individual responses manifest by comparison of variation between two groups; control and intervention [@hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019].
@@
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-Thus, the $SD_{ir}$ reflects a comparison of the absolute variance in change scores between control and intervention groups.
+```
+Thus, the $SD_{ir}$ reflects a comparison of the absolute variation in change scores between control and intervention groups.
 However, a potential concern with the $SD_{ir}$ is its potential to violate assumptions of normality, which is not the case for other effect size statistics such as $\textrm{ln}VR$ and $\textrm{ln}CVR$.
@@
 ## Log ratio of standard deviations ($\textrm{ln}VR$)
-A similar effect size statistic for the comparison of absolute variance between groups, and one that has had wide applications in more than just intervention response variability within fields such as ecology and evolution, is the $\textrm{ln}VR$ [@hedgesSexDifferencesMental1995; @nakagawaMetaanalysisVariationEcological2015; @seniorRevisitingExpandingMetaanalysis2020].
+
+A similar effect size statistic for the comparison of absolute variation between groups, and one that has had wide applications in more than just intervention response variability within fields such as ecology and evolution, is the $\textrm{ln}VR$ [@hedgesSexDifferencesMental1995; @nakagawaMetaanalysisVariationEcological2015; @seniorRevisitingExpandingMetaanalysis2020].
 An unbiased estimator of the natural logarithm of a population standard deviation ($\textrm{ln}\sigma$), and its sampling variance, $s^2_{\textrm{ln}\sigma}$ is given by:
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
+```
+where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$.
+Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), an unbiased estimator of the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
 
-where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$.
-Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
-
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparison group (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$).
 In this case, it is likely that the standard deviation will be larger in the group with the larger mean (e.g., $s_{E}$ is larger than $s_{C}$).
-This mean-variance relationship is common for many variables and datasets[^11] and to highlight this we provide examples below.
+This mean-variation relationship is common for many variables and datasets[^11] and to highlight this we provide examples below.
 They also assume constant measurement error over the range of values for the mean, which can impact their utility for examining response variation [@tenanCommentMethodStop2020].
@@
 ## Log ratio of coefficient of variation ($\textrm{ln}CVR$)
-The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variance relationship is present.
-\textcolor{red}{In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.} The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
-Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), the $\textrm{ln}CVR$ is given by:
 
+The coefficient of variation is the ratio of the standard deviation to the mean; therefore, comparison of the coefficient of variation between groups will identify whether standard deviations differ more, or less, than would be predicted by their difference in means where a mean-variation relationship is present.
+In essence, the coefficient of variation is a means of standardising the standard deviation against the mean such that the relative variation in an effect is expressed.
+The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
+Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), an unbiased estimator of the $\textrm{ln}CVR$ is given by:
+
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 where $CV_{E}$ and $CV_{C}$ are ${s_{E}}/{\overline{x}_{E}}$ and ${s_{C}}/{\overline{x}_{C}}$ respectively.
@@
 \end{equation}
+```
+# Examples using resistance training studies
 
-# Examples using resistance training studies
 As noted, to facilitate understanding for those new to examination of variation, we provide primary examples of the approaches presented using data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021].
@@
 All analysis examples were performed in R (version 4.2.1, "Funny-Looking Kid", The R Foundation for Statistical Computing, 2022) using the **metafor** package [@viechtbauerConductingMetaAnalysesMetafor2010].
-The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (https://osf.io/2h9ma/) \textcolor{red}{or the GitHub repository (https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git)}.
+The extracted dataset, analysis scripts, models, data summaries, and supplementary materials are available on the Open Science Framework (<https://osf.io/2h9ma/>) or the GitHub repository (<https://github.com/jamessteeleii/Meta-Analysis-of-Variation-in-Resistance-Training.git>).
 
@@
 Given that Polito et al.
-[-@politoModeratorsStrengthGains2021] included only studies with a non-training control group, their study selection offers a unique context to examine variation of interindividual responses specifically by comparing the variances in change scores between the RT intervention groups(s) and control group.
+[-@politoModeratorsStrengthGains2021] included only studies with a non-training control group, their study selection offers a unique context to examine variation of interindividual responses specifically by comparing the variation in change scores between the RT intervention groups(s) and control group.
 Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies.
-Indeed, this highlights the typically small samples used in sport and exercise science, and thus low power to detect difference in both means and variances in in individual studies [@yangLowStatisticalPower2022], emphasising the value of meta-analysis to explore variation.
+Indeed, this highlights the typically small samples used in sport and exercise science, and thus low power to detect difference in both means and variation in in individual studies [@yangLowStatisticalPower2022], emphasising the value of meta-analysis to explore variation.
 Table \@ref(tab:summary-characteristics-tab) shows the study and participant characteristics.
 
-[^13]: \textcolor{red}{Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
+[^13]: Regression analyses are likely familiar to most readers where in the simplest form they try to predict the value of some dependent variable from some independent variable(s).
 This can be extended to meta-analytic synthesis where the independent variables reflect characteristics associated with the effects included.
-For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.}
+For example, they may reflect characteristics of the sample in the study for which the effect was extracted such as age or sex, or they might reflect characteristics of the intervention received such as the dose or frequency of exposure.
 
@@
 
-## Detecting the presence of interindividual response variation to resistance training intervention \textcolor{red}{with absolute variance statistics}
+## Detecting the presence of interindividual response variation to resistance training intervention with absolute variation statistics
+
 First we conducted a traditional SMD and $\textrm{ln}RR$ based effect size[^14] meta-analysis to explore the effects of RT interventions compared to controls for strength and hypertrophy (i.e., muscle mass/size) outcomes[^15].
@@
 However, the data we extracted were hierarchical in nature.
-\textcolor{red}{Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.} Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
-\textcolor{red}{This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
-We can therefore describe the overall model as:}
+Thus, as opposed to there being only the assumption that studies are a random effect, due to there being multiple outcomes measured within each arm in the studies (i.e., intervention group(s) and control group, within each study), and that in some studies there were multiple interventions examined, there is the additional assumption that must be included that both the intervention groups and effects also come from overarching distributions.
+Thus a multilevel mixed-effects meta-analysis model [@vandennoortgateThreelevelMetaanalysisDependent2013] with cluster-robust variance estimation [@hedgesRobustVarianceEstimation2010] was used with random intercepts for study, arm[^16], and effect.
+This model then includes additional $\tau^2$ terms for each of the levels and assigns weights appropriately given this and the clustering of effects within arms within studies.
+We can therefore describe the overall model as:
 
-[^14]: It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental intervention, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure).
+[^14]: It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental interventions, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure).
 We presented the SMD and $\textrm{ln}RR$ effect sizes in equations \@ref(eq:SMD-eq) and \@ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., [-@gurevitchInteractionCompetitionPredation2000]; Morris et al., [-@morrisDirectInteractiveEffects2007]; Morris [-@morrisEstimatingEffectSizes2008]; Lajeunesse [-@lajeunesseMetaanalysisResponseRatios2011; -@lajeunesseBiasCorrectionLog2015]) and these are the effect sizes used in the meta-analyses referred to here.
 
-[^15]: We also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variance based effect sizes in the field implies that they might have more influence over such biases.
-There did not appear to be any obvious small study bias in the dataset (see https://osf.io/stqr3).
+[^15]: We also explored for signs of small study bias, including publication bias favouring the finding of intervention effects, for the SMDs given that the relative lack of awareness for variation based effect sizes in the field implies that they might have more influence over such biases.
+There did not appear to be any obvious small study bias in the dataset (see <https://osf.io/stqr3>).
 
@@
 \end{equation}
+```
+where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study, $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
+This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects[^17].
+The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).
 
-\textcolor{red}{where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study,  $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
-This model is referred to as 'mixed' effects because of the presence of both fixed ($\theta$), and random ($\tau_{(1)i},\tau_{(2)j},\tau_{(3)k}$) effects}[^17].
-\textcolor{red}{The main term in the model we are interested in is $\theta$ which is our estimate for the overall weighted average effect (i.e., $\hat\theta$).}
+[^17]: Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.
 
-[^17]: \textcolor{red}{Technically then the *random* effects model presented earlier is also a *mixed* effects model. It is traditionally referred to as the random-effects model though.}
-
-We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variance in responses differed between RT interventions and controls.
+We then fit the same model for the $SD_{ir}$ and $\textrm{ln}VR$ effect sizes for change scores (i.e., post-intervention minus pre-intervention scores) in order to explore how absolute variation in responses differed between RT interventions and controls.
 A positive SMD or $\textrm{ln}RR$ would indicate that RT interventions produced greater improvements in outcomes compared to controls, whilst a positive $SD_{ir}$ and $\textrm{ln}VR$ would indicate that the introduction of the RT intervention increased variation in responses (i.e., change scores) compared to controls (i.e., suggests the presence of interindividual response variation).
@@
 As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%).
-Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, \textcolor{red}{indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure }\@ref(fig:forest-SMD-plot)) \textcolor{red}{were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).}
+Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).
 
@@
 
-In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[3]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[4]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[7]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[8]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[8]`%) indicating that exposure to the RT interventions may have introduced additional variance over and above random error, potentially suggesting the presence of interindividual response variation.
+In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[3]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[4]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[7]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[8]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[8]`%) indicating that exposure to the RT interventions may have introduced additional variation over and above random error, potentially suggesting the presence of interindividual response variation.
 Although, heterogeneity across the models and levels (study, arm, effect) were again relatively large and quite varied.
 
-This additional variance might support previous perspectives [@carpinelliInterindividualHeterogeneityAdaptations2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* independent of the random error).
-However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variance over values of the mean \textcolor{red}{(i.e., that the variance is similar whether mean values are low or high)}.
+This additional variation might support previous perspectives [@carpinelliInterindividualHeterogeneityAdaptations2017] that the considerable variation in responses to RT interventions typically observed are due to 'true' interindividual response variation over and above the random error that occurs from pre- and post-intervention measurements (i.e., the variation is *detectable* independent of the random error).
+However, as noted, both the $SD_{ir}$ and $\textrm{ln}VR$ assume constant variation over values of the mean (i.e., that the variation is similar whether mean values are low or high).
 As we have seen from the SMD and $\textrm{ln}RR$ models, RT interventions increase mean scores.
-Thus, if there is a mean-variance relationship in the data, an increase in the mean alone may be fully responsible for any apparent increase in variation.
-As such, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
+Thus, if there is a mean-variation relationship in the data, an increase in the mean alone may be fully responsible for any apparent increase in variation.
+As such, we cannot rely solely on absolute comparisons of variation such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
 The $\textrm{ln}CVR$ can be used to overcome this issue, and below we re-analyse this dataset using this effect size statistic.
-First though, we present data demonstrating the ubiquity of the mean-variance relationship in typical RT study outcome measures \textcolor{red}{to emphasise the need to consider this assumption,} and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ \textcolor{red}{and provide flexibility to accomodate wider applications}.
+First though, we present data demonstrating the ubiquity of the mean-variation relationship in typical RT study outcome measures to emphasise the need to consider this assumption, and introduce a modelling approach that can also be used to overcome some possible limitations with the $\textrm{ln}CVR$ and provide flexibility to accomodate wider applications.
 
@@
 ```
+
 ```{r forest-logVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$VR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}
@@
 
-## Mean-variance relationships in muscular strength and hypertrophy
+## Mean-variation relationships in muscular strength and hypertrophy
+
 With meta-analytic models of variation we are not limited to solely exploring variation in responses to interventions (e.g., @halperinAccuracyPredictingRepetitions2022; @nuzzoMaximalNumberRepetitions2023).
-We can explore the relationships between variance in a number of outcomes and the impact of certain predictors on this in the form of meta-regression.
+We can explore the relationships between variation in a number of outcomes and the impact of certain predictors on this in the form of meta-regression.
 For example, as noted, one possible predictor of variance is the mean itself.
-As such, we can model variance of each effect as the response itself with the mean of the effect as the predictor.
-The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models \textcolor{red}{(i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).} Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
-In the following example we explore the mean-variance relationship in the pre-intervention scores for outcomes in the data set from Polito et al.
+As such, we can model variation of each effect as the response itself with the mean of the effect as the predictor.
+The standard deviation is, however, bounded at zero and so in many cases it may not conform to assumptions of normality which are required for regression models (i.e., that the residuals, the difference between the estimated and actual data, are normally distributed).
+Therefore, we instead can use $\textrm{ln}\hat\sigma$, which is unbounded.
+In the following example we explore the mean-variation relationship in the pre-intervention scores for outcomes in the data set from Polito et al.
 [-@politoModeratorsStrengthGains2021].
 
-```{r mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes."}
+```{r mean-variation-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes."}
 
@@
 
-As can be seen in figure \@ref(fig:mean-variance-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
-\textcolor{red}{The variance in standard deviations increases with higher mean values.} This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
+As can be seen in figure \@ref(fig:mean-variation-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
+The range of standard deviations increases with higher mean values.
+This is similar to what is known as Taylor's law in ecology, or the power law; in essence, an empirically derived relationship stating that the variance is a power function of the mean in many biological and physical systems [@taylorAggregationVarianceMean1961].
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 where $a$ and $b$ are some constants.
 When this relationship holds, under most circumstances the standard deviation is not proportional to the mean.
-However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear \textcolor{red}{based upon the product and power logarithmic rules:}
+However, when the mean and standard deviation are transformed to the log scale this relationship becomes linear based upon the product and power logarithmic rules:
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
-Figure \@ref(fig:mean-variance-pre-plot)(B) and (D) shows that the relationship between the mean and variance on the log scale better meets the assumption of normality.
+```
+Figure \@ref(fig:mean-variation-pre-plot)(B) and (D) shows that the relationship between the mean and variation on the log scale better meets the assumption of normality.
 Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from multiple outcomes within multiple arms within studies we can also estimate this relationship using a multilevel mixed-effects meta-regression model similar to that applied above.
@@
 \end{equation}
-
+```
 where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^18]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept or overall mean of the effects, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study, $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
-
-\textcolor{red}{These kinds of models are highly flexible.} Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
 
@@
 
+These kinds of models are highly flexible.
+Additional predictor terms could be added; for example, we could model a categorical variable for the outcome type and include $\beta_{2}Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the slope or regression coefficient for $Outcome$ (most intuitively thought of as the difference between the two outcome types)[^19].
+
 [^19]: We do not have to limit ourselves to only fixed effect predictor terms as we have here.
-Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes \textcolor{red}{(i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences} [@barrRandomEffectsStructure2013].
+Indeed, for mixed effect models generally some argue that models should use a *maximal* random effects structure including both random intercepts and slopes (i.e., that the effect of the predictor term can vary within different levels of the model and is also assumed to come from an overarching distribution of slopes), and their correlations, to enhance generalisability of inferences [@barrRandomEffectsStructure2013].
 We could model a categorical variable for the outcome type and using random effects include $(\beta_{2} + \varphi_{i})Outcome$ or $(\beta_{2} + \varphi_{i} + \varphi_{j})Outcome$ in the model with $Outcome$ as a dummy coded variable for the outcome type (i.e., hypertrophy = 0, and strength = 1), where $\beta_{2}$ is the overall average slope or regression coefficient for $Outcome$, and $\varphi_{i}$ is the deviation (random slope) from $\beta_{2}$ for the $i\textrm{th}$ study and $\varphi_{j}$ is the deviation for the $j\textrm{th}$ arm.
-\textcolor{red}{These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms.
+These model specifications do not assume that the difference between outcomes is fixed, but can vary between studies and arms.
 We could also do the same and include random slopes for $(\beta_{1}$ on $\textrm{ln}\overline{x}$ thus allowing for the strength of the relationship between $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ to also vary between studies and arms.
 Indeed, we fit a range of models using $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Outcome$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Outcome$ by study, (5) the inclusion of correlated random slopes for $Outcome$ by study and arm, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study, (7) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and arm.
-The comparison of these models using $2\times\textrm{log}BF$} [@kassBayesFactors1995] \textcolor{red}{from approximate Bayesian information criterion} [@wagenmakersPracticalSolutionPervasive2007] \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (https://osf.io/3tv6x).
-There was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.}
+The comparison of these models using $2\times\textrm{log}BF$ [@kassBayesFactors1995] from approximate Bayesian information criterion [@wagenmakersPracticalSolutionPervasive2007] to determine under which is the observed data most likely is included in the supplementary materials (<https://osf.io/3tv6x>).
+There was very strong evidence supporting the random intercepts plus correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study model compared to all others and so this is presented here.
 
-Figure \@ref(fig:model-mean-variance-pre-plot) shows this model fit visually where the size of the points reflects their weight in the model.
-Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight but systematically greater degree of variance in strength measures compared to hypertrophy for a given mean score.
+Figure \@ref(fig:model-mean-variation-pre-plot) shows this model (note, see $\textrm{footnote}^{19}$) fit visually where the size of the points reflects their weight in the model.
+Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight but systematically greater degree of variation in strength measures compared to hypertrophy for a given mean score.
 
-```{r model-mean-variance-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores."}
+```{r model-mean-variation-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores."}
 
@@
 
-The presence of Taylor's law type relationships should be examined in datasets prior to deciding on which variance effect size statistic should be employed.
-Returning to the context of interindividual response variation to interventions, the presence of a mean-variance relationship in the data would imply that we cannot rely on absolute comparisons of variance (i.e., $SD_{ir}$ or $\textrm{ln}VR$) to determine whether interindividual response variation is actually present.
-So we should also explore this assumption for the change-scores in the RT and control groups and determining the appropriate effect sizes to explore.
+The presence of Taylor's law type relationships should be examined in datasets prior to deciding on which variation effect size statistic should be employed.
+Returning to the context of interindividual response variation to interventions, the presence of a mean-variation relationship in the data would imply that we cannot rely on absolute comparisons of variation (i.e., $SD_{ir}$ or $\textrm{ln}VR$) to determine whether interindividual response variation is actually present.
+So we should also explore this assumption for the change-scores in the RT and control groups and determine the appropriate effect sizes to explore.
 
 ## Reanalysis of interindividual response variation using $\textrm{ln}CVR$
-As can be seen in figures \@ref(fig:mean-variance-delta-plot)(A) and (C) there is also a mean-variance relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
-Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures \@ref(fig:mean-variance-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seen with the pre-intervention scores above.
-As such, in this case, we cannot rely solely on absolute comparisons of variance such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
 
-```{r mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes."}
+As can be seen in figures \@ref(fig:mean-variation-delta-plot)(A) and (C) there is also a mean-variation relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
+Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures \@ref(fig:mean-variation-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seen with the pre-intervention scores above.
+As such, in this case, we cannot rely solely on absolute comparisons of variation such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
+
+```{r mean-variation-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes."}
 
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue though.
-Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before (see equation \@ref(eq:ml-model-eq)) using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variance comparisons using $SD_{ir}$ or $\textrm{ln}VR$.
+Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before (see equation \@ref(eq:ml-model-eq)) using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variation comparisons using $SD_{ir}$ or $\textrm{ln}VR$.
 The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[28]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[28]`%) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[88]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[88]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[88]`%) and further there is lower relative heterogeneity between studies in the effect estimates (figure \@ref(fig:forest-logCVR-plot)).
@@
 Firstly, it is limited to the use of ratio scale data (which is not the case for the $\textrm{ln}\hat\sigma$ or $\textrm{ln}VR$); hence the need to transform the change scores to be positively signed in this specific case.
-Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variance relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
-Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variance-delta-plot)) an alternative yet \textcolor{red}{comparable} approach might be desirable.
-\textcolor{red}{Lastly, this statistic is limited to examination of pairwise comparisons of variance.
+Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variation relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
+Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variation-delta-plot)) an alternative yet comparable approach might be desirable.
+Lastly, this statistic is limited to examination of pairwise comparisons of variation.
 In this example this approach poses no issue as we are comparing RT intervention group(s) to a control group.
-But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a \textcolor{red}{comparable} effect of interest to the $\textrm{ln}CVR$ one just applied.}
+But, as seen with the pre-score example above, alternative modelling approaches provide greater flexibility and can still offer estimation of a comparable effect of interest to the $\textrm{ln}CVR$ one just applied.
 
@@
 # load the summary dataframes to the environment
-load(here("models/RobuEstMultiLevelModel_ri_only_log_mean_mod_strength"))
-load(here("models/RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy"))
+load(here("models/RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength"))
+load(here("models/RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy"))
 
@@
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
-\textcolor{red}{This kind of meta-regression model with  $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variance in the RT groups vs the control groups.
-This reflects the pairwise nature of the $\textrm{ln}CVR$.}
+This kind of meta-regression model with $Group$ as a predictor is comparable to the $\textrm{ln}CVR$ model in the previous section where the $\beta_{2}Group$ is the slope or regression coefficient for $Group$ and reflects the difference i.e., variation in the RT groups vs the control groups.
+This reflects the pairwise nature of the $\textrm{ln}CVR$.
 
-```{r model-mean-variance-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores."}
+```{r model-mean-variation-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores."}
 
@@
 
-Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variance-delta-plot)), we fit this model and other model specifications[^20]\textsuperscript{,}[^21] to the dataset (see figure \@ref(fig:model-mean-variance-delta-plot)).
-The results were largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_only_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
-Both models suggest that the introduction of the RT interventions *reduced* variance in change scores.
+Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variation-delta-plot)), we fit this model and other model specifications[^20]\textsuperscript{,}[^21] to the dataset (see figure \@ref(fig:model-mean-variation-delta-plot)).
+The results from these models (note, see $\textrm{footnote}^{20}$) were largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
+Both models suggest that the introduction of the RT interventions *reduced* variation in change scores.
 
-[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with \textcolor{red}{(1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Outcome$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).
-The comparison of these models using $2\times\textrm{log}BF$} [@kassBayesFactors1995] \textcolor{red}{from approximate Bayesian information criterion} [@wagenmakersPracticalSolutionPervasive2007] \textcolor{red}{to determine under which is the observed data most likely is included in the supplementary materials (see https://osf.io/b5deh for strength and https://osf.io/5ektd for hypertrophy).
-Similar to the models including $Outcome$ there was very strong evidence supporting the random intercepts only model compared to all others and so this is presented here.
-All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (https://osf.io/5g7ce) all of which revealed similar conclusions.}
+[^20]: Note, as with the models examining $Outcome$ upon baseline scores, we similarly explored $\textrm{ln}\hat\sigma$ with $\textrm{ln}\overline{x}$ and $Group$ as a predictor with (1) random intercepts only for study and arm, (2) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study, (3) the inclusion of correlated random slopes for $\textrm{ln}\overline{x}$ by study and arm, (4) the inclusion of correlated random slopes for $Group$ by study, (5) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study, and (6) the inclusion of correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study and $\textrm{ln}\overline{x}$ by arm (we do not include the models with random slopes for $Group$ by arm as in this model each arm refers to a particular group, RT or CON, and so no arm provides data for both).
+The comparison of these models using $2\times\textrm{log}BF$ [@kassBayesFactors1995] from approximate Bayesian information criterion [@wagenmakersPracticalSolutionPervasive2007] to determine under which is the observed data most likely is included in the supplementary materials (see <https://osf.io/b5deh> for strength and <https://osf.io/5ektd> for hypertrophy).
+Similar to the models including $Outcome$ there was very strong evidence supporting the random intercepts plus correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study and $\textrm{ln}\overline{x}$ by arm model for strength, and random intercepts plus correlated random slopes for both $\textrm{ln}\overline{x}$ and $Group$ by study for hypertrophy, compared to all other models and so these are presented here.
+All estimates for the difference between RT and CON, where positive values indicate RT *increased* variation in changes scores and negative values indicate it *decreased* variation, can be seen in the supplementary materials (<https://osf.io/5g7ce>) all of which revealed similar conclusions.
 
-[^21]: \textcolor{red}{It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variance relationship.
+[^21]: It is perhaps worth explaining the assumptions that the different models explored make regarding the mean-variation relationship.
 For example, the $\textrm{ln}VR$ and $\textrm{ln}CVR$ models can be thought of as similar in that they both make fixed assumptions about the relationship between mean and variance; in the $\textrm{ln}VR$ it is assumed to be zero, and in the $\textrm{ln}CVR$ it is assumed to be proportional i.e., one.
 In both however this is a *strong* assumption.
-The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in model where random slopes are included this is also estimated allowing it to vary between studies and arms (i.e., in some studies there may be a more or less strong relationship compared to others).
-Mean-variance relationships are important to consider when exploring variance effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression models allow).
+The multilevel meta-regressions on the other hand actually *estimate* this relationship (i.e., the value of $\beta_{1}$, the slope or regression coefficient for $\textrm{ln}\overline{x}$) and in models where random slopes are included this is also estimated allowing it to vary between studies and/or arms (i.e., in some studies there may be a more or less strong relationship compared to others).
+Mean-variation relationships are important to consider when exploring variation effects, but it is also important to consider whether or not this relationship is assumed to be some fixed proportional value (i.e., as the $\textrm{ln}CVR$ does) or whether or not this should be estimated from the data and whether it might also vary across studies and arms (i.e., as the multilevel meta-regression models allow).
 It should also be noted that these models all assume that the $\overline{x}$ is estimated without error which is clearly not the case.
@@
 This is beyond the scope of this paper to discuss.
-It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.}
+It is not necessarily clear which model should be preferred here, and fortunately substantive conclusions are impacted little by model specification, but thought should be given to the assumptions each makes and the fit of each model to the data.
 
 ## Moderators of variance effects
-\textcolor{red}{Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variance through a meta-analytic framework.
-Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variance-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)}[^22].
-\textcolor{red}{But these are limited to the pairwise comparisons of a categorical variable as the effect size.
-The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.}
 
-[^22]: See supplementary materials (https://osf.io/e6vpr) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderator analysis results) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
+Hopefully it is clear from the meta-regression models here, where we have included both fixed and random predictors as both categorical (i.e., $Outcome$, or $Group$) and continuous (i.e., $\textrm{ln}\overline{x}$) variables, that there is considerable flexibility in the inclusion of predictors when exploring variation through a meta-analytic framework.
+Of course, the pairwise models described can also be extended to meta-regressions to explore not only how study, arm, or effect level characteristics moderate effect size estimates when considering effect sizes such as SMDs or $\textrm{ln}RR$, but also when considering the variation-based effect size statistics and models employed in this article (i.e., $SD_{ir}$, $\textrm{ln}VR$, $\textrm{ln}CVR$, and $\textrm{ln}\hat\sigma$)[^22].
+But these are limited to the pairwise comparisons of a categorical variable as the effect size.
+The meta-regression models presented here for $\textrm{ln}\hat\sigma$ allow for comparisons to be extended beyond two categories (e.g., intervention, placebo, and control) including any number of arbitrary predictors, fixed and random and assumptions about their correlations.
+
+[^22]: See supplementary materials (<https://osf.io/e6vpr>) for examples from model estimates for both SMD and $\textrm{ln}CVR$, (used for simplicity of presenting moderator analysis results) across a range of categorical and continuous predictors for both strength and hypertrophy outcomes.
 There were no obvious moderators of $\textrm{ln}CVR$ in particular.
@@
 # Discussion
-Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variance, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
+
+Given the apparent lack of awareness of the utility of meta-analytic frameworks for exploring variation, and the potential value such analyses can offer for the sport and exercise sciences, we have presented some existing effect size statistics and models pertinent to this topic that hopefully will encourage and support researchers in the field to embrace more than just the mean when engaging in quantitative evidence synthesis.
 Indeed, for a field such as sport and exercise science where sample sizes are typically small, meta-analysis becomes even more valuable as such small samples in primary studies have even lower statistical power to detect differences in variation as compared to means[^23] [@yangLowStatisticalPower2022].
@@
 It is of particular interest to note the different conclusions drawn here are dependent on the approach taken to determine from non-training control and RT intervention data whether or not there is *detectable* inter-individual response variation present.
-\textcolor{red}{We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
+We deliberately presented these varying approaches to highlight their assumptions and to aid in readers understanding of their applications.
 Given the different conclusions drawn from the examples provided, we recommend that researchers consider whether assumptions of simpler approaches are met, or indeed seem reasonable, before their application in specific situations.
-Where for example there is not an obvious mean-variance relationship it may be appropriate to utilise the $SD_{ir}$ or $\textrm{ln}VR$.
+Where for example there is not an obvious mean-variation relationship it may be appropriate to utilise the $SD_{ir}$ or $\textrm{ln}VR$.
 However, when this is present in the data then the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ may be more appropriate, though assumptions regarding whether this relationship is proportional and fixed should also be considered.
-The structure of the data also impacts the specific modelling approach to be employed and whether or not assumptions that the mean-variance relationship varies within study or arm levels can be incorporated.
+The structure of the data also impacts the specific modelling approach to be employed and whether or not assumptions that the mean-variation relationship varies within study or arm levels can be incorporated.
 Further, depending on the exact research questions it is worth considering balancing model complexity with its ability to provide an answer.
@@
 As such, it is worth considering the use of so called *multiverse* approaches to exploring questions such as those pertaining to variation with meta-analysis.
-With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions} [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
+With multiverse approaches, instead of having to choose one modelling approach, researchers use the many possible approaches (such as presented here) and explore how sensitive their substantive conclusions might be to nuances of model specifications and assumptions [@steegenIncreasingTransparencyMultiverse2016; @olsson-collentineMetaanalyzingMultiversePeek2023].
 
 The examples presented herein used data from RT studies included in a recent meta-analysis published in the *Journal of Sport Sciences* [@politoModeratorsStrengthGains2021], which hopefully makes them more relatable for researchers in sport and exercise sciences.
-Examining this data using absolute comparisons of variance such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely *increased* variance above random error, suggesting the presence of inter-individual response variation.
+Examining this data using absolute comparisons of variation such as $SD_{ir}$ and $\textrm{ln}VR$ gave the impression that the introduction of the RT intervention likely *increased* variation above random error, suggesting the presence of inter-individual response variation.
 In the case of RT interventions there is evidently an average intervention effect for strength and hypertrophy which is positive (as seen in the SMD and $\textrm{ln}RR$ models), yet combined with the results from $SD_{ir}$ and $\textrm{ln}VR$ we might conclude that while all likely benefit, some benefit more so than others.
-Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variance could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
+Indeed, even if for an intervention there was not clear evidence for average intervention effects, exploring variation in their absence might still be important as interventions with large enough variation could imply that the intervention is at least beneficial to some [@usuiMetaanalysisVariationSuggests2021].
 Such results might lead researchers to consider that further research exploring subgroup- or participant-by-intervention interactions is required to maximise successful practical application of such an intervention to avoid negative effects for some, and ensure positive effects for others.
 
-However, similar to the cross-sectional pre-intervention score models reported here (and indeed most physical and biological variables), change scores also demonstrated a mean-variance relationship in addition to heteroskedasticity.
-The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite substantive direction of the absolute variance comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variance, implying that there is likely little to no interindividual response variation to explain.
+However, similar to the cross-sectional pre-intervention score models reported here (and indeed most physical and biological variables), change scores also demonstrated a mean-variation relationship in addition to heteroskedasticity.
+The likely more appropriate analyses in this case using the $\textrm{ln}CVR$ or meta-regression of $\textrm{ln}\hat\sigma$ upon $\textrm{ln}\overline{x}$ revealed conclusions in the opposite substantive direction of the absolute variation comparisons; essentially, that the introduction of the RT intervention may have slightly *decreased* change score variation, implying that there is likely little to no interindividual response variation to explain.
 Interventions, such as RT interventions explored here, which induce both meaningful average treatment effects and also show little evidence suggestive of interindividual variation, are likely to be widely generalisable and so from a practical perspective might offer considerable value in that we can have high expectations that everyone receiving them will likely improve [@usuiMetaanalysisVariationSuggests2021]; that is to say we can assume a constant effect and that the average intervention effect is indicative of the individual intervention effect [@cortesmartinezConstantEffectRandomized2021].
@@
 The reason for the apparent reduction in variation after introduction of an RT intervention observed here is not necessarily discernible from this analysis.
-Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variance (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021].
+Perhaps the introduction of an RT intervention has indirect effects that reduce other sources of random variation (e.g., diet, other physical activity etc.; @hallidayResistanceTrainingAssociated2017), or a ceiling effect on change (i.e., plateau in response; @steeleLongTermTimeCourseStrength2022) has a constraining effect [@cortesmartinezConstantEffectRandomized2021].
 However, this potentially represents another interesting area of future study regarding variation; specifically, how to produce interventions that actually reduce variance in an outcome.
@@
 # Conclusion
+
 Embracing variability and focusing on more than merely the mean differences between groups or conditions, such as intervention and control comparisons, has the potential to inform experimental design and lead to changes in both the approach and direction of follow-up studies.
-Whether there is evidence of meaningful average intervention effects or not, where considerable variance effects are present it suggests that a meaningful line of research would be to aim at identifying subgroup- or participant-by-intervention interactions using appropriate study designs [@heckstedenIndividualResponseExercise2015].
-Where variance effects are limited this instead suggests that translational work towards generalisable implementation might be the most meaningful line of future research.
-Finally, there may be cases where it is in fact desirable to identify interventions that actually reduce variance; for example, improvements in methodological approaches to enhance research [@usuiMetaanalysisVariationSuggests2021], or interventions to reduce variation in sport performances.
-Thus, researchers in sport and exercise science should consider exploring variance more systematically, and indeed utilise the meta-analytic framework to support this.
-This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variances utilising the established approaches [@nakagawaMetaanalysisVariationEcological2015; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] presented here and echoing the efforts of other recent work [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
-\textcolor{red}{Researchers should be aware though that the meta-analytic modelling of variance, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make.
-Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.}
+Whether there is evidence of meaningful average intervention effects or not, where considerable variation effects are present it suggests that a meaningful line of research would be to aim at identifying subgroup- or participant-by-intervention interactions using appropriate study designs [@heckstedenIndividualResponseExercise2015; @robinsonOptimizingResearchMethodology2023].
+Where variation effects are limited this instead suggests that translational work towards generalisable implementation might be the most meaningful line of future research.
+Finally, there may be cases where it is in fact desirable to identify interventions that actually reduce variation; for example, improvements in methodological approaches to enhance research [@usuiMetaanalysisVariationSuggests2021], or interventions to reduce variation in sport performances.
+Thus, researchers in sport and exercise science should consider exploring variation more systematically, and indeed utilise the meta-analytic framework to support this.
+This could include the re-analysis of past meta-analyses as we have done here, and indeed researchers conducting future meta-analyses in the field of sport and exercise science should consider the value of concomitantly exploring means and variation utilising the established approaches [@nakagawaMetaanalysisVariationEcological2015; @hopkinsIndividualResponsesMade2015; @atkinsonTrueFalseInterindividual2015; @atkinsonIssuesDeterminationResponders2019; @usuiMetaanalysisVariationSuggests2021; @millsDetectingHeterogeneityIntervention2021] presented here and echoing the efforts of other recent work [@kelleyAreThereInterindividual2022; @kelleyAreThereInterIndividual2020; @estevesIndividualParticipantData2021; @bonafigliaInterindividualDifferencesTrainability2022; @steeleSlowSteadyHard2021; @fisherRoleSupervisionResistance2022].
+Researchers should be aware though that the meta-analytic modelling of variation, or indeed any effect size statistics, requires careful consideration of the research question and assumptions that different models make.
+Substantive conclusions may be sensitive to subtle differences in the approaches we have presented.
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::

### 2023-11-13 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Second round of revisions

@@
 Thus, in a field with small samples such as sport and exercise science, exploration of variance through a meta-analytic framework is appealing.
-Despite the value of embracing and exploring variation alongside mean effects in sport and exercise science it is rarely applied to research synthesis through meta-analysis.We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs."
+Despite the value of embracing and exploring variation alongside mean effects in sport and exercise science it is rarely applied to research synthesis through meta-analysis.
+We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs."
 thanks: |

### 2023-11-13 paper/Revised Paper 2 - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Second round of revisions

--- 2023-11-13 paper/Revised Paper - Meta-Analysis of Variation in Sport and Exercise Science - anonymous.Rmd
@@
 subtitle: "Examples of Application Within Resistance Training Research"
+author: 
+  - James Steele$^a$
+  - James Fisher$^a$
+  - Dave Smith$^b$
+  - Brad Schoenfeld$^c$
+  - Yefeng Yang$^d$
+  - Shinichi Nakagawa$^d$
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 Despite the value of embracing and exploring variation alongside mean effects in sport and exercise science it is rarely applied to research synthesis through meta-analysis.We introduce and evaluate different effect size calculations along with models for meta-analysis of variation using relatable examples from resistance training RCTs."
+thanks: |
+  Address for correspondence: james.steele@solent.ac.uk
+  Affiliations:
+  $^a$Faculty of Sport, Health, and Social Sciences, Solent University, UK;
+  $^b$Research Centre for Musculoskeletal and Sports Medicine, Manchester Metropolitan University, Manchester, UK;
+  $^c$Health Sciences Department, CUNY Lehman College, Bronx, New York, USA;
+  $^d$Evolution & Ecology Research Centre and School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, New South Wales, Australia
 output: 
@@
    - \captionsetup[figure]{font=small}
-   - \usepackage{lineno}
-   - \linenumbers
 ---
+
+------------------------------------------------------------------------
 

### 2023-11-13 paper/Revised Pre-print 2 - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
Second round of revisions

--- 2023-11-13 paper/Revised Paper 2 - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
@@
 thanks: |
+  Preprint, please cite as: Steele, J., Fisher, J., Smith, D., Schoenfeld, B., Yang, R., & Nakagawa, S.
+(2022).
+Meta-Analysis of Variation in Sport and Exercise Science: Examples of Application Within Resistance Training Research.
+$\textrm{sportr}\chi\textrm{iv}$ DOI: https://doi.org/10.51224/SRXIV.214.
   Address for correspondence: james.steele@solent.ac.uk

### 2023-11-13 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -1128,16 +1128,15 @@ model_mean_variance_pre_model_comparisons
 
 ggsave("plots/mean_variance_pre_model_comparisons.tiff", width = 15, height = 7.5, device = "tiff", dpi = 300)
 
-
 ### Meta-analytic scatter plot
 
 # get the predicted log values
 Data_long_pre <- Data_long_pre %>%
   filter(!is.na(SD_log))
 
-Data_long_pre <- cbind(Data_long_pre, pred = predict(RobuEstMultiLevelModel_ri_only_log_mean_variance)$pred,
-                      ci.lb =  predict(RobuEstMultiLevelModel_ri_only_log_mean_variance)$ci.lb,
-                      ci.ub =  predict(RobuEstMultiLevelModel_ri_only_log_mean_variance)$ci.ub) %>%
+Data_long_pre <- cbind(Data_long_pre, pred = predict(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_variance)$pred,
+                      ci.lb =  predict(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_variance)$ci.lb,
+                      ci.ub =  predict(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_variance)$ci.ub) %>%
   mutate(wi = 1/sqrt(SD_log_vi),
          size = 0.5 + 3.0 * (wi - min(wi))/(max(wi) - min(wi))) 
 
@@ -1627,7 +1626,7 @@ ggsave("plots/mean_mod_strength_model_comparisons.tiff", width = 15, height = 7.
 ### Meta-analytic scatter plot
 
 # get the predicted log values
-Data_long_strength_log <- cbind(Data_long_strength, predict(RobuEstMultiLevelModel_ri_only_log_mean_mod_strength )) %>%
+Data_long_strength_log <- cbind(Data_long_strength, predict(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength )) %>%
     mutate(wi = 1/sqrt(SD_log_vi),
            size = 0.5 + 3.0 * (wi - min(wi))/(max(wi) - min(wi)))
 
@@ -1843,7 +1842,7 @@ ggsave("plots/mean_mod_hypertrophy_model_comparisons.tiff", width = 15, height =
 ### Meta-analytic scatter plot
 
 # get the predicted log values
[…]

### 2023-11-18 paper/Final Paper Word Doc - Meta-Analysis of Variation in Sport and Exercise Science.qmd
Render docx file and prep figures for journal publication

--- 2023-11-13 paper/Revised Paper 2 - Meta-Analysis of Variation in Sport and Exercise Science.Rmd
@@
   $^d$Evolution & Ecology Research Centre and School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, New South Wales, Australia
-output: 
-  bookdown::pdf_document2:
-    extra_dependencies: ["flafter"]
+format: 
+  docx:
+    toc: false
+    number-sections: true
+    highlight-style: github
+    prefer-html: true
 bibliography: references.bib
 csl: apa.csl
-toc: false
 header-includes:
@@
 
-------------------------------------------------------------------------
-
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.align="center")
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 options(knitr.kable.NA = '')
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
+```{r}
 # Load relevant packages
@@
 
-```{r, message=FALSE,warning=FALSE,echo=FALSE}
+```{r}
 # load the raw data to the environment
@@
 Although the quantitative synthesis of results across studies has existed since the 17th century [@plackettStudiesHistoryProbability1958], the modern-day term "meta-analysis" was coined by Gene Glass [-@glassPrimarySecondaryMetaAnalysis1976].
-Since that time, the use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@haggerMetaanalysis2022], with resistance training (RT) accounting for a considerable proportion of this growth (figure \@ref(fig:trend-plot)).
+Since that time, the use of meta-analysis as a tool for the synthesis of research in sport and exercise science has increased considerably [@haggerMetaanalysis2022], with resistance training (RT) accounting for a considerable proportion of this growth (@fig-trend-plot).
 Accordingly, throughout the paper we use RT studies as a hopefully familiar example for sport and exercise science researchers.
@@
 
-```{r trend-plot, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Trends in meta-analyses published in sport and exercise science since 1976."}
+```{r}
+#| label: fig-trend-plot 
+#| fig-width: 6
+#| fig-height: 4
+#| fig-cap: Trends in meta-analyses published in sport and exercise science since 1976.
 
@@
 
-```{=tex}
+$$
 \begin{equation}
 \hat\theta_{i}=\theta+m_{i} 
-(\#eq:fixed-model-eq)
 \end{equation}
-```
+$$ {#eq-fixed-model}
+
 where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta$ is the intercept or overall mean (i.e., the fixed-effect), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
@@
 
-```{=tex}
+$$
 \begin{equation}
 w_{i}=\frac{1}{s^2_{i}}
-(\#eq:fixed-weights-eq)
 \end{equation}
-```
+$$ {#eq-fixed-weights}
 Where $s_{i}$ is the standard deviation for the effect estimate and thus $s^2_{i}$ is the variance.
@@
 
-```{=tex}
+$$
 \begin{equation}
 \hat\theta=\frac{\sum^I_{i=1}\hat\theta_{i}w_{i}}{\sum^I_{i=1}w_{i}}
-(\#eq:fixed-average-eq)
 \end{equation}
-```
+$$ {#eq-fixed-average}
 Contrastingly, the random effects model does not make the assumption that there is only one fixed-effect.
@@
 
-```{=tex}
+$$
 \begin{equation}
 \hat\theta_{i}=\theta_{i}+\tau_{i}+m_{i}
-(\#eq:random-model-eq)
 \end{equation}
-```
+$$ {#eq=random-model}
+
 where $\hat\theta_{i}$ is the $i\textrm{th}$ effect size ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies and thus effect sizes), $\theta_i$ is the intercept or overall mean of the effects across each study, $\tau_{i}$ is the deviation from $\theta_{i}$ for the $i\textrm{th}$ study (i.e., the random-effects), and $m_{i}$ is the sampling error for each effect size normally distributed with $\sigma^2_{m_{i}}$.
@@
 
-```{=tex}
+$$
 \begin{equation}
 w^*_{i}=\frac{1}{s^2_{i}+\tau^2}
-(\#eq:random-weights-eq)
 \end{equation}
-```
-Then the average of the distribution of effects can be calculated as per equation \@ref(eq:fixed-average-eq) substituting $w_{i}$ for $w^*_{i}$.
+$$ {#eq-random-weights}
+
+Then the average of the distribution of effects can be calculated as per @eq-fixed-average substituting $w_{i}$ for $w^*_{i}$.
 
@@
 
-```{=tex}
+$$
 \begin{equation}
 SMD=\frac{\overline{x}_E - \overline{x}_C}{s_{pooled}}J
-(\#eq:SMD-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-SMD}
+
+$$
 \begin{equation}
 J=1-\frac{3}{4(n_{C}+n_{E})-2)-1}
-(\#eq:biasJ-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-biasJ}
+$$
 \begin{equation}
 s_{pooled}=\sqrt{\frac{(n_{C}-1)s^2_{C}+(n_{E}-1)s^2_{E}}{n_{C}+n_{E}-2}}
-(\#eq:s-pooled-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-s-pooled}
+
+$$
 \begin{equation}
 s^2_{SMD}=\frac{n_{C}+n_{E}}{n_{C}n_{E}}+\frac{SMD^2}{2(n_{E}+n_{C})}
-(\#eq:SMDvar-eq)
 \end{equation}
-```
+$$ {#eq-SMDvar}
 where $\overline{x}_C$ and $\overline{x}_E$ are the sample means of the control group (C) and experimental (E) or intervention group respectively, $s_C$ and $s_E$ are the standard deviations of the two groups, $n_C$ and $n_E$ are the sample sizes of the two groups, and $J$ is a bias correction for small sample sizes.
@@
 
-```{=tex}
+$$
 \begin{equation}
 \textrm{ln}RR=\textrm{ln}\frac{\overline{x}_E}{\overline{x}_C} + \frac{1}{2}(\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}-\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}})
-(\#eq:lnRR-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-lnRR}
+$$
 \begin{equation}
 s^2_{\textrm{ln}RR}=\frac{s^2_{C}}{n_{C}\overline{x}^2_{C}}+\frac{s^4_{C}}{2n_{C}\overline{x}^2_{C}\overline{x}^4_{C}}+\frac{s^2_{E}}{n_{E}\overline{x}^2_{E}}+\frac{s^4_{E}}{2n_{E}\overline{x}^2_{E}\overline{x}^4_{E}}
-(\#eq:lnRRvar-eq)
 \end{equation}
-```
+$$ {#eq-lnRRvar}
 Due to its calculation the SMD is affected not only by the difference in means of the two groups, but also by the standard deviations of both groups due to the standardisation of the effect size by $s_{pooled}$ in the denominator.
-In contrast, the $\textrm{ln}RR$ is uninfluenced in magnitude by the standard deviations in either groups except for the bias correction factor (see second term in equation \@ref(eq:lnRR-eq)), and the sampling variance (see equation \@ref(eq:lnRRvar-eq)).
+In contrast, the $\textrm{ln}RR$ is uninfluenced in magnitude by the standard deviations in either groups except for the bias correction factor (see second term in @eq-lnRR), and the sampling variance (see @eq-lnRRvar).
 Despite this, the use of effect sizes like the $\textrm{ln}RR$ has been limited in previous meta-analyses in sport and exercise science [@debQuantifyingEffectsAcute2018; @nuzzoEccentricConcentricStrength2023] and to our knowledge only a couple of meta-analyses of RT interventions has used this kind of effect size [@swintonInterpretingMagnitudeChange2022; @wolfPartialVsFull2023].
@@
 
-```{=tex}
+$$
 \begin{equation}
 SD_{ir}=\sqrt{s^2_{E} - s^2_{C}}
-(\#eq:SDir-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-SDir}
+$$
 \begin{equation}
 s^2_{SD_{ir}}=2\biggl(\frac{s^4_{E}}{n_{E}-1}+\frac{s^4_{C}}{n_{C}-1}\biggl)
-(\#eq:sdIRvar-eq)
 \end{equation}
-```
+$$ {#eq-sdIRvar}
 Thus, the $SD_{ir}$ reflects a comparison of the absolute variation in change scores between control and intervention groups.
@@
 
-```{=tex}
+$$
 \begin{equation}
 \textrm{ln}\hat\sigma=\textrm{ln}s+\frac{1}{2(n-1)}
-(\#eq:lnsigma-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-lnsigma}
+$$
 \begin{equation}
 s^2_{\textrm{ln}\hat\sigma}=\frac{1}{2(n-1)}
-(\#eq:lnsigmavar-eq)
 \end{equation}
-```
+$$ {#eq-lnsigmavar}
+
 where $\textrm{ln}\hat\sigma$ is an estimate of $\textrm{ln}\sigma$, and it is assumed with sufficiently large sample size and value of $\sigma$ that $\textrm{ln}\sigma$ is normally distributed with variance $s^2_{\textrm{ln}\sigma}$.
-Given equations \@ref(eq:lnsigma-eq) and \@ref(eq:lnsigmavar-eq), an unbiased estimator of the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
+Given equations @eq-lnsigma and @eq-lnsigmavar, an unbiased estimator of the logarithm of the ratio of standard deviations of two groups, such as a control and intervention, the $\textrm{ln}VR$, and its sampling variance, $s^2_{\textrm{ln}VR}$ is given by:
 
-```{=tex}
+$$
 \begin{equation}
 \textrm{ln}VR=\textrm{ln}\biggl(\frac{s_{E}}{s_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
-(\#eq:lnVR-eq)
 \end{equation}
-```
-```{=tex}
+$$ {#eq-lnVR}
+$$
 \begin{equation}
 s^2_{\textrm{ln}VR}=\frac{1}{2}\biggl(\frac{n_{C}}{(n_{C}-1)^2}+\frac{n_{E}}{(n_{E}-1)^2}\biggl)
-(\#eq:lnVRvar-eq)
 \end{equation}
-```
+$$ {#eq-lnVRvar}
 However, due to both $SD_{ir}$ and $\textrm{ln}VR$ being comparisons of absolute variance, they may find limited applicability where the mean of one group is larger than the comparison group (e.g., when $\overline{x}_E$ is larger than $\overline{x}_C$).
@@
 The natural logarithm of the ratio between the coefficients of variation from two groups, the $\textrm{ln}CVR$ is thus a more generally applicable effect size statistic for examining variability between groups.
-Considering equations \@ref(eq:lnRR-eq) and \@ref(eq:lnVR-eq), an unbiased estimator of the $\textrm{ln}CVR$ is given by:
+Considering equations @eq-lnRR and @eq-lnVR, an unbiased estimator of the $\textrm{ln}CVR$ is given by:
 
-```{=tex}
+$$
 \begin{equation}
 \textrm{ln}CVR=\textrm{ln}\biggl(\frac{CV_{E}}{CV_{C}}\biggl)+\frac{1}{2(n_{E}-1)}-\frac{1}{2(n_{C}-1)}
-(\#eq:lnCVR-eq)
 \end{equation}
-```
+$$ {#eq-lnCVR}
 where $CV_{E}$ and $CV_{C}$ are ${s_{E}}/{\overline{x}_{E}}$ and ${s_{C}}/{\overline{x}_{C}}$ respectively.
@@
 
-```{=tex}
+$$
 \begin{equation}
@@
 \end{split}
-(\#eq:lnCVRvar-eq)
 \end{equation}
-```
+$$ {#eq-lnCVRvar}
 # Examples using resistance training studies
@@
 
-```{r sample-size-tab, message=FALSE, warning=FALSE, echo=FALSE}
+```{r}
+#| label: tbl-sample-size 
+#| tbl-cap-location: "top"
+#| tbl-cap: Sample sizes for resistance training and non training control groups for dataset.
 
@@
 knitr::kable(
-  sample_sizes,
-  caption = 'Sample sizes for resistance training and non training control groups for dataset.'
+  sample_sizes
   ) %>%
@@
 [-@politoModeratorsStrengthGains2021] included only studies with a non-training control group, their study selection offers a unique context to examine variation of interindividual responses specifically by comparing the variation in change scores between the RT intervention groups(s) and control group.
-Table \@ref(tab:sample-size-tab) shows the total sample size, along with the median and range by group, across the included studies.
+@tbl-sample-size shows the total sample size, along with the median and range by group, across the included studies.
 Indeed, this highlights the typically small samples used in sport and exercise science, and thus low power to detect difference in both means and variation in in individual studies [@yangLowStatisticalPower2022], emphasising the value of meta-analysis to explore variation.
-Table \@ref(tab:summary-characteristics-tab) shows the study and participant characteristics.
+@tbl-summary-characteristics shows the study and participant characteristics.
 
@@
 
-```{r summary-characteristics-tab, message=FALSE, warning=FALSE, echo=FALSE}
+```{r}
+#| label: tbl-summary-characteristics 
+#| tbl-cap-location: "top"
+#| tbl-cap: Summary of study and participant characteristics.
 
@@
   summary_table,
-  caption = 'Summary of study and participant characteristics.',
   align = c("l","c") 
@@
 [^14]: It is worth noting that in the sport and exercise sciences, similarly to other fields that examine the effects of experimental interventions, the most common study design for testing or estimating intervention effects is the randomised pretest-postest-control design (i.e., an intervention and control, or other intervention, group randomly allocated and measured pre- and post-exposure).
-We presented the SMD and $\textrm{ln}RR$ effect sizes in equations \@ref(eq:SMD-eq) and \@ref(eq:lnRR-eq) merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., [-@gurevitchInteractionCompetitionPredation2000]; Morris et al., [-@morrisDirectInteractiveEffects2007]; Morris [-@morrisEstimatingEffectSizes2008]; Lajeunesse [-@lajeunesseMetaanalysisResponseRatios2011; -@lajeunesseBiasCorrectionLog2015]) and these are the effect sizes used in the meta-analyses referred to here.
+We presented the SMD and $\textrm{ln}RR$ effect sizes in equations @eq-SMD and @eq-lnRR merely for simplicity in the introduction, but note that extension of these for such 2x2 (i.e., condition x time) study designs have been presented in detail elsewhere (see: Gurevitch et al., [-@gurevitchInteractionCompetitionPredation2000]; Morris et al., [-@morrisDirectInteractiveEffects2007]; Morris [-@morrisEstimatingEffectSizes2008]; Lajeunesse [-@lajeunesseMetaanalysisResponseRatios2011; -@lajeunesseBiasCorrectionLog2015]) and these are the effect sizes used in the meta-analyses referred to here.
 
@@
 
-```{=tex}
+$$
 \begin{equation}
 \hat\theta_{ijk}=(\theta+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+m_{ijk}
-(\#eq:ml-model-eq)
 \end{equation}
-```
+$$ {#eq-ml-model}
 where $\hat\theta_{ijk}$ is the $k\textrm{th}$ effect size, here the SMD or $\textrm{ln}RR$, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\theta$ is the intercept or overall mean of the effects, $\tau_{i}$ is the deviation from $\theta$ for the $i\textrm{th}$ study, $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\theta_{ijk}}$.
@@
 
-```{r forest-SMD-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SMD effect sizes for strength (A) and hypertrophy (B) outcomes."}
-
-# "Caterpillar plots of SMD effect sizes for strength and hypertrophy outcomes.
-Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+```{r}
+#| label: fig-forest-SMD-plot 
+#| fig-width: 10
+#| fig-height: 10
+#| fig-cap: Caterpillar plots of SMD effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval.
 
@@
 
-The pattern of results from our models examining SMDs (figure \@ref(fig:forest-SMD-plot)) were similar to those reported by Polito et al.
+The pattern of results from our models examining SMDs (@fig-forest-SMD-plot) were similar to those reported by Polito et al.
 [-@politoModeratorsStrengthGains2021], albeit with slightly lower estimates for both outcome types; possibly due to our use of a multilevel mixed-effects meta-analysis model that allowed for each individual effect size to be more appropriately weighted (the relative amount of heterogeneity between effects for each level is presented as the $I^2$ statistic).
 As might be expected, in comparison to non-training controls the RT interventions produced increases in both strength (SMD = `r SMD_RR_SDir_logVR$Estimate[1]` [95%CI: `r SMD_RR_SDir_logVR$Lower[1]` to `r SMD_RR_SDir_logVR$Upper[1]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[1]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[1]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[1]`%) and hypertrophy outcomes (SMD = `r SMD_RR_SDir_logVR$Estimate[5]` [95%CI: `r SMD_RR_SDir_logVR$Lower[5]` to `r SMD_RR_SDir_logVR$Upper[5]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[5]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[5]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[5]`%).
-Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see figure \@ref(fig:forest-SMD-plot)) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).
+Confidence intervals on the overall effects were precise for both outcomes though prediction intervals, indicating the range over which we might expect future estimates of effects to fall based on this evidence, for SMD estimates (see @fig-forest-SMD-plot) were fairly wide and relative heterogeneity was fairly high mostly as a result of between-study variance (i.e., effects were more similar *within* studies and arms than *between* them).
 
-```{r forest-RR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of exponentiated RR effect sizes for strength (A) and hypertrophy (B) outcomes."}
-
-# "Caterpillar plots of exponentiated RR effect sizes for strength and hypertrophy outcomes.
-Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+```{r}
+#| label: fig-forest-RR-plot 
+#| fig-width: 10
+#| fig-height: 10
+#| fig-cap: Caterpillar plots of exponentiated RR effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval.
 
@@
 For the $\textrm{ln}RR$ results we exponentiated them and converted to percentages to be more interpretable.
-These were similar, with greater proportional increases in strength compared with hypertrophy (figure \@ref(fig:forest-RR-plot)).
+These were similar, with greater proportional increases in strength compared with hypertrophy (@fig-forest-RR-plot).
 Increases were seen for both strength ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[2]` [95%CI: `r SMD_RR_SDir_logVR$Lower[2]` to `r SMD_RR_SDir_logVR$Upper[2]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[2]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[2]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[2]`%) and hypertrophy ($\textrm{exp}RR$ = `r SMD_RR_SDir_logVR$Estimate[6]` [95%CI: `r SMD_RR_SDir_logVR$Lower[6]` to `r SMD_RR_SDir_logVR$Upper[6]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[6]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[6]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[6]`%).
@@
 
-In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (figure \@ref(fig:forest-SDir-plot)) and $\textrm{ln}VR$ (figure \@ref(fig:forest-logVR-plot)) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[3]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[4]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[7]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[8]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[8]`%) indicating that exposure to the RT interventions may have introduced additional variation over and above random error, potentially suggesting the presence of interindividual response variation.
+In addition to the SMD and $\textrm{ln}RR$ results, both the $SD_{ir}$ (@fig-forest-SDir-plot) and $\textrm{ln}VR$ (@fig-forest-logVR-plot) were also positive for both strength ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[3]` [95%CI: `r SMD_RR_SDir_logVR$Lower[3]` to `r SMD_RR_SDir_logVR$Upper[3]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[3]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[3]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[3]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[4]` [95%CI: `r SMD_RR_SDir_logVR$Lower[4]` to `r SMD_RR_SDir_logVR$Upper[4]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[4]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[4]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[4]`%) and hypertrophy outcomes ($SD_{ir}$ = `r SMD_RR_SDir_logVR$Estimate[7]` [95%CI: `r SMD_RR_SDir_logVR$Lower[7]` to `r SMD_RR_SDir_logVR$Upper[7]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[7]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[7]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[7]`%; $\textrm{ln}VR$ = `r SMD_RR_SDir_logVR$Estimate[8]` [95%CI: `r SMD_RR_SDir_logVR$Lower[8]` to `r SMD_RR_SDir_logVR$Upper[8]`]; $I^2_{study}$ = `r SMD_RR_SDir_logVR$I.2.study[8]`%, $I^2_{arm}$ = `r SMD_RR_SDir_logVR$I.2.arm[8]`%, $I^2_{effect}$ = `r SMD_RR_SDir_logVR$I.2.effect[8]`%) indicating that exposure to the RT interventions may have introduced additional variation over and above random error, potentially suggesting the presence of interindividual response variation.
 Although, heterogeneity across the models and levels (study, arm, effect) were again relatively large and quite varied.
@@
 
-```{r forest-SDir-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of SDir effect sizes for strength (A) and hypertrophy (B) outcomes."}
-
-# "Caterpillar plots of SDir effect sizes for strength and hypertrophy outcomes.
-Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+```{r}
+#| label: fig-forest-SDir-plot 
+#| fig-width: 10
+#| fig-height: 10
+#| fig-cap: Caterpillar plots of SDir effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval.
 
@@
 
-```{r forest-logVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$VR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}
-
-# "Caterpillar plots of $\textrm{ln}RR$ effect sizes for strength and hypertrophy outcomes.
-Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+```{r}
+#| label: fig-forest-logVR-plot 
+#| fig-width: 10
+#| fig-height: 10
+#| fig-cap: Caterpillar plots of $\textrm{ln}RR$ effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval.
 
@@
 
-```{r mean-variation-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes."}
+```{r}
+#| label: fig-mean-variation-pre-plot 
+#| fig-width: 10
+#| fig-height: 8
+#| fig-cap: Scatter plots of raw mean and standard deviation of pre-intervention scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of pre-intervention scores for (C) strength outcomes and (D) hypertrophy outcomes.
 
@@
 
-As can be seen in figure \@ref(fig:mean-variation-pre-plot)(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
+As can be seen in @fig-mean-variation-pre-plot(A) and (C), there is considerable heteroskedasticity in the relationship between the raw mean ($\overline{x}$) and standard deviation ($s$).
 The range of standard deviations increases with higher mean values.
@@
 
-```{=tex}
+$$
 \begin{equation}
 s^2=a\overline{x}^b
-(\#eq:taylors-eq)
 \end{equation}
-```
+$$ {#eq-taylors}
 where $a$ and $b$ are some constants.
@@
 
-```{=tex}
+$$
 \begin{equation}
 2\textrm{ln}s=\textrm{ln}a+b\textrm{ln}\overline{x}
-(\#eq:taylorsln-eq)
 \end{equation}
-```
-Figure \@ref(fig:mean-variation-pre-plot)(B) and (D) shows that the relationship between the mean and variation on the log scale better meets the assumption of normality.
+$$ {#eq-taylorsln}
+@fig-mean-variation-pre-plot(B) and (D) shows that the relationship between the mean and variation on the log scale better meets the assumption of normality.
 Given these the observations we have for $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$ come from multiple outcomes within multiple arms within studies we can also estimate this relationship using a multilevel mixed-effects meta-regression model similar to that applied above.
@@
 
-```{=tex}
+$$
 \begin{equation}
 \textrm{ln}\hat\sigma_{ijk}=(\beta_{0}+\tau_{(1)i}+\tau_{(2)j}+\tau_{(3)k})+\beta_{1}\textrm{ln}\overline{x}_{ijk}+\epsilon_{ijk}+m_{ijk} 
-(\#eq:lnm-lns-model-eq)
 \end{equation}
-```
-where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in equation \@ref(eq:lnsigma-eq), from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^18]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept or overall mean of the effects, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study, $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
+$$ {#eq-lnm-lns-model}
+where $\textrm{ln}\hat\sigma_{ijk}$ is the $k\textrm{th}$ effect size, as in @eq-lnsigma, from the $j\textrm{th}$ arm ($j = 1,2,\cdots,N_{j}$; where $N_{j}$ is the number of arms[^18]) in the $i\textrm{th}$ study ($i = 1,2,\cdots,N_{i}$; where $N_{i}$ is the number of studies), $\textrm{ln}\overline{x}_{ijk}$ is the mean estimate for each effect size, $\beta_{0}$ is the intercept or overall mean of the effects, $\beta_{1}$ is the slope or regression coefficient for $\textrm{ln}\overline{x}$, $\tau_{i}$ is the deviation from $\beta_{0}$ for the $i\textrm{th}$ study, $\tau_{j}$ is the deviation for the $j\textrm{th}$ arm, $\tau_{k}$ is the deviation for the $k\textrm{th}$ effect, $\epsilon_{ijk}$ is the residual for each effect size which is normally distributed with $\sigma^2_{\epsilon}$, and $m_{ijk}$ is the sampling error for each effect size normally distributed with $\sigma^2_{\textrm{ln}\hat\sigma_{ijk}}$.
 
@@
 
-Figure \@ref(fig:model-mean-variation-pre-plot) shows this model (note, see $\textrm{footnote}^{19}$) fit visually where the size of the points reflects their weight in the model.
+@fig-model-mean-variation-pre-plot shows this model (note, see $\textrm{footnote}^{19}$) fit visually where the size of the points reflects their weight in the model.
 Both strength and hypertrophy outcomes show strong linearity between the mean and standard deviation on the log scale, though there is a small difference in intercepts between the two outcome types suggesting a slight but systematically greater degree of variation in strength measures compared to hypertrophy for a given mean score.
 
-```{r model-mean-variation-pre-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=6, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores."}
-
-# "Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores.
-Note, fitted lines are the regression slopes for strength and hypertophy outcomes each with their 95% confidence interval band."
+```{r}
+#| label: fig-model-mean-variation-pre-plot
+#| fig-width: 6
+#| fig-height: 4
+#| fig-cap: Meta-analytic scatter plot of the log mean and log standard deviation of pre-intervention scores.
+Note, fitted lines are the regression slopes for strength and hypertophy outcomes each with their 95% confidence interval band.
 
@@
 
-As can be seen in figures \@ref(fig:mean-variation-delta-plot)(A) and (C) there is also a mean-variation relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
-Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures \@ref(fig:mean-variation-delta-plot)(B) and (D) that the log transformation exhibits similar linearity as seen with the pre-intervention scores above.
+As can be seen in figures @fig-mean-variation-delta-plot(A) and (C) there is also a mean-variation relationship in the change score data about zero whereby an increase in the mean alone (i.e., greater mean change score in the intervention compared to the control) may be fully responsible for any apparent increase in variation.
+Further, when transforming change scores to absolute changes (i.e., converting all to positive numeric scores) we see that in figures @fig-mean-variation-delta-plot(B) and (D) that the log transformation exhibits similar linearity as seen with the pre-intervention scores above.
 As such, in this case, we cannot rely solely on absolute comparisons of variation such as the $SD_{ir}$ and $\textrm{ln}VR$ to determine whether interindividual response variation is actually present.
 
-```{r mean-variation-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=8, fig.width=10, fig.cap="Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes."}
+```{r}
+#| label: fig-mean-variation-delta-plot
+#| fig-width: 10
+#| fig-height: 8
+#| fig-cap: Scatter plots of raw mean and standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes, and of the log mean and log standard deviation of change scores for (C) strength outcomes and (D) hypertrophy outcomes.
 
@@
 The $\textrm{ln}CVR$ can be used to overcome this issue though.
-Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before (see equation \@ref(eq:ml-model-eq)) using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variation comparisons using $SD_{ir}$ or $\textrm{ln}VR$.
-The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[28]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[28]`%) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[88]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[88]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[88]`%) and further there is lower relative heterogeneity between studies in the effect estimates (figure \@ref(fig:forest-logCVR-plot)).
+Fitting the same multilevel mixed-effects meta-analysis model with cluster-robust variance estimation and random intercepts for study, arm, and effect as before (see @eq-ml-model) using the $\textrm{ln}CVR$ as the effect size statistic leads to different conclusions compared to absolute variation comparisons using $SD_{ir}$ or $\textrm{ln}VR$.
+The introduction of an RT intervention actually *reduces* the relative variation seen in change scores for strength ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[28]` [95%CI: `r mods_SMD_logCVR$Lower[28]` to `r mods_SMD_logCVR$Upper[28]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[28]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[28]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[28]`%) and hypertrophy ($\textrm{ln}CVR$ = `r mods_SMD_logCVR$Estimate[88]` [95%CI: `r mods_SMD_logCVR$Lower[88]` to `r mods_SMD_logCVR$Upper[88]`]; $I^2_{study}$ = `r mods_SMD_logCVR$I.2.study[88]`%, $I^2_{arm}$ = `r mods_SMD_logCVR$I.2.arm[88]`%, $I^2_{effect}$ = `r mods_SMD_logCVR$I.2.effect[88]`%) and further there is lower relative heterogeneity between studies in the effect estimates (@fig-forest-logCVR-plot).
 
-```{r forest-logCVR-plot, message=FALSE, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, fig.cap="Caterpillar plots of ln$CVR$ effect sizes for strength (A) and hypertrophy (B) outcomes."}
-
-# "Caterpillar plots of $\textrm{ln}CVR$ effect sizes for strength and hypertrophy outcomes.
-Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval"
+```{r}
+#| label: fig-forest-logCVR-plot
+#| fig-width: 10
+#| fig-height: 10
+#| fig-cap: Caterpillar plots of $\textrm{ln}CVR$ effect sizes for strength and hypertrophy outcomes.
+Note, each estimate and 95% confidence interval is a single observed effect size within each arm within each study, the diamond reflects the model overall estimate and 95% confidence interval, and the horizontal line with vertical bars at the end reflect the 95% prediction interval.
 
@@
 Secondly, whilst the $\textrm{ln}CVR$ is useful in situations where there is a mean-variation relationship, the use of the $CV$ in the effect size statistic assumes proportionality between standard deviation and mean.
-Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (figure \@ref(fig:mean-variation-delta-plot)) an alternative yet comparable approach might be desirable.
+Where we see the kind of heteroskedasticity in the relationship between mean and standard deviation as we do for the change scores here (@fig-mean-variation-delta-plot) an alternative yet comparable approach might be desirable.
 Lastly, this statistic is limited to examination of pairwise comparisons of variation.
@@
 
-```{r, message=FALSE,warning=FALSE,echo=FALSE}
+```{r}
 # load the summary dataframes to the environment
@@
 
-Instead, we can use a version of the meta-regression model described above (see equation \@ref(eq:lnm-lns-model-eq) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
+Instead, we can use a version of the meta-regression model described above (see @eq-lnm-lns-model) and the paragraph which followed it) to compare the variability in change scores between intervention and control groups using $\textrm{ln}\hat\sigma$ and $\textrm{ln}\overline{x}$.
 In this case, the categorical variable for the outcome type used previously is instead swapped for the group type and the new model term included becomes $\beta_{2}Group$ with $Group$ as a dummy coded variable for the group (i.e., non-training control = 0, and RT intervention = 1), where $\beta_{2}$ is the slope or regression coefficient for $Group$.
@@
 
-```{r model-mean-variation-delta-plot, message=FALSE, warning=FALSE, echo=FALSE,fig.height=4, fig.width=8, fig.cap="Meta-analytic scatter plot of the log mean and log standard deviation of change scores."}
-
-# "Meta-analytic scatter plot of the log mean and log standard deviation of change scores.
-Note, fitted lines are the regression slopes for control group (CON) and resistance training intervention (RT) groups each with their 95% confidence interval band."
+```{r}
+#| label: fig-model-mean-variation-delta-plot
+#| fig-width: 8
+#| fig-height: 4
+#| fig-cap: Meta-analytic scatter plot of the log mean and log standard deviation of change scores for (A) strength outcomes and (B) hypertrophy outcomes.
+Note, fitted lines are the regression slopes for control group (CON) and resistance training intervention (RT) groups each with their 95% confidence interval band.
 
@@
 
-Given the heteroskedasticity in the change scores means and standard deviations (see figure \@ref(fig:mean-variation-delta-plot)), we fit this model and other model specifications[^20]\textsuperscript{,}[^21] to the dataset (see figure \@ref(fig:model-mean-variation-delta-plot)).
+Given the heteroskedasticity in the change scores means and standard deviations (see @fig-mean-variation-delta-plot), we fit this model and other model specifications[^20]\textsuperscript{,}[^21] to the dataset (see @fig-model-mean-variation-delta-plot).
 The results from these models (note, see $\textrm{footnote}^{20}$) were largely similar to those found using the $\textrm{ln}CVR$ model for strength ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_arm_log_mean_mod_strength$ci.ub[3],2)`]) and hypertrophy ($\beta_{\textrm{ln}\hat\sigma[Group \textrm{ for RT}]}$ = `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy$b[3],2)` [95%CI: `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy$ci.lb[3],2)` to `r round(RobuEstMultiLevelModel_ri_mean_group_slope_study_log_mean_mod_hypertrophy$ci.ub[3],2)`]).
@@
 
-[^23]: Indeed, it can be seen from figures \@ref(fig:forest-SDir-plot), \@ref(fig:forest-logVR-plot), and \@ref(fig:forest-logCVR-plot) that many of the individual study effect estimates have very large sampling errors.
+[^23]: Indeed, it can be seen from figures @fig-forest-SDir-plot, @fig-forest-logVR-plot, and @fig-forest-logCVR-plot that many of the individual study effect estimates have very large sampling errors.
 

### 2023-11-18 analysis/Analysis - Meta-Analysis of Variation in Sport and Exercise Science.R

@@ -37,7 +37,7 @@ trends <- rbind(SES_meta_trend, RT_meta_trend) %>%
 
 trends_plot <- trends %>%
   ggplot(aes(x = year, y = (query_hits / all_hits)*100, linetype = `Search string`)) +
-  geom_line(size = 1) +
+  geom_line(linewidth = 1) +
   labs(x = "Year", y = "Proportion of Published Articles (%)",
        title = "Articles in Europe Pub Med Central Database") +
   theme_classic() +
@@ -45,6 +45,9 @@ trends_plot <- trends %>%
         legend.text = element_text(size = 6))
 
 save(trends_plot, file = "plots/trends_plot")
+
+ggsave("plots/trends_plot.tiff", width = 6, height = 4, device = "tiff", dpi = 300)
+
 
 ##### Read csv as data frame into environment - Note: change source address
 Data <- read.csv(here::here("data","Polito et al. RT Extracted Data.csv"), na.strings=c(""," ","NA"))
